### 2016-08-22 code/test_code.Rmd
intial test code rmd

---
title: "Test_code"
author: "Nick Lesniak"
date: "August 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r, message=FALSE}
pack_used <- c('randomForest','miscTools','tidyr','ggplot2','dplyr', 'pROC', 'knitr', 'cowplot')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

```{r}
# setup environment

# read in files
meta_file   <- read.table('../data/process/human_CdGF_metadata.txt', sep = '\t',
                        header = T)
shared_file <- read.table('../data/process/human_CdGF.an.unique_list.0.03.subsample.shared',
                          sep='\t',header = T)
tax_file <- read.table('../data/mothur/gf_cdiff.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy',
                       sep='\t',header = T, row.names = 'OTU')
# function to label otu with tax
source('tax_level.R')

unused_samples <- c("P2", "CON", "CON2", "CON3")

# remove samples not used or with other cdiff strains
meta_file <- 
  meta_file %>% 
  filter(sample_type == 'stool',
         !cage_id %in% unused_samples) %>% 
  select(Mouse_ID,Group = group, day, cdiff_cfu, Euthanized) %>% 
  mutate(log_cfu = log10(cdiff_cfu + 1)) %>% 
  filter(!is.na(log_cfu))

# create df with day 0 communities with column of mouse_id
day_0_otu_by_mouse <- 
  meta_file %>% 
  filter(day == 0 ) %>% 
  select(Group, Mouse_ID) %>% 
  left_join(shared_file)
day_0_otu_over_01 <- 
  day_0_otu_by_mouse %>% 
  select(contains('Otu')) %>% 
  apply(. , 2, function(i){mean(i)/mean(rowSums(.))}) > 0.01
day_0_otu_over_01 <-  names(day_0_otu_over_01)[day_0_otu_over_01]
day_0_otu_by_mouse_01 <- 
  day_0_otu_by_mouse %>% 
  select(Mouse_ID, one_of(day_0_otu_over_01))

seed <- 18
n_trees <- 2000
set.seed(seed)

rf_plot_list <- list()
n_features <- 10

rf_plot_list <- 
lapply(1:10, function(i){
rf_data <- 
  meta_file %>%
  filter(day == i) %>% 
  select(log_cfu, Mouse_ID) %>%
  inner_join(day_0_otu_by_mouse_01) %>% 
  select(-Mouse_ID) %>% 
  tbl_df()

rf_all <- randomForest(log_cfu ~ ., data = rf_data,
                   importance = TRUE, ntree = n_trees)
importance_sorted <- sort(importance(rf_all)[,1], decreasing = T)
rf_n_features <- randomForest(log_cfu ~ ., data=rf_data[ ,c( 'log_cfu',
                              names(importance_sorted)[1:n_features])], 
                           importance=TRUE, ntree=n_trees)

predict_plot_day <- paste('prediction_day_', i, sep = '')
importance_plot_day <- paste('importance_day_', i, sep = '')

rf_prediction <- cbind(rf_data[, 'log_cfu'], 
                       data.frame(rf_n_features$predicted))
colnames(rf_prediction) <- c('observed', 'predicted')

predict_plot_day <-  
  ggplot(data = rf_prediction, aes(x=observed, y=predicted)) + geom_point() + 
        labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)') + 
        theme_bw() + ggtitle(paste("Random Forest Prediction of C Diff CFU \n on Day ",
                                   i, " from Day 0 Community", sep = '')) +
        annotate('text', x=0.15*(max(rf_prediction$observed)), 
                 y=0.95*(max(rf_prediction$predicted)), 
                 label=paste('R^2 = ', round(tail(rf_n_features$rsq, 1), 2),'\nMSE = ', 
                             round(tail(rf_n_features$mse, 1), 2), sep = ''))

top_important_OTU <- data.frame(head(importance_sorted, n_features))
colnames(top_important_OTU) <- 'Importance'
top_important_OTU$OTU <- rownames(top_important_OTU)
otu_taxa <- get_tax(1, top_important_OTU$OTU, tax_file)

importance_plot_day <- 
  ggplot(data = top_important_OTU, aes(x = factor(OTU), y = Importance)) + 
        geom_point() + scale_x_discrete(limits = rev(top_important_OTU$OTU),
                                        labels = paste(otu_taxa[,1],' (',
                                                       rownames(otu_taxa),')',
                                                       sep='')) +
        labs(x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()

rf_plot_list[[i]] <- list(predict_plot_day,importance_plot_day)

})
```

### Predict Day 1 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[1]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[1]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 2 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[2]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[2]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 3 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[3]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[3]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 4 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[4]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[4]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 5 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[5]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[5]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 6 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[6]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[6]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 7 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[7]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[7]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 8 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[8]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[8]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 9 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[9]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[9]][[2]], 0.5, 0, 0.5, 1)

```

***

### Predict Day 10 CFU from Day 0 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[10]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[10]][[2]], 0.5, 0, 0.5, 1)

```