### 2016-09-02 code/HumanCdGF_entero_predict.Rmd
initial entero-based prediction

---
title: "humanCdGF_entero_predict"
author: "Nicholas Lesniak"
date: "September 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, fig.width=10, fig.height=5)
```

```{r}
pack_used <- c('randomForest','tidyr','ggplot2','dplyr','knitr', 'cowplot', 'vegan', 'cluster')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

```{r}
# setup environment

# read in files
meta_file   <- read.table('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt', sep = '\t',
                        header = T)
shared_file <- read.table('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.an.unique_list.0.03.subsample.shared',
                          sep='\t',header = T, row.names = 'Group')
tax_file <- read.table('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/mothur/gf_cdiff.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy',
                       sep='\t',header = T, row.names = 'OTU')
# function to label otu with tax
source('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R')

# identify c difficile otus
cdiff_otu <- rownames(tax_file[grep('Clostridium_XI', tax_file$Taxonomy),])
# remove samples not used or with other cdiff strains
meta_file <- 
  meta_file %>% 
  filter(sample_type == 'stool', cdiff_strain == 431) %>% 
  select(Mouse_ID, cage_id, group, day, cdiff_cfu, Euthanized) %>% 
  mutate(log_cfu = log10(cdiff_cfu + 1)) %>% 
  filter(!is.na(log_cfu))

# create enterotypes
# first generate distance matrices
dist.JSD <- function(inMatrix, pseudocount=0.000001, ...) {
	KLD <- function(x,y) sum(x *log(x/y))
	JSD<- function(x,y) sqrt(0.5 * KLD(x, (x+y)/2) + 0.5 * KLD(y, (x+y)/2))
	matrixColSize <- length(colnames(inMatrix))
	matrixRowSize <- length(rownames(inMatrix))
	colnames <- colnames(inMatrix)
	resultsMatrix <- matrix(0, matrixColSize, matrixColSize)

  inMatrix = apply(inMatrix,1:2,function(x) ifelse (x==0,pseudocount,x))

	for(i in 1:matrixColSize) {
		for(j in 1:matrixColSize) { 
			resultsMatrix[i,j]=JSD(as.vector(inMatrix[,i]),
			as.vector(inMatrix[,j]))
		}
	}
	colnames -> colnames(resultsMatrix) -> rownames(resultsMatrix)
	as.dist(resultsMatrix)->resultsMatrix
	attr(resultsMatrix, "method") <- "dist"
	return(resultsMatrix) 
}

shared_dist <- dist.JSD(shared_file)