### 2016-09-02 code/HumanCdGF_entero_predict.Rmd
added enterotype files to rmd

---
title: "humanCdGF_entero_predict"
author: "Nicholas Lesniak"
date: "September 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, fig.width=10, fig.height=5)
```

```{r}
pack_used <- c('randomForest','tidyr','ggplot2','dplyr','knitr', 'cowplot', 'vegan', 'cluster')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

```{r}
# setup environment

# read in files
meta_file   <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt'
shared_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.an.unique_list.0.03.subsample.shared'
tax_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/mothur/gf_cdiff.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy'
entero_list <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.0.03.subsample.shared.for.0.03.dmm.5.mix.posterior'
entero_abund <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.0.03.subsample.shared.for.0.03.dmm.5.mix.relabund'

meta_file   <- read.table(meta_file, sep = '\t', header = T)
shared_file <- read.table(shared_file, sep='\t',header = T, row.names = 'Group')
tax_file <- read.table(tax_file, sep='\t',header = T, row.names = 'OTU')
entero_list <- read.table(entero_list, sep = '\t', header = T)
entero_abund <- read.table(entero_abund, sep = '\t', header = T)

# function to label otu with tax
source('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R')

# identify c difficile otus
cdiff_otu <- rownames(tax_file[grep('Clostridium_XI', tax_file$Taxonomy),])
# remove samples not used or with other cdiff strains
meta_file <- 
  meta_file %>% 
  filter(sample_type == 'stool', cdiff_strain == 431) %>% 
  select(Mouse_ID, cage_id, group, day, cdiff_cfu, Euthanized) %>% 
  mutate(log_cfu = log10(cdiff_cfu + 1)) %>% 
  filter(!is.na(log_cfu))

# create enterotypes
entero_list <-
  unique(meta_file[meta_file$cage_id != 'inoculum',c('human_source','cage_id')])  
  gather(entero_list, 'Partition', 2:6)