### 2016-10-14 code/HumanCdGF_microbiota_strains.Rmd
fix median and axis label

Strain independence
========================================================
``` {r, include = FALSE}
# title: "Strain independence"
# author: "Nicholas Lesniak"
# date: "September 12, 2016"
# output: html_document
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r, message=FALSE}
pack_used <- c('ggplot2','knitr','dplyr')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
non_431   <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt'

# read in files
non_431   <- read.table(non_431, sep = '\t', header = T, row.names = 'sample_id')

# select mice infected with two strains of c diff
cdiff_strains <- unique(non_431$cdiff_strain)
non_431_strains <- cdiff_strains[!unique(cdiff_strains) %in% c('431', NA)]
non_431_inocula <- unique(non_431[non_431$cdiff_strain %in% non_431_strains, 'human_source'])
non_431 <- non_431[non_431$human_source %in% non_431_inocula, ]
```

```{r fig.height= 5, fig.width=10}
non_431 %>% 
  filter(day %in% c(0:10), !is.na(log_cfu) ) %>%
  select(mouse_id, day, human_source, cdiff_strain, cage_id, log_cfu) %>% 
  ggplot(aes(x = factor(day), y = log_cfu, color = factor(cdiff_strain))) + geom_point() +
            stat_summary(fun.y=median, geom="line", aes(group = cage_id)) + 
            facet_grid(. ~ human_source) + theme_bw() + 
            labs(x = 'Day', y = expression(paste(italic('C. Difficile'), ' Log'[10], ' CFU', sep = ''))) +
            theme(legend.justification=c(1,0.5), legend.position=c(1,0.5), 
                  legend.title=element_blank())
```