### 2016-11-17 code/CdGF_AUCRF.Rmd

---
title: "CdGF_aucrf"
author: "Nicholas Lesniak"
date: "September 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
pack_used <- c('randomForest','ggplot2', 'pROC', 'knitr','dplyr','AUCRF', 'tidyr', 'caret')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

## Can Day 0 Community members predict C. difficile CFU?
```{r setup data, fig.height= 10, fig.width=10}

meta_file   <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt'
shared_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.an.unique_list.0.03.subsample.shared'
tax_file <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/gf_new.an.taxonomy'
tax_function <- '~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R'

# read in files
meta_file   <- read.table(meta_file, sep = '\t', header = T, row.names = 'sample_id')
shared_file <- read.table(shared_file, sep = '\t', header = T, row.names = 'sample_id')
tax_file <- read.table(tax_file, sep = '\t', header = T, row.names = 'OTU')
source(tax_function)

# select only mice with 431 cdiff
meta_file <- meta_file[meta_file$cdiff_strain == '431', ]

# remove cage NP1 (no confirmed infection)
meta_file <- meta_file[!meta_file$cage_id == 'NP1', ]

# subset shared
#Create df with relative abundances
stool_shared <- shared_file[rownames(meta_file[meta_file$day %in% 0, ]), ]
# use %in% instead of == because to ignore NA
# rename day 0 community rows by mouse id - since meta rows have day in row names, use mouse id for both
rownames(stool_shared) <- meta_file$mouse_id[rownames(meta_file) %in% 
                                               rownames(stool_shared) ]
rel_abund_stool <- 100 * stool_shared / unique(apply(stool_shared, 1, sum))
#Create vector of OTUs with median abundances >1%
OTUs_stool_0_01 <- apply(rel_abund_stool, 2, max) > 1
#df of OTUs with abundances >1% - by cage and inoculum
rel_abund_stool <- rel_abund_stool[ ,OTUs_stool_0_01]

# remove correlative otus
rel_abund_cor <- cor(rel_abund_stool, method = 'spearman')
cor_cutoff <- 0.75
cor_otus <- findCorrelation(rel_abund_cor, cutoff = cor_cutoff)
rel_abund_stool <- select(rel_abund_stool, -cor_otus)

#random forest regression model to determine Day 0 OTU correlated with increased CFU
seed <- 1
n_trees <- 2001
iters <- 100
```

##Can we predict moribundity with the community structure at Day 0?

```{r roc, fig.width=8, fig.height=8, results= 'hide'}
#create dataframe for RF-classification
meta_file$Euth_Early <- factor(ifelse(meta_file[ , 'Early_Euth'] == FALSE, 0, 1))

Predict_df <- merge(meta_file[meta_file$day %in% 0, 
                              c('human_source', 'mouse_id', 'cage_id', 'Euth_Early')],
                    rel_abund_stool, by.x = 'mouse_id', by.y = 'row.names')

# remove cage/inocula OTUs from prediction dataframes
Predict_early_euth_df <- select(Predict_df, -cage_id, -mouse_id, -human_source)

# create RF model
set.seed(seed)
rf_early_aucrf <- AUCRF(Euth_Early ~ ., data = Predict_early_euth_df,
                              ntree = n_trees, pdel = 0.05, ranking = 'MDA')

otu_euth_probs <- predict(rf_early_aucrf$RFopt, type = 'prob')
otu_euth_roc <- roc(Predict_early_euth_df$Euth_Early ~ otu_euth_probs[ , 2])
euth_otu_feat <- rf_early_aucrf$Xopt

aucrf_data <- Predict_early_euth_df[,c('Euth_Early',euth_otu_feat)]

#### 10-fold Cross validated - 10 repetitions
cv10f_aucs <- c()
cv10f_all_resp <- c()
cv10f_all_pred <- c()
for(j in 1:iters){
  set.seed(j)
  sampling <- sample(1:nrow(aucrf_data),nrow(aucrf_data),replace=F)
  cv10f_probs <- rep(NA,52)
  for(i in seq(1,50,5)){
    train <- aucrf_data[sampling[-(i:(i+4))],]
    test <- aucrf_data[sampling[i:(i+4)],]
    set.seed(seed)
    temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
    cv10f_probs[sampling[i:(i+4)]] <- predict(temp_model$RFopt, test, type='prob')[,2]
  }
  cv10f_roc <- roc(aucrf_data$Euth_Early~cv10f_probs)
  cv10f_all_pred <- c(cv10f_all_pred, cv10f_probs)
  cv10f_all_resp <- c(cv10f_all_resp, aucrf_data$Euth_Early)
  cv10f_aucs[j] <- cv10f_roc$auc
}
cv10f_roc <- roc(cv10f_all_resp~cv10f_all_pred)

#### leave-one-out
LOO_probs <- NULL
LOO_probs <- sapply(1:nrow(aucrf_data), function(i){
 train <- aucrf_data[-i,]
  test <- aucrf_data[i,]
  set.seed(seed)
  temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
  LOO_probs[i] <- predict(temp_model$RFopt, test, type='prob')[,2]
})
LOO_roc <- roc(aucrf_data$Euth_Early~LOO_probs)

#### leave-one-cage-out
LOcO_df <- Predict_df %>% 
                select(Euth_Early, cage_id, one_of(euth_otu_feat))
cage_list <- as.character(unique(LOcO_df$cage_id))

LOcO_probs <- c()
LOcO_rf <- c()
for(i in 1:length(cage_list)){
  by_cage <- LOcO_df$cage_id == cage_list[i]
  LOcO_predict <- select(LOcO_df, -cage_id)
  train <- LOcO_predict[!by_cage,]
  test <- LOcO_predict[by_cage,]
  set.seed(seed)
  temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
  LOcO_rf[[i]] <- temp_model$RFopt
  LOcO_probs[LOcO_df$cage_id == cage_list[i]] <- predict(temp_model$RFopt, test, type='prob')[,2]
}
LOcO_roc <- roc(LOcO_df$Euth_Early~LOcO_probs)

#### leave-2-out
LOdO_df <- Predict_df %>% 
                select(Euth_Early, human_source, one_of(euth_otu_feat))
donor_list <- as.character(unique(LOdO_df$human_source))

LOdO_probs <- c()
LOdO_rf <- c()
for(i in seq(1,length(donor_list),2)){
  set.seed(seed)
  donors <- sample(donor_list)
  by_donor <- LOdO_df$human_source %in% donors[i:(i+1)]
  LOdO_predict <- select(LOdO_df, -human_source)
  train <- LOdO_predict[!by_donor,]
  test <- LOdO_predict[by_donor,]
  set.seed(seed)
  temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
  LOdO_rf[[i]] <- temp_model$RFopt
  LOdO_probs[LOdO_df$human_source %in% donors[i:(i+1)]] <- predict(temp_model$RFopt, test, type='prob')[,2]
}
LOdO_roc <- roc(LOdO_df$Euth_Early~LOdO_probs)

#### plot rocs
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='black', lwd=2, add=T, lty=1)
plot(cv10f_roc, col='gray', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=2)
plot(LOcO_roc, col='purple', lwd=2, add=T, lty=3)
plot(LOdO_roc, col='red', lwd=2, add=T, lty=3)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('Leave-1 cage-out (AUC: %.3g)',LOcO_roc$auc),
                               sprintf('Leave-2 cages-out (AUC: %.3g)',LOdO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc) #,
#                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
#                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(1,2,2,2,1,0,0), lwd=2, col=c('black','blue','purple','red','gray'), bty='n')
```

### Partial Dependency with Relative Abundance Overlayed Plots
###### Using optimal RF model generated from whole dataset
###### Plot labels contain MDA of each OTU

```{r rel abund partial depend, fig.width=10, fig.height=10}
layout(matrix(c(1:9), ncol=3, byrow=TRUE))

for (i in seq_along(euth_otu_feat)) {
  partial_otu <- partialPlot(rf_early_aucrf$RFopt, aucrf_data, euth_otu_feat[i], xlab='',
                main=paste(get_tax(row_list = euth_otu_feat)$tax_label[i]))
points(Predict_early_euth_df[Predict_early_euth_df$Euth_Early == 1 ,euth_otu_feat[i]], 
       jitter(rep(0.25*(diff(range(partial_otu$y))) + min(partial_otu$y), 
                  sum(Predict_early_euth_df$Euth_Early == 1)), 
              amount = (0.1*(diff(range(partial_otu$y))))), 
              pch = 21, bg = 'red')
points(Predict_early_euth_df[Predict_early_euth_df$Euth_Early == 0 ,euth_otu_feat[i]], 
       jitter(rep(0.75*(diff(range(partial_otu$y))) + min(partial_otu$y), 
                  sum(Predict_early_euth_df$Euth_Early == 0)), 
              amount = (0.1*(diff(range(partial_otu$y))))), 
              pch = 21, bg = 'gray')
  axis(side = 4, labels = c('Severe', 'Mild'),
       at = c(0.25*(diff(range(partial_otu$y))) + min(partial_otu$y),
              0.75*(diff(range(partial_otu$y))) + min(partial_otu$y)))
}
plot(NULL, xlim = c(0,1), ylim = c(0,1), xlab = '', ylab = '', axes = F)
legend('topleft', c('Mild', 'Severe'), pch = 21, pt.bg = c('gray','red'), cex = 2)
```

Would adding initial increase/slope of CFU or day 1 CFU help predict?

* Adding CFU decreases AUC and improves Random model, likely due to samples lost due to lack of cfu data which are biased towards the severe samples
    - Day 1 CFU loses 2 samples  [AUCRF model with Day 1 CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day1.html)
    - Day 2 cfu loses 14 samples, 10 are severe  [AUCRF model with Day 2 CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day2.html)

Would it help to add community information such as diversity or enterotype?

* When adding enterotype, diversity, and CFU, invsimpson and cfu appear as features, but when cfu is removed, invsimpson is no longer a predictor

But how to show since model is already approaching 1?

* Rerun AUCRF with added features and see if they replace any OTUs

Leave 1 cage out RF with all otus gives and AUC of 0.70, but using the reduced features selected by AUCRF on the whole dataset gives an AUC of 0.80

To remove any cage effect RF could be picking out instead of the actual outcome, I first run RF on all OTUs to predict cage and separately human source. Then select the OTUS that are necessary for the lowest error rate (I went slightly above) from both cage and human souce RFs. Remove those OTUs from the OTU data frame and predict Euthanization with AUCRF

* Of note, I tried a few different seeds, and one otu (Turicibacter OTU 120) was identified as a classifier of cage/inocula but wasnt in others and was identified as a classifier of Euthanization. Also we occasionally lose OTU 50 as a classifier

####### Model run with both relative and absolute abundances and no significant difference in their output

***  
***