### 2016-12-05 submission/Manuscript.tex
got citations working, not done yet

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microbiota predict Clostridium difficile severity in germ-free mice colonized with human feces},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{\textbf{Microbiota predict \emph{Clostridium difficile} severity in
germ-free mice colonized with human feces}}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\begin{document}

\maketitle

\vspace{35mm}

Running title: Microbiota predict \emph{C. difficile} severity in
humanized mice

\vspace{35mm}

Kaitlin J. Flynn\textsuperscript{1+}, Nicholas
Lesniak\textsuperscript{1+}, Alyxandria M. Schubert\textsuperscript{2},
Hamide Sinani\textsuperscript{?}, and Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

\vspace{35mm}

\(\dagger\) Corresponding author:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

+ These authors contributed equally to this work

1. Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, Michigan 48109

2. Food and Drug Administration?

3. Department for Hamide?

\newpage
\linenumbers

\subsubsection{Abstract}\label{abstract}

\emph{Clostridium difficile} causes diarrheal disease when it
successfully colonizes a disrupted gut microbial community. Current
mouse models to study \emph{C. difficile} infection (CDI) rely on
pre-treatment with antibiotics to alter the mouse microbiome prior to
inoculation. This model does not allow for analysis of human-associated
microbial community members that modulate \emph{C. difficile}
colonization and expansion. To study human-associated microbes in the
context of CDI, we inoculated germ-free C57BL/6 mice with one of 16
human fecal samples from diarrheal or healthy donors and challenged with
\emph{C. difficile} 14 days later. Five unique donor-mice combinations
resulted in severe CDI while the remaining 11 only experienced mild
disease. Both healthy and diarrheal donors were susceptible to
colonization and severe symptoms of CDI. To determine if specific
microbes were associated with disease severity outcomes, we built a
classification Random Forest machine learning model based on relative
abundance data of the communities prior to infection. The model
identified a number of bacterial populations associated with the
development of severe CDI, including \emph{Bacilliales, Ruminococcaceae,
Ruminococcus, Staphylococcus, Streptococcus and Bacteriodetes}.
Additionally, a regression model accurately predicted colonization
levels of \emph{C. difficile} at one to ten days post-infection. This
model explained 99\% of the variance in the number of CFU isolated from
mouse stool. Members of \emph{Lachnospiraceae, Parabacteroides,
Bacteroidales, Bacteroidetes, Porphyromonadaceae} and unclassified
\emph{Bacteria} families were predictive of future \emph{C. difficile}
colonization levels. Finally, challenging these mice with different
strains of \emph{C. difficile} revealed that susceptible
human-associated microbial communities were prone to severe disease
independent of strain type. Taken together these results suggest that
human-associated microbial communities can be recapitulated in germ-free
mice and used to characterize dynamics of CDI. Because both healthy and
diarrheal patients were susceptible to severe CDI, machine-learning
models are useful to identify bacterial populations that allow
colonization and contribute to the development of \emph{C. difficile}
associated disease in humans.

\newpage

\subsubsection{Introduction}\label{introduction}

\emph{Clostridium difficile} is an opportunistic pathogen of the human
lower gastrointestinal tract. \emph{C. difficile} forms spores that can
persist on abiotic surfaces and are not readily killed by ethanol-based
hand-sanitizers, putting hospital patients particularly at risk. Indeed,
\textasciitilde{}12\% of hospital acquired infections in the United
States are due to \emph{C. difficile} and result in up to 15,000 deaths
annually {[}1{]}. Disruption of the native microbial community is the
most common risk factor for development of \emph{C. difficile} infection
(CDI) {[}2{]}. Antibiotic use and inflammatory bowel diseases are
associated with loss of colonization resistance to \emph{C. difficile}
through the loss of potentially protective bacterial families such as
\emph{Barnesiella} and \emph{Lachnospiraceae} in both mouse models and
human association experiments {[}3,4{]}. The composition of the
community is clearly important for the acquisition, resistance and
treatment of \emph{C. difficile} infection, as giving patients a healthy
fecal microbiome transplant is the most effective treatment for this
disease (cite Anna 2014). The precise mechanisms of colonization
resistance and restoration of a healthy community are yet to be
discovered.

Murine models to study CDI typically rely on treating
conventionally-raised mice with antibiotics either in drinking water or
by injection to induce susceptibility (3, 4). This model provides a
convenient way to study \emph{C. difficile} pathogenesis and virulence
factors. Numerous microbiome studies have been performed using this
model to determine the antibiotic classes (5), starting microbial
community (6) and metabolites (7) that impact development and severity
of \emph{C. difficile} infection. While informative, these studies are
somewhat removed from human disease because they only examine
mouse-associated microbial communities.

Gnotobiotic or germ-free mouse models have been used for a range of
studies of CDI, including assessment of species-specific interactions
between \emph{C. difficile} and competing microbial community members
(8), analysis of nutrient restriction (9), \emph{in vivo}
transcriptomics of \emph{C. difficile} and examination of host immune
response to CDI (10). Further, CDI therapeutics such as antibiotics and
fecal microbiota transplants have been tested extensively in a
gnotobiotic-piglet or piglet-to-gnotobiotic-mouse model of disease (11),
(12). Pigs have a longer digestive tract with components more similar to
humans than mice and are typically infected by strains typical in human
infection (13). However, the murine and porcine microbiomes typically do
not resemble those of the human gut.

The power of the gnotobiotic models to study CDI has been further
realized by first inoculating germ-free mice and piglets with human
stool microbes. In one study, germ-free piglets were acutely colonized
with human feces for one week and then treated with tigecycline. After
challenge with \emph{C. difficile} none of the antibiotic-treated
piglets succumbed to infection, while some of the untreated
human-colonized pigs did (11). In another study, germ-free mice
colonized with human feces were bred over several generations to create
a cohort of mice with identical human-derived microbiomes (14). These
mice were subsequently treated with a five-antibiotic cocktail to induce
dysbiosis and then were successfully colonized by \emph{C. difficile}
(14). While informative, these studies were limited in their use of only
one human donor as input inoculum.

Here we designed a laboratory system that allows for studying and
modeling microbial coumminity interactions with \emph{C. difficile}
infection in mice with human derived microbiomes. To test the impact of
individual human microbiomes on CDI, we colonized germ-free mice with 16
different human stool donors. We then characterized human-associated
microbiome response to \emph{C. difficile} challenge. Further, we built
a predictive model based on bacterial composition that classified
``at-risk'' microbiomes prior to infection with \emph{C. difficile}.
These findings show that human-associated microbiomes can be at risk for
CDI even in the absence of antibiotic-induced dysbiosis. Finally, our
results indicate that study of mice colonized with human feces can
replicate a range of clinical outcomes.

\subsubsection{Results}\label{results}

\textbf{Germ-free mice inoculated with human feces as model for C.
difficile infection.} To generate mice with human-derived microbiomes,
we inoculated one cage each of gnotobiotic C57/BL6 mice with one of 16
different human fecal donors. Five donors were patients that had
diarrhea that was not attributable to \emph{C. difficile} infection
while 11 donors were healthy at time of donation. Stool from a patient
that was colonized with virulent \emph{C. difficile} was used as a
positive control. After inoculation with human stool, mice microbiomes
were allowed to equilibrate for 14 days. Prior to infection, stool
samples were taken from each mouse to establish baseline microbiome
composition. Then, 100 spores of \emph{C. difficile} strain 430
(isolated from the positive control patient) were used to orally infect
each mouse. Mice were monitored for weight loss and clinical signs of
disease. Fecal samples were taken to enumerate \emph{C. difficile} CFU
and for microbiome analysis each day up to 10 days post-infection (Fig
1A). To ensure that the donors we selected represented a diverse array
of human microbiomes, we sequenced the 16S rRNA genes from donor fecal
inocula. Ordination of the distances between donor communities showed
that the donors each had distinctly different communities, independent
of whether the sample came from a sick or healthy person (Fig 1B).
Likewise, the microbial communities of the mice on day 0 were
characterized by sequencing of fecal pellet DNA prior to infection
(Table S1). Ordination of day 0 communities shows that mice were similar
to each other within each cage and donor, but distinct from other donors
(Fig 1C). This result confirmed that human-associated microbes were able
to colonize gnotobiotic mice and provide various initial communities to
test \emph{C. difficile} dynamics.

\textbf{C. difficile infection in mice with human-derived microbiota
cause a range of outcomes.} \emph{C. difficile} colonization was
monitored by daily plating of stool pellets for \emph{C. difficile} CFU.
Nearly all of the mice were colonized to 10\textsuperscript{5} --
10\textsuperscript{7} CFU by one day post-infection and remained
colonized at that level until the end of the experiment (Fig 2A). As one
indicator of disease, mouse weights were taken each day post-infection
and weight-loss was monitored alongside clinical signs. When mice were
judged to be too ill to continue they were humanely euthanized. Overall,
disease phenotypes fell into two classes. Mice that became severely ill
and lost 20\% or more of their starting body weight within one to two
days post-infection were classified as ``severe'' whereas mice that were
colonized with \emph{C. difficile} but did not show signs of disease or
severe weight loss were considered to have ``mild'' disease (Fig 2A,
2B). Interestingly, \emph{C. difficile} was able to cause severe disease
in both mice that had been colonized with healthy stool and those
colonized with diarrheal stool, suggesting susceptibility to CDI is
dependent on the composition of the starting microbiome and not
associated with donor clinical metadata.

\textbf{Nick will be updating the next two sections} \textbf{Microbes
present in the gut prior to infection are predictive of \emph{C.
difficile} colonization levels} Previous work in our group has
demonstrated that the microbes present in the mouse microbiome prior to
\emph{C. difficile} challenge can predict future colonization levels
(cite Alyx). To determine if the human-associated microbiome in the
mouse could similarly predict \emph{C. difficile} colonization we
employed a Random Forest machine learning algorithm similar to the one
used in previous studies. Relative abundance of OTUs with greater than
1\% abundance at day 0 were used as input to the model. We found that
the model explained X\% of the variance in colonization levels of
\emph{C. difficile} (Fig 3A). By refining the model to include only the
top X predictive OTUs, the model explained X\% of the variance. Partial
dependency plots of the top X OTUs predictive of \emph{C. difficile}
colonization reveal that colonization is positively associated with
\emph{Bacteriodetes}, \emph{Parabacteriodes} and
\emph{Porphyromonadaceae} (really?) (Fig 3B). Colonization was
negatively associated with the presence of \emph{Lachnospiraceae} (Fig
3B) though none of the human-associated mouse communities were
completely resistant to colonization by \emph{C. difficile} (Fig 2A,
3B). These results confirm that as in mouse microbiome studies, the
human-associated gut community prior to challenge is predictive of
\emph{C. difficile} colonization levels.

\textbf{The microbiome predicts \emph{C. difficile} infection severity.}
As \emph{C. difficile} colonization levels can be predicted by the
microbiome composition on day 0, we postulated that severity of disease
could also be predicted from day 0. Using weight loss as a proxy for
severity, we built a Random Forest classification model based on
relative abundance to determine which community members were associated
with either mild or severe disease. We optimized the model using the
AUC-RF algorithm to generate a Reciever Operating Curve (ROC) with
maximal Area Under the Curve (AUC) (cite Calle 2011), as used previously
in our group to classify colon cancer outcomes (cite Niel 2015 gen med).
The optimized model used X number of OTUs and identified several that
predicted increased weightloss and subsequent disease severity (Fig 4A,
4B). The populations that were most commonly found in mice that
succumbed to \emph{C. difficile} infection were members of the
\emph{Clostridiales}, \emph{Bacilliales}, \emph{Streptococcus}
(Lactobacilles?) and \emph{Bacteriodes}. (need more OTUs here, or number
that were used for model and then group by rank). The AUC was calculated
using the Out-Of-Bag (OOB) error for each sample (Fig 4A).
Cross-validation using 10-fold, leave-one-mouse-out and
leave-one-cage-out resulted in AUCs not significantly different from the
OOB (Fig 4A). This model reveals that the levels of X OTUs on day 0 can
predict disease outcome in mice with human-associated microbiomes.
(probably also going to want a supplemental table for predictive OTUs
for this and Fig 3)

\textbf{Susceptibility to infection is community-dependent.} \emph{C.
difficile} strains are known to vary in disease phenotypes based both on
pathogenicity factors of the strain and host microbiome composition
(cite someone). To test the susceptibility of human-derived microbiomes
to \emph{C. difficile} infection, two donor-mouse microbiomes that were
previously shown to be resistant to \emph{C. difficile} strain 430
induced weight loss (Fig 2B) (DA00396, DA00430-wait isnt this a control)
and one that was susceptible to severe disease (DA00578) were
independently challenged with one of three unique \emph{C. difficile}
strains (\emph{table or reference for toxin prod}, etc). As done
previously, \emph{C. difficile} stool CFU was enumerated daily and mouse
weight-loss tracked. Infection of resistant mice with virulent strain
458 or the weaker strain 299 resulted in colonization to levels near
that of 431 (Fig 5A). Accordingly, these mice did not experience severe
weight loss or show clinical signs of disease (Fig 5B). In contrast,
infection of susceptible mice (DA00578, Fig 1B) with pathogenic strain
395 resulted in colonization and severe weight loss by two days
post-infection (Fig 5A, 5B). Strain 395 is similarly virulent as strain
431, and shows a similar infection phenotype in these mice. These
preliminary results suggest that human-associated microbiomes may be
more predictive of disease severity than the \emph{C. difficile} strain.

\subsubsection{Discussion}\label{discussion}

Study of human-associated microbiomes and their interactions with
\emph{C. difficile} has been limited by the lack of an appropriate model
and available samples. Here we colonized 16 groups of mice with human
donor stool from patients with different clinical profiles or healthy
volunteers. This created a diverse set of human-associated mouse
microbiomes that allowed for testing \emph{C. difficile} infection
dynamics and community interactions. Our results showed that both
healthy and diarrheal-associated human communities were susceptible to
severe disease. Additionally, our results confirmed previous findings
that resistance to \emph{C. difficile} disease is dependent on the
entire microbial community and not just one or two taxa (cite alyx).
Further, we were able to learn classification models that identified
specific OTUs within susceptible communities that predispose mice to
infection. By subsequently infecting susceptible or resistant
communities with varying strains of \emph{C. difficile} we confirmed
that the initial state of the human-associated microbiome is more
predictive of disease severity than the strain used.

Here we colonized germ-free mice with a variety of human stool
communities and allowed these communities to colonize the mice for 2
weeks, prior to \emph{C. difficile} challenge. It is important to note
that the mouse gut environment is different from that of the human, and
as such exerts a selective pressure on the human stool microbial
community that stabilizes over the 2 weeks. Thus these mice cannot be
considered ``humanized'' but rather colonized by human-associated
microbes. Interestingly, clinical status was not associated with
proclivity to severe \emph{C. difficile} infection as some healthy donor
mice were susceptible. A surprising finding was that none of the
human-associated mouse communities were completely resistant to
colonization by \emph{C. difficile}. This is in contrast to numerous
reports of mouse microbiome studies (cite) and the knowledge that in
general healthy humans without gut perturbations are resistant to
\emph{C. difficile} infection (cite). On the other hand, it is known
that humans, especially health care employees, can be asymptomatic
carriers of \emph{C. difficile} (Furyura-Kanamari). This may explain the
fact that the majority of our groups of mice were colonized by \emph{C.
difficile} but did not display outward signs of illness or
infection-associated weight loss (Fig 2).

• Discuss prediction methods and outcomes - Nick - we have bugs that
predict severity, but do we have bugs that predict mild/asymptomatic?
probably, could discuss here • Discuss potential mechanisms for
interesting OTUs - Nick

Finally, our pilot experiments testing the efficiency of infection with
different \emph{C. difficile} strains have allowed for testing community
versus pathogen factors of infection. Our results are consistent with
previous reports that the composition of the gut microbiome determines
if an individual is successfully infected or not (cite Alyx and others).
\emph{C. difficile} strains are known to vary in pathogenicity,
primarily via the amount of toxin that is produced during infection
(cite). To confirm that disease risk is dependent on the microbiome, we
challenged susceptibile mice with an additional pathogenic strain of
\emph{C. difficile} and found that the mice indeed fell ill. In
contrast, we exposed an asymptomatic community to a relatively weak
strain of \emph{C. difficile} and found that the community could be
colonized by this one as well-- but not cause disease. This suggests
that this particular colony is prone to colonization by any \emph{C.
difficile} strain. Further experiments will test if infecting
susceptible communities with weak strains allows the mice to survive. It
has been recently reported that carriage of a non-toxigenic strain of
\emph{C. difficile} can prevent additional colonization by a pathogenic
strain, suggesting a role of niche-filling or immunity this
pre-colonization provides (cite cite). One interesting experiment would
be to recapitulate this finding in human-associated mouse microbiomes by
making a susceptible community resistant to infection. Further work with
human-associated communities in mice and mathematical modeling will
allow for uncovering of complex human-associated disease processes and
dynamics as they relate to human \emph{C. difficile} infection.

\subsubsection{Materials and Methods}\label{materials-and-methods}

\textbf{Mice} The experiments in this study used 6-8 week-old male and
female gnotobiotic C57/BL6 mice. The mice were born and maintainted in a
completely sterile environment in the Germ Free Mouse Facility at the
University of Michigan. The University Committee on Use and Care of
Animals approved all protocols and experiments conducted in this study.

\textbf{Human samples} Donor stool samples were chosen from a sample
bank collected from patients from October 2010 to November 2012 at the
University of Michigan. Sample collection methods and study was approved
by the Institutional Review Board at the University of Michigan. Stool
samples from University of Michigan Health System patients were
submitted to the microbiology lab for \emph{C. difficile} testing.
\emph{C. difficile} presence was measured using WHATEVER- Cdiff quik
check? toxin AB?. Samples from these patients were stored in Cary-Blair
medium at -80C. All patients consented to participation in the study.
Donor stool for this study was selected from one patient that tested
positive for \emph{C. difficile} infection and five patients that
presented with diarrhea and tested negative for \emph{C. difficile}.
Healthy donor stool was collected from adult volunteers living in the
Ann Arbor, MI area. 100ul of stool slurry was gavaged into each mouse.

\textbf{\emph{Clostridium difficile} and sample processing} For all
experiments, 100 spores of \emph{Clostridium difficile} were used to
orally infect mice as described previously (cite alyx). The \emph{C.
difficile} strains used were acquired from spore stocks of (WHERE) and
comprised of strains 431, 395, 299 and 468 (cite?). After gavage, the
remaining inoculum was diluted and plated to confirm infection dose.
Bacteria were cultured anaerobically on TCCFA plates and incubated at
37C (cite plate recipe). During infection, two fresh stool pellets were
collected from each mouse daily. One pellet was resuspended in anaerobic
phosphate-buffered saline and diluted and plated on TCCFA plates to
enumerate \emph{C. difficile} CFU. The other pellet was frozen at -80C
until the end of the experiment.

\textbf{DNA isolation and sequencing} Bacterial DNA was isolated from
banked donor stool and each mouse stool pellet using the MO-BIO
PowerSoil DNA isolation kit. This DNA was used as template for
amplification of the V4 region of the 16S rRNA gene and fragments were
sequenced on an Illumina MiSeq as previously described (cite Jim paper).

• Sequence curation/mothur, data analysis, code + availability of
data(SRA) and code (Github) - Nick

\subsection{Acknowledgments}\label{acknowledgments}

We would like to thank members of the Schloss lab and Jhansi Leslie for
analysis advice. We would also like to thank the Microbial Systems
Laboratory at the University of Michigan for sequencing assistance.

\newpage

\subsubsection{Figure Legends}\label{figure-legends}

\textbf{Figure 1. Germ-free mice inoculated with human feces as a model
for \emph{C. difficile infection.}} A) Experimental design. Stool from
16 healthy, diarrheal and CDI patients were independently inoculated
into 3-4 germ-free mice by oral gavage. 14 days later mice were orally
infected with 100 spores of C. difficile strain 431. Weight and stool
CFU were monitored for up to 10 days post infection. B) NMDS ordination
of donor stool communities prior to inoculating mice. Each point
represents one donor and are colored by clinical diagnosis. C) NDMS
ordination of the stool communities on day 0. Each symbol represents one
mouse and is colored by donor. Circles represent mice that experienced
mild disease and triangles represent those that suffered severe disease.

\textbf{Figure 2. \emph{C. difficile} infection results in mild or
severe disease.} A) \emph{C. difficile CFU} was enumerated by plating of
mouse stool pellets daily. Each point represents a mouse and the lines
represent the median CFU in each group. Error bars are interquartile
ranges. Red lines and points correspond to mice that succumbed to severe
disease whereas black lines and points correspond to mice that had mild
or no disease. B) Mouse weights were recorded and daily percent weight
loss calculated for each mouse. Data is presented as the median of each
group and interquartile ranges. Mice that succumbed to severe infection
typically lost a significant amount of weight by day 1 or 2 post
infection. Red lines correspond to severely ill mice, black to mice with
mild disease.

\textbf{Figure 3. Random Forest prediction of \emph{C. difficile}
colonization level.} A) OTUs above 1\% relative abundance on day 0 were
used to predict median log\textsubscript{10} CFU of \emph{C. difficile}
after colonization. OTUs were chosen such that they were not predictive
of cage or donor. Each point is a mouse colored by cage. B) Partial
dependency plots of the top six predictive OTUs. Line displats the
partial dependence of log\textsubscript{10} CFU on the relative
abundance of each predictive OUT. Each median log\textsubscript{10} CFU
is plotted against its relative abundance for each predictive OTU.

\textbf{Figure 4. Random Forest prediction of CDI severity.} OTUs above
1\% relative abundance on day 0 were used to predict disease severity.
OTUs were chosen such that they were not predictive of cage or donor.
Predictive classification tested via 10-fold (gray), leave-one-cage-out
(purple dashed) or leave-one-mouse-out (blue dashed) models are
displayed in (A). B) Partial dependency plots of most predictive OTUs.
Line displays the partial dependence of log\textsubscript{10} CFU on OTU
relative abundance. Points are the OTU relative abundance of each mouse
colored by outcome (red, severe, black, mild).

\textbf{Figure 5. Infection of mice with different \emph{C. difficile
strains}.} 3 strains of \emph{C. difficile} were used to infect mice
colonized with susceptible (DA00578) or resistant (DA00369, DA00430)
human donor stool. A) \emph{C. difficile} stool CFU was enumerated over
10 days. B) Percent weight loss was calculated each day for each mouse.
In both plots, each mouse is a point and lines represent the mean of
each cage.

\paragraph{Supplement}\label{supplement}

\textbf{Table S1: Mouse day 0 communities by donor genera (avg + stdev
of cage)}

\newpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

1. Lessa FC, Mu Y, Bamberg WM, Beldavs ZG, Dumyati GK, Dunn JR, et al.
Burden ofClostridium difficileInfection in the united states. New
England Journal of Medicine. New England Journal of Medicine (NEJM/MMS);
2015;372: 825--834.
doi:\href{http://dx.doi.org/10.1056/nejmoa1408913}{10.1056/nejmoa1408913}

2. Britton RA, Young VB. Interaction between the intestinal microbiota
and host in clostridium difficile colonization resistance. Trends in
Microbiology. Elsevier BV; 2012;20: 313--319.
doi:\href{http://dx.doi.org/10.1016/j.tim.2012.04.001}{10.1016/j.tim.2012.04.001}

3. Schubert AM, Rogers MAM, Ring C, Mogle J, Petrosino JP, Young VB, et
al. Microbiome data distinguish patients with clostridium difficile
infection and non-c. difficile-associated diarrhea from healthy
controls. mBio. American Society for Microbiology; 2014;5:
e01021--14--e01021--14.
doi:\href{http://dx.doi.org/10.1128/mbio.01021-14}{10.1128/mbio.01021-14}

4. Vincent C, Stephens DA, Loo VG, Edens TJ, Behr MA, Dewar K, et al.
Reductions in intestinal clostridiales precede the development of
nosocomial clostridium difficile infection. Microbiome. Springer Nature;
2013;1: 18.
doi:\href{http://dx.doi.org/10.1186/2049-2618-1-18}{10.1186/2049-2618-1-18}

\end{document}