### 2020-03-03 exploratory/scratch/previous_analysis/submission/Schubert_humanCdGF_2015.Rmd
move old analysis files from 2016 to exploratory/scratch

---
title: 'Human-derived microbiota challenged with C.diff'
author: 'Alyxandria Schubert'
date: '10 June 2015'
---

# Colonization and Infection Resistance by Human-dervied Microbiotas against _Clostridium difficile_

**Running title:** Colonization and Infection Resistance against _Clostridium difficile_

Alyxandria M. Schubert, Jhansi Leslie, and Patrick D. Schloss\*

Correspondence:

> Department of Microbiology and Immunology
> University of Michigan
> 1520A Medical Science Research Bldg. I
> 1500 W. Medical Center Dr.
> Ann Arbor, MI 48109
> pschloss@umich.edu
> 734.647.5801

```{r, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
opts_chunk$set("dev" = c("png", "pdf"))
opts_chunk$set(results = "hold")
opts_chunk$set(fig.show = "hold")
opts_chunk$set(fig.align = "center")
opts_chunk$set(cache = FALSE)
opts_chunk$set(echo = FALSE)
opts_chunk$set(warning = FALSE)
```

### Abstract

### (Importance)

### Introduction

### Results

```{r summary_stats}

metadata <- read.table(file="~/Documents/Schloss/Data/GF/GFmetadata.txt", header=T, na.strings="NA", stringsAsFactors = FALSE)
num_sources <- length(unique(metadata$human_source)) 
## add source group column

num_cages <- length(unique(metadata$cage_id))
num_cages <- length(unique(metadata$cage_id))
num_mice <- length(unique(metadata$mouse_id))
plot( metadata$day, metadata$cdiff_cfu, log="y", ylim=c(.01, 10e11))
```

Stool from `r num_sources` human subjects were selected for gavage into germ-free mice. Of these subjects, 
X were inpatient and X were healthy subjects.

### Discussion

### Materials and Methods

### Acknowledgements

### Figure Legends