### 2021-05-26 exploratory/notebook/ml_analysis/ml_analysis_notebook.Rmd
start notebook analyzing ml performance

---
title: "ml_output_notebook"
author: "Nick Lesniak"
date: "5/21/2021"
output: rmarkdown::github_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_knit$set(root.dir = '../../..')
```

```{r read in data}
library(tidyverse)

# load in data
performance_df <- read_tsv('../../../data/process/ml/combined_ml_performance.tsv',
                     col_type = 'dddddddddddddddcdc') 
features_df <- read_tsv('../../../data/process/ml/combined_ml_feature_imp.tsv',
                     col_type = 'cccccDddddDcdddddcdl') 
hyperparameter_df <- read_tsv('../../../data/process/ml/combined_ml_hp_performance.tsv',
                     col_type = 'ddcdcc')

training_frac_df <- map_dfr(c(50, 60, 70, 80, 90), function(x){
  read_tsv(
      paste0('../../../data/process/ml/combined_ml_performance_', x, '.tsv'),
      col_type = 'dddddddddddddddcdc') %>% 
    mutate(training_frac = x/100)
})
```

## Model Performance

```{r}
performance_df %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  ggplot(aes(x = method, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_wrap(dataset~.) + 
    theme_bw() + 
    coord_flip()

```

```{r}
training_frac_df %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, training_frac) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  ggplot(aes(x = method, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~training_frac) + 
    theme_bw() + 
    coord_flip()

```

## Hyperparameter Performance

```{r}
hyperparameter_df %>% 
  #filter(model == 'glmnet', dataset == 'same_day_toxin') %>% 
  #filter(model == 'glmnet', dataset == 'day_0_predict_future_toxin') %>% 
  #filter(model == 'glmnet', dataset == 'day_0_moribund') %>% 
  #filter(model == 'glmnet', dataset == 'day_10_histology') %>% 
  filter(model == 'glmnet') %>% 
    ggplot(aes(x = value, y = AUC, color = params)) + 
    geom_point(alpha = 0.25)+ 
    stat_summary(fun = median, geom = 'line') + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(model~dataset, scales = 'free_x') + 
    theme_bw() + 
    scale_x_log10()

hyperparameter_df %>% 
  #filter(model == 'rf', dataset == 'same_day_toxin') %>% 
  #filter(model == 'rf', dataset == 'day_0_predict_future_toxin') %>% 
  #filter(model == 'rf', dataset == 'day_0_moribund') %>% 
  #filter(model == 'rf', dataset == 'day_10_histology') %>% 
  filter(model == 'rf') %>% 
  ggplot(aes(x = value, y = AUC, color = params)) + 
    geom_point(alpha = 0.25)+ 
    stat_summary(fun = median, geom = 'line') + 
    ylim(c(0.5,1)) + 
    facet_grid(dataset~model, scales = 'free') + 
    theme_bw()

hyperparameter_df %>% 
  filter(model == 'glmnet') %>% 
  pull(value) %>% 
  unique()
```