### 2021-06-09 exploratory/notebook/ml_analysis/ml_analysis_notebook.Rmd
update with otu only models

---
title: "ml_output_notebook"
author: "Nick Lesniak"
date: "5/21/2021"
output: rmarkdown::github_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_knit$set(root.dir = '../../..')
```

```{r read in data}
library(tidyverse)

# load in hyperparameter/training data
training_frac_df <- map_dfr(c(50, 60, 70, 80, 90), function(x){
  read_tsv(
      paste0('../../../data/process/ml/training_frac/combined_ml_performance_', x, '.tsv'),
      col_type = 'dddddddddddddddcdc') %>% 
    mutate(training_frac = x/100)
})

training_frac_hp_df <- map_dfr(c(50, 60, 70, 80, 90), function(x){
  read_tsv(
      paste0('../../../data/process/ml/training_frac/combined_ml_hp_performance_', x, '.tsv'),
      col_type = 'ddcdcc') %>% 
    mutate(training_frac = x/100)
})
```

## Hyperparameter Performance

```{r}
training_frac_df %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, training_frac) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_10_histology'))) %>% 
  ggplot(aes(x = method, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~training_frac) + 
    theme_bw() + 
    coord_flip()

```

```{r}
training_frac_hp_df %>% 
  filter(model == 'glmnet') %>% 
  mutate(dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_10_histology'))) %>% 
    ggplot(aes(x = value, y = AUC, color = params)) + 
    geom_point(alpha = 0.25)+ 
    stat_summary(fun = median, geom = 'line') + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~training_frac, scales = 'free_x') + 
    theme_bw() + 
    scale_x_log10()

training_frac_hp_df %>% 
  filter(model == 'rf') %>% 
  mutate(dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_10_histology'))) %>% 
  ggplot(aes(x = value, y = AUC, color = params)) + 
    geom_point(alpha = 0.25)+ 
    stat_summary(fun = median, geom = 'line') + 
    ylim(c(0.5,1)) + 
    facet_grid(dataset~training_frac, scales = 'free') + 
    theme_bw()
```

## Model Performance

Features  
-  same day toxin presence - CFU, taxonomic abundances  
-  day 0 future toxin presence - taxonomic abundances  
-  day 0 future moribund - max cfu, median cfu, max toxin, median toxin, toxin presence, taxonomic abundances  
-  day 10 histology (high/low) - cfu, max toxin, median toxin, toxin level, toxin presence, taxonomic abundances

```{r}

performance_df <- read_tsv('../../../data/process/ml/all_taxa_levels_all_features_ml_performance.tsv',
                     col_type = 'dddddddddddddddcdcc') 
features_df <- read_tsv('../../../data/process/ml/all_taxa_levels_all_features_ml_feature_imp.tsv',
                     col_type = 'ddcccdcc') 
hyperparameter_df <- read_tsv('../../../data/process/ml/all_taxa_levels_all_features_ml_hp_performance.tsv',
                     col_type = 'ddcdccc') 
taxonomy_df <- read_tsv('../../../data/process/final.taxonomy.tidy.tsv',
                        col_type = cols(.default = col_character()))

performance_df %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_10_histology')),
         taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() 
```

*day 10 histology is only predicting the high/low histology scores

```{r}
hyperparameter_df %>% 
  filter(model == 'glmnet') %>% 
  mutate(dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_10_histology')),
         taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = value, y = AUC, color = taxonomic_level)) + 
    geom_point(alpha = 0.25) + 
    stat_summary(fun = median, geom = 'line') + 
    ylim(c(0.5,1)) + 
    facet_grid(dataset~taxonomic_level, scales = 'free') + 
    theme_bw() + 
    scale_x_log10()

```

```{r}
hyperparameter_df %>% 
  filter(model == 'rf') %>% 
  mutate(dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_10_histology')),
         taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = value, y = AUC, color = taxonomic_level)) + 
    geom_point(alpha = 0.25) + 
    stat_summary(fun = median, geom = 'line') + 
    ylim(c(0.5,1)) + 
    facet_grid(dataset~taxonomic_level, scales = 'free') + 
    theme_bw() + 
    scale_x_log10()

```

## Models and features

```{r}
performance_df %>% 
  filter(dataset == 'same_day_toxin') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() +
    geom_rect(data = data.frame(ymin = 0.7, ymax = 0.95, method = 'glmnet', taxonomic_level = 6, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
```

Using Logistic regression at the Phylum level the features with median differences greater than 0 are:

```{r}

features_df %>% 
  filter(dataset == 'same_day_toxin',
         method == 'glmnet',
         taxonomic_level == 'Family') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>%   
  ggplot(aes(x = reorder(names, median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()

```

```{r}
performance_df %>% 
  filter(dataset == 'day_0_predict_future_toxin') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() +
    geom_rect(data = data.frame(ymin = 0.35, ymax = 0.98, method = 'glmnet', taxonomic_level = 2, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA) +
    geom_rect(data = data.frame(ymin = 0.48, ymax = 0.99, method = 'rf', taxonomic_level = 6, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
```

Using Logistic regression at the Genus level or RF at the Phylum level the features with median differences greater than 0 are:

```{r}
features_df %>% 
  filter(dataset == 'day_0_predict_future_toxin',
         (method == 'rf' & taxonomic_level == 'Phylum') | (method == 'glmnet' & taxonomic_level == 'Genus')) %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset, scales= 'free_y', space = 'free_y') + 
    labs(x = NULL, y = 'Decrease AUC') +
    theme_bw()
```

```{r}
performance_df %>% 
  filter(dataset == 'day_0_moribund') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() +
    geom_rect(data = data.frame(ymin = 0.65, ymax = 1.01, method = 'glmnet', taxonomic_level = 5, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
```

Using Logistic regression at the Class level the features with median differences greater than 0 are:

```{r}
features_df %>% 
  filter(dataset == 'day_0_moribund',
         method == 'glmnet',
         taxonomic_level == 'Class') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()
```

## Using only OTUs

```{r}

performance_df <- read_tsv('../../../data/process/ml/otu_only_all_taxa_levels_ml_performance.tsv',
                     col_type = 'dddddddddddddddcdcc') 
features_df <- read_tsv('../../../data/process/ml/otu_only_all_taxa_levels_ml_feature_imp.tsv',
                     col_type = 'ddcccdcc') 
hyperparameter_df <- read_tsv('../../../data/process/ml/otu_only_all_taxa_levels_ml_hp_performance.tsv',
                     col_type = 'ddcdccc')

performance_df %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_10_histology')),
         taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() 
```

```{r}
performance_df %>% 
  filter(dataset == 'same_day_toxin') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw()  +
    geom_rect(data = data.frame(ymin = 0.71, ymax = 0.94, taxonomic_level = 3, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)

```

Using either Logistic regression or RF at the Family level the features with median differences greater than 0 are:

```{r}

features_df %>% 
  filter(dataset == 'same_day_toxin',
         method == 'glmnet',
         taxonomic_level == 'Family') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>%   
  ggplot(aes(x = reorder(names, median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()

features_df %>% 
  filter(dataset == 'same_day_toxin',
         method == 'rf',
         taxonomic_level == 'Family') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>%   
  ggplot(aes(x = reorder(names, median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()

```

```{r}
performance_df %>% 
  filter(dataset == 'day_0_predict_future_toxin') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw()  +
    geom_rect(data = data.frame(ymin = 0.38, ymax = 0.97, method = 'glmnet', taxonomic_level = 2, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA) +
    geom_rect(data = data.frame(ymin = 0.48, ymax = 0.99, method = 'rf', taxonomic_level = 6, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)

```

Using Logistic regression at the Genus level or RF at the Phylum level the features with median differences greater than 0 are:

```{r}
features_df %>% 
  filter(dataset == 'day_0_predict_future_toxin',
         (method == 'rf' & taxonomic_level == 'Phylum') | (method == 'glmnet' & taxonomic_level == 'Genus')) %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>% 
  ggplot(aes(x = reorder(names, median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset, scales= 'free_y', space = 'free_y') + 
    labs(x = NULL, y = 'Decrease AUC') +
    theme_bw()
```

```{r}
performance_df %>% 
  filter(dataset == 'day_0_moribund') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw()  +
    geom_rect(data = data.frame(ymin = 0.63, ymax = 1.01, taxonomic_level = 5, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
```

Using Logistic regression or RF at the Class level features with median differences greater than 0 are:

```{r}
features_df %>% 
  filter(dataset == 'day_0_moribund',
         method == 'glmnet',
         taxonomic_level == 'Class') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()

features_df %>% 
  filter(dataset == 'day_0_moribund',
         method == 'rf',
         taxonomic_level == 'Class') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()
```

## Use multiclass for endpoint histology scores

```{r}
hist_performance_df <- read_tsv('../../../data/process/ml/otu_only_all_taxa_levels_hist_ml_performance.tsv')  %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum')))
hist_features_df <- read_tsv('../../../data/process/ml/otu_only_all_taxa_levels_hist_ml_feature_imp.tsv') 
hist_hyperparameter_df <- read_tsv('../../../data/process/ml/otu_only_all_taxa_levels_hist_ml_hp_performance.tsv')

hist_performance_df %>% 
  select(method, seed, dataset, training =cv_metric_logLoss, test = logLoss, taxonomic_level, null_model) %>% 
  pivot_longer(cols = c(training, test), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(metric = ifelse(grepl('NULL', dataset), paste0(metric, '_NULL'), metric)) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    #geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(.~method, scales = 'free_y') + 
    theme_bw() + 
    coord_cartesian(ylim = c(0,2)) +
    geom_rect(data = data.frame(ymin = 0.1, ymax = 0.93, method = 'rf', taxonomic_level = 6, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
```

Using multiclass RF at the Phylum level the features with median differences greater than 0 are:

```{r}
hist_features_df %>% 
  filter(dataset == 'day_10_histology',
         method == 'rf',
         taxonomic_level == 'Phylum') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') + 
    theme_bw()

```