### 2021-10-21 exploratory/notebook/ml_analysis/ml_analysis_notebook.Rmd
add manuscript figures to rmd notebooks

---
title: "ml_output_notebook"
author: "Nick Lesniak"
date: "5/21/2021"
output: rmarkdown::github_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r read in data}
library(tidyverse)
library(here)

# load in hyperparameter/training data
training_frac_df <- read_tsv(
      here(paste0('data/process/ml/ml_tune_performance.tsv')),
      col_type = 'dddddddddddddddcdccd')

training_frac_hp_df <- read_tsv(
      here(paste0('data/process/ml/ml_tune_hp_performance.tsv')),
      col_type = 'ddcdcdcc')

```

## Hyperparameter Performance

```{r}
tax_levels <- c('Phylum', 'Class', 'Order', 'Family', 'Genus', 'OTU')
plot_model_performance <- function(plot_dataset){
  training_frac_df %>% 
    filter(dataset == plot_dataset,
           method == 'rf') %>% 
    select(method, seed, cv_auc = cv_metric_AUC, test_auc = AUC, training_fraction, taxonomic_level) %>% 
    pivot_longer(cols = c(cv_auc, test_auc), 
                 names_to = 'metric', 
                 values_to = 'Performance') %>% 
    mutate(taxonomic_level = factor(taxonomic_level, tax_levels)) %>% 
    ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
      geom_boxplot() + 
      geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
      facet_grid(method~training_fraction) + 
      theme_bw() + 
      coord_flip()
}

plot_lr_hp <- function(plot_dataset){
  training_frac_hp_df %>% 
    filter(model == 'glmnet',
           dataset == plot_dataset) %>% 
    mutate(taxonomic_level = factor(taxonomic_level, tax_levels)) %>% 
      ggplot(aes(x = value, y = AUC, color = params)) + 
      geom_point(alpha = 0.1)+ 
      stat_summary(fun = median, geom = 'line', color = 'black') + 
      #geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
      facet_grid(taxonomic_level~training_fraction, scales = 'free') + 
      theme_bw() + 
      scale_x_log10()
}
plot_rf_hp <- function(plot_dataset){
  training_frac_hp_df %>% 
    filter(model == 'rf',
           dataset == plot_dataset) %>% 
    mutate(taxonomic_level = factor(taxonomic_level, tax_levels)) %>% 
    ggplot(aes(x = value, y = AUC, color = params)) + 
      geom_point(alpha = 0.1)+ 
      stat_summary(fun = median, geom = 'line', color = 'black') + 
      xlim(c(0,10)) + 
      facet_grid(taxonomic_level~training_fraction, scales = 'free') + 
      theme_bw()
}
```

#### Same day toxin
```{r}
plot_model_performance('same_day_toxin')
plot_lr_hp('same_day_toxin')
plot_rf_hp('same_day_toxin')
```

#### Day 0 predict future toxin
```{r}
plot_model_performance('day_0_predict_future_toxin')
plot_lr_hp('day_0_predict_future_toxin')
plot_rf_hp('day_0_predict_future_toxin')
```

#### Day 0 predict moribund outcome
```{r}
plot_model_performance('day_0_moribund')
plot_lr_hp('day_0_moribund')
plot_rf_hp('day_0_moribund')
```

#### Day 0 predict high/low clinical score on day 10 
```{r}
plot_model_performance('day_0_histology')
plot_lr_hp('day_0_histology')
plot_rf_hp('day_0_histology')
```

## Model Performance

Features  
-  same day toxin presence - CFU, taxonomic abundances   
-  day 0 future toxin presence - taxonomic abundances

-  day 0 future moribund - median cfu, day w/toxin presence (prior to day 3), taxonomic abundances   
  -    to avoid duplicating similar data, i sumarized cfu/toxin to be a single feature to represent the amount or presence over the first two days of challenge since all mice were present for at least 2 days  
-  day 0 histology (high vs low) - median cfu, days w/toxin, taxonomic abundances

```{r}

performance_df <- read_tsv(here('data/process/ml/ml_performance.tsv'),
                     col_type = 'dddddddddddddddcdcc') %>% 
  filter(dataset != 'day_0_histology_50') %>% 
  mutate(dataset = gsub('_80', '', dataset),
         dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_0_histology')),
         taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum')))  
features_df <- read_tsv(here('data/process/ml/ml_feature_imp.tsv'),
                     col_type = 'ddcccddcc') %>% 
  filter(dataset != 'day_0_histology_50') %>% 
  mutate(dataset = gsub('_80', '', dataset))
hyperparameter_df <- read_tsv(here('data/process/ml/ml_hp_performance.tsv'),
                     col_type = 'ddcdccc') %>% 
  filter(dataset != 'day_0_histology_50') %>% 
  mutate(dataset = gsub('_80', '', dataset),
         dataset = factor(dataset, c('same_day_toxin', 'day_0_predict_future_toxin',
                                     'day_0_moribund', 'day_0_histology')),
         taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum')))   
taxonomy_df <- read_tsv(here('data/process/final.taxonomy.tidy.tsv'),
                        col_type = cols(.default = col_character()))

metadata <- read_tsv(here('data/process/metadata_tidy.tsv'),
					 col_type = 'cccccDddddDcdddddcdl') %>% 
	filter(cdiff_strain == 431) %>% 
	mutate(cdiff_cfu = ifelse(cdiff_cfu == 0, log10(60), log10(cdiff_cfu)))

toxin <- read_tsv(here('data/process/toxin_tidy.tsv'),
				  col_type = 'cccccdc') %>% 
	mutate(toxin = ifelse(Log_repiricoal_dilution > 1, 'present', 'absent'))

histology <- read_tsv(here('data/process/histology_tidy.tsv'),
					  col_type = 'ccccdddcdcc') %>% 
	mutate(hist_score = case_when(summary_score < 5 ~ 'low',
								  summary_score > 5 ~ 'high',
								  summary_score == 5 ~ 'mid',
								  T ~ 'NA'))

shared <- read_tsv(here('data/mothur/sample.final.0.03.subsample.shared'),
         col_types = cols(.default = col_double(),
                          Group = col_character())) %>% 
  mutate(Group = ifelse(grepl('_D|DA', Group), Group,
      gsub('D', '_D', Group))) %>% 
  select(-label, -numOtus)

same_day_toxin_df <- toxin %>% 
	select(group, toxin) %>% 
	inner_join(shared, by = c('group' = 'Group')) %>% 
	left_join(select(metadata, group, C_difficile_CFU = cdiff_cfu), by = c('group'))

#	from day 0?
day_0_predict_future_toxin_df <- toxin %>% 
	group_by(mouse_id) %>% 
	summarise(toxin = ifelse(max(Log_repiricoal_dilution) > 1, 'present', 'absent')) %>% 
	left_join(select(metadata, mouse_id, day, group), 
			  by = c('mouse_id')) %>% 
	filter(day == 0) %>% 
	inner_join(shared, by = c('group' = 'Group'))

# setup data to predict moribunidty (microbiome, cdiff, toxin)
toxin_summary <- toxin %>% 
	filter(toxin_sample_day  %in% c(1,2)) %>% 
	group_by(mouse_id) %>% 
	summarise(toxin_positive_days = sum(Log_repiricoal_dilution > 1))
cfu_summary <- metadata %>% 
	filter(day %in% c(1,2)) %>% 
	group_by(mouse_id, early_euth) %>% 
	summarise(median_CFU = median(cdiff_cfu))
day_0_moribund <- metadata %>% 
	select(group, mouse_id, day) %>% 
	filter(day == 0) %>% 
	left_join(cfu_summary, by = c('mouse_id')) %>% 
	left_join(toxin_summary, by = c('mouse_id')) %>% 
	inner_join(shared, by = c('group' = 'Group')) %>% 
	mutate(early_euth = ifelse(early_euth, 'severe', 'mild'))

# setup data to predict histology severity of colonized mice (microbiome, cdiff, toxin)
toxin_summary <- toxin %>% 
	group_by(mouse_id) %>% 
	summarise(toxin_positive_days = sum(Log_repiricoal_dilution > 1))
day_0_histology <- metadata %>% 
	select(group, mouse_id, day) %>% 
	filter(day == 0) %>% 
	left_join(cfu_summary, by = c('mouse_id')) %>% 
	left_join(toxin_summary, by = c('mouse_id')) %>% 
	inner_join(shared, by = c('group' = 'Group')) %>% 
	ungroup %>% 
	filter(early_euth == F) %>% 
	inner_join(select(histology, mouse_id, hist_score), by = c('mouse_id'))
```

```{r}
performance_df %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() 
```

```{r}
# did not run logistic regression in final run as using rf for final model
#hyperparameter_df %>% 
#  filter(model == 'glmnet') %>% 
#  ggplot(aes(x = value, y = AUC, color = taxonomic_level)) + 
#    geom_point(alpha = 0.25) + 
#    stat_summary(fun = median, geom = 'line') + 
#    ylim(c(0.5,1)) + 
#    facet_grid(dataset~taxonomic_level, scales = 'free') + 
#    theme_bw() + 
#    scale_x_log10() +
#    theme(legend.position = 'none')

```

```{r}
hyperparameter_df %>% 
  filter(model == 'rf') %>% 
  ggplot(aes(x = value, y = AUC, color = taxonomic_level)) + 
    geom_point(alpha = 0.25) + 
    stat_summary(fun = median, geom = 'line') + 
    ylim(c(0.5,1)) + 
    facet_grid(dataset~taxonomic_level, scales = 'free') + 
    theme_bw() + 
    scale_x_log10() + 
    theme(legend.position = 'none')

```

## Models and features

#### Modeling future production of toxin based on initial community

```{r}
day_0_toxin_perf_plot <- performance_df %>% 
  filter(dataset == 'day_0_predict_future_toxin') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() +
#    geom_rect(data = data.frame(ymin = 0.35, ymax = 0.98, method = 'glmnet', taxonomic_level = 2, Performance = 1),
#              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
#              fill = NA) +
    geom_rect(data = data.frame(ymin = 0.48, ymax = 0.99, method = 'rf', taxonomic_level = 6, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
day_0_toxin_perf_plot
```

Using Logistic regression at the Genus level or RF at the Phylum level the features with median differences greater than 0 are:

```{r}
day_0_toxin_production <- features_df %>% 
  filter(dataset == 'day_0_predict_future_toxin',
         (method == 'rf' & taxonomic_level == 'Phylum')) %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) 
day_0_toxin_feature_plot <- day_0_toxin_production %>%
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset, scales= 'free_y', space = 'free_y') + 
    labs(x = NULL, y = 'Decrease AUC') +
    theme_bw()
```

```{r}
day_0_toxin_features <- day_0_toxin_production %>% 
  select(names, median_diff) %>% 
  unique %>% 
  arrange(median_diff) %>% 
  pull(names)

day_0_toxin_feature_abundance_plot <- day_0_predict_future_toxin_df %>% 
  pivot_longer(cols = starts_with('Otu'), names_to = 'OTU', values_to = 'counts') %>% 
  left_join(taxonomy_df, by = "OTU") %>% 
  filter(Phylum %in% day_0_toxin_features) %>% 
  group_by(group, toxin, Phylum) %>% 
  summarise(counts = sum(counts)/21.07) %>% 
  mutate(taxon = factor(Phylum, day_0_toxin_features)) %>% 
  ggplot(aes(x = taxon, y = counts, color = toxin)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
                 position = position_dodge(width = 0.7)) +
    geom_jitter(position = position_jitterdodge(dodge.width = 0.7, jitter.width = 0.3), 
                alpha = 0.2) + 
    scale_y_log10() + 
    theme_bw() + 
    coord_flip() + 
    scale_color_manual(values = c('red', 'black'),
                       breaks = c('absent', 'present'), 
                       labels = c('False', 'True')) + 
    labs(x = NULL, y = 'Relative Abundance (%)', color  = 'Toxin Detected')

```

#### Modeling production of toxin with the community and CFU data from that same day

```{r}
same_day_toxin_perf_plot <- performance_df %>% 
  filter(dataset == 'same_day_toxin') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() +
    geom_rect(data = data.frame(ymin = 0.7, ymax = 0.95, method = 'rf', taxonomic_level = 6, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
same_day_toxin_perf_plot
```

Using Random Forest at the Phylum level the features with median differences greater than 0 are:

```{r}

same_day_toxin <- features_df %>% 
  filter(dataset == 'same_day_toxin',
         method == 'rf',
         taxonomic_level == 'Phylum') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) 
same_day_toxin_feature_plot <- same_day_toxin %>%   
  ggplot(aes(x = reorder(names, median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()
same_day_toxin_feature_plot
```

```{r}
same_day_toxin_features <- same_day_toxin %>% 
  select(names, median_diff) %>% 
  unique %>% 
  arrange(median_diff) %>% 
  pull(names)

same_day_toxin_df %>% 
  ggplot(aes(x = toxin, y = C_difficile_CFU, color = toxin)) + 
    geom_jitter(alpha = 0.2, width  = 0.3) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
                 position = position_dodge(width = 0.7)) +
    coord_flip() + 
    theme_bw() + 
    labs(x = NULL, y = 'C. difficile CFU', color  = 'Toxin Detected')
```

```{r}
same_day_toxin_feature_abundance_plot <- same_day_toxin_df %>% 
  pivot_longer(cols = starts_with('Otu'), names_to = 'OTU', values_to = 'counts') %>% 
  left_join(taxonomy_df, by = "OTU") %>% 
  filter(Phylum %in% same_day_toxin_features) %>% 
  group_by(group, toxin, Phylum) %>% 
  summarise(counts = sum(counts)/21.07) %>% 
  mutate(taxon = factor(Phylum, same_day_toxin_features)) %>% 
  ggplot(aes(x = taxon, y = counts, color = toxin)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
                 position = position_dodge(width = 0.7)) +
    geom_jitter(position = position_jitterdodge(dodge.width = 0.7, jitter.width = 0.3), 
                alpha = 0.05) + 
    scale_y_log10() + 
    theme_bw() + 
    coord_flip() + 
    scale_color_manual(values = c('red', 'black'),
                       breaks = c('absent', 'present'), 
                       labels = c('False', 'True')) + 
    labs(x = NULL, y = 'Relative Abundance (%)', color  = 'Toxin Detected')

```

#### Modeling severe disease (moribund) from the initial community

```{r}
day_0_moribund_perf_plot <- performance_df %>% 
  filter(dataset == 'day_0_moribund') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() +
    geom_rect(data = data.frame(ymin = 0.65, ymax = 1.01, method = 'rf', taxonomic_level = 5, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
day_0_moribund_perf_plot
```

Using random forest at the Class level the features with median differences greater than 0 are:

```{r}
day_0_moribund_plot_df <- features_df %>% 
  filter(dataset == 'day_0_moribund',
         method == 'rf',
         taxonomic_level == 'Class') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) 
day_0_moribund_feature_plot <- day_0_moribund_plot_df %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()
day_0_moribund_feature_plot
```

```{r}
day_0_moribund_features <- day_0_moribund_plot_df %>% 
  select(names, median_diff) %>% 
  unique %>% 
  arrange(median_diff) %>% 
  pull(names)

day_0_moribund_feature_abundance_plot <- day_0_moribund %>% 
  pivot_longer(cols = starts_with('Otu'), names_to = 'OTU', values_to = 'counts') %>% 
  left_join(taxonomy_df, by = "OTU") %>% 
  filter(Class %in% day_0_moribund_features) %>% 
  group_by(group, early_euth, Class) %>% 
  summarise(counts = sum(counts)/21.07) %>% 
  mutate(taxon = factor(Class, day_0_moribund_features)) %>% 
  ggplot(aes(x = taxon, y = counts, color = early_euth)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
                 position = position_dodge(width = 0.7)) +
    geom_jitter(position = position_jitterdodge(dodge.width = 0.7, jitter.width = 0.3), 
                alpha = 0.2) + 
    scale_y_log10() + 
    theme_bw() + 
    coord_flip() + 
#    scale_color_manual(values = c('red', 'black'),
 #                      breaks = c('absent', 'present'), 
  #                     labels = c('False', 'True')) + 
    labs(x = NULL, y = 'Relative Abundance (%)', color  = 'Disease Severity')
day_0_moribund_feature_abundance_plot
```

#### Modeling endpoint histology scores

```{r, include = F, eval = F}
hist_performance_df <- read_tsv('../../../data/process/ml/hist_ml_performance.tsv')
hist_features_df <- read_tsv('../../../data/process/ml/hist_ml_feature_imp.tsv') 
hist_hyperparameter_df <- read_tsv('../../../data/process/ml/hist_ml_hp_performance.tsv')

hist_performance_df %>% 
  select(method, seed, dataset, training =cv_metric_logLoss, test = logLoss, taxonomic_level, null_model) %>% 
  pivot_longer(cols = c(training, test), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(metric = ifelse(grepl('NULL', dataset), paste0(metric, '_NULL'), metric)) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    #geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    facet_grid(.~method, scales = 'free_y') + 
    theme_bw() + 
    coord_cartesian(ylim = c(0,2)) +
    labs(x = NULL, y = 'Performance - logLoss') + 
    geom_rect(data = data.frame(ymin = 0.1, ymax = 0.93, method = 'rf', taxonomic_level = 6, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
# helpful logloss explanation https://stats.stackexchange.com/questions/276067/whats-considered-a-good-log-loss
```

```{r, include = F, eval = F}
#Using multiclass RF at the Phylum level the features with median differences greater than 0 are:

hist_features_plot_df <- hist_features_df %>% 
  filter(dataset == 'day_10_histology',
         method == 'rf',
         taxonomic_level == 'Phylum') %>% 
  group_by(names) %>% 
  mutate(median_diff = median(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0)
hist_features_plot_df %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') + 
    theme_bw()

```

```{r, include = F, eval = F}
hist_features <- hist_features_plot_df %>% 
  select(names, median_diff) %>% 
  unique %>% 
  arrange(median_diff) %>% 
  pull(names)

day_10_histology_df %>% 
  pivot_longer(cols = starts_with('Otu'), names_to = 'OTU', values_to = 'counts') %>% 
  left_join(taxonomy_df, by = "OTU") %>% 
  filter(Phylum %in% hist_features) %>% 
  group_by(group, hist_score, Phylum) %>% 
  summarise(counts = sum(counts)/21.07) %>% 
  mutate(Phylum = factor(Phylum, hist_features),
         hist_score = factor(hist_score, c('low', 'mid', 'high'))) %>% 
  ggplot(aes(x = Phylum, y = counts, color = hist_score)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
                 position = position_dodge(width = 0.7)) +
    geom_jitter(position = position_jitterdodge(dodge.width = 0.9, jitter.width = 0.3), 
                alpha = 0.2) + 
    scale_y_log10() + 
    theme_bw() + 
    coord_flip() + 
#    scale_color_manual(values = c('red', 'black'),
 #                      breaks = c('absent', 'present'), 
  #                     labels = c('False', 'True')) + 
    labs(x = NULL, y = 'Relative Abundance (%)', color  = 'Histology Score')

```

```{r}
day_0_hist_perf_plot <- performance_df %>% 
  filter(dataset == 'day_0_histology') %>% 
  select(method, seed, dataset, cv_auc = cv_metric_AUC, test_auc = AUC, taxonomic_level) %>% 
  pivot_longer(cols = c(cv_auc, test_auc), 
               names_to = 'metric', 
               values_to = 'Performance') %>% 
  mutate(taxonomic_level = factor(taxonomic_level, 
                                  levels = c('OTU', 'Genus', 'Family', 'Order', 'Class', 'Phylum'))) %>% 
  ggplot(aes(x = taxonomic_level, y = Performance, color = metric)) + 
    geom_boxplot() + 
    geom_hline(yintercept = 0.5, linetype = 'dashed', size = 0.25) +
    #facet_grid(dataset~method, scales = 'free_y') + 
    theme_bw() +
    geom_rect(data = data.frame(ymin = 0.48, ymax = 1.01, method = 'rf', taxonomic_level = 2, Performance = 1),
              aes(xmin = taxonomic_level - 0.5, xmax = taxonomic_level + 0.5, ymin = ymin, ymax = ymax, color = NA), 
              fill = NA)
day_0_hist_perf_plot
```

Using Logistic regression at the Class level the features with median differences greater than 0 are:

```{r}
day_0_histology_plot_df <- features_df %>% 
  filter(dataset == 'day_0_histology',
         method == 'rf',
         taxonomic_level == 'Genus') %>% 
  group_by(names) %>% 
  mutate(median_diff = mean(perf_metric_diff)) %>% 
  ungroup() %>% 
  filter(median_diff > 0) 
day_0_hist_feature_plot <- day_0_histology_plot_df %>% 
  ggplot(aes(x = reorder(names,median_diff), y = perf_metric_diff)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    geom_jitter(alpha = 0.1, width = 0.1) + 
    coord_flip() + 
    facet_grid(method~dataset) + 
    labs(x = NULL, y = 'AUC Diff with feature permuted') +
    theme_bw()
day_0_hist_feature_plot
```

```{r}
day_0_histology_features <- day_0_histology_plot_df %>% 
  select(names, median_diff) %>% 
  separate(names, into = c('A', 'B')) %>% 
  unique %>% 
  arrange(median_diff) %>% 
  pivot_longer(cols = c('A', 'B'), values_to = 'names') %>% 
  pull(names) %>% 
  unique

day_0_hist_feature_abundance_plot <- day_0_histology %>% 
  filter(hist_score %in% c('high', 'low')) %>% 
  pivot_longer(cols = starts_with('Otu'), names_to = 'OTU', values_to = 'counts') %>% 
  left_join(taxonomy_df, by = "OTU") %>% 
  filter(Genus %in% day_0_histology_features) %>% 
  group_by(group, hist_score, Genus) %>% 
  summarise(counts = sum(counts)/21.07) %>% 
  mutate(taxon = factor(Genus, day_0_histology_features)) %>% 
  ggplot(aes(x = taxon, y = counts, color = hist_score)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
                 position = position_dodge(width = 0.7)) +
    geom_jitter(position = position_jitterdodge(dodge.width = 0.7, jitter.width = 0.3), 
                alpha = 0.2) + 
    scale_y_log10() + 
    theme_bw() + 
    coord_flip() + 
#    scale_color_manual(values = c('red', 'black'),
 #                      breaks = c('absent', 'present'), 
  #                     labels = c('False', 'True')) + 
    labs(x = NULL, y = 'Relative Abundance (%)', color  = 'Disease Severity')
day_0_hist_feature_abundance_plot  
```

```{r}
day_0_toxin_perf_plot
day_0_hist_perf_plot
day_0_moribund_perf_plot
same_day_toxin_perf_plot

ggsave(here('results/figures/Figure_S2.jpg'),
  cowplot::plot_grid(
      cowplot::plot_grid(day_0_toxin_perf_plot + coord_flip(), day_0_toxin_feature_plot, nrow = 1),
      #cowplot::plot_grid(same_day_toxin_perf_plot + coord_flip(), same_day_toxin_feature_plot, nrow = 1),
      cowplot::plot_grid(day_0_moribund_perf_plot + coord_flip(), day_0_moribund_feature_plot, nrow = 1),
      cowplot::plot_grid(day_0_hist_perf_plot + coord_flip(), day_0_hist_feature_plot, nrow = 1),
      ncol = 1),  height = 9, width = 6.875, unit = 'in')

ggsave(here('results/figures/Figure_5.jpg'),
  cowplot::plot_grid(
#    cowplot::plot_grid(
     cowplot::plot_grid(NULL,
      day_0_toxin_feature_abundance_plot + 
        theme(legend.position = 'right') + 
        labs(color = 'Toxin\nDetected'),
      nrow = 1, rel_widths = c(1, 25)),
#      same_day_toxin_feature_abundance_plot + 
#        theme(legend.position = 'none'),
#      ncol = 1),
#    cowplot::plot_grid(  
      day_0_moribund_feature_abundance_plot + 
        theme(legend.position = 'right') + 
        labs(color = 'Disease\nSeverity'),
     cowplot::plot_grid(NULL,     
      day_0_hist_feature_abundance_plot + 
        theme(legend.position = 'right') + 
        scale_color_manual(values = c('dark green', 'light green')) + 
        labs(color = 'Clinical\nScore'),
      nrow = 1, rel_widths = c(1, 20)),
      rel_heights = c(9, 15, 14),
      labels = c('A', 'B', 'C'),
      ncol = 1), 
#    nrow = 1),
  height = 9, width = 6, unit = 'in')

rf_performance <- performance_df %>% 
  filter((dataset == 'day_0_predict_future_toxin' & taxonomic_level == 'Phylum') | 
         (dataset == 'day_0_moribund' & taxonomic_level == 'Class') |
         (dataset == 'day_0_histology' & taxonomic_level == 'Genus')) %>% 
  group_by(dataset) %>% 
  summarise(AUC = round(mean(AUC), 2)) %>% 
  pivot_wider(names_from = dataset, values_from = AUC)

```

## What are the temporal dynamics of the taxa identified in the Lefse/RF models?

```{r}
taxonomy_df %>% 
  filter(Genus == 'Helicobacter') %>% 
  left_join(pivot_longer(shared, 
                         cols = -Group, names_to = 'OTU', values_to = 'counts')) %>% 
  left_join(metadata, by = c('Group' = 'group')) %>% 
  #filter(early_euth == F) %>% 
  group_by(mouse_id, day, Genus, early_euth) %>% 
  summarise(counts = sum(counts)) %>% 
  inner_join(select(histology, mouse_id, hist_score), by = 'mouse_id') %>% 
  mutate(outcome = ifelse(early_euth, 'severe', hist_score),
         outcome = factor(outcome, levels = c('low', 'high', 'severe'))) %>% 
  filter(outcome != 'mid') %>% 
  ggplot(aes(x = day, y = counts, group = interaction(day, outcome), color = outcome)) + 
    geom_boxplot() + 
    facet_wrap(Genus~.) + 
    theme_bw()

taxonomy_df %>% 
  filter(Family %in% c('Enterobacteriaceae', 'Pseudomonadaceae', 'Neisseriaceae', 'Sutterellaceae')) %>% 
  left_join(pivot_longer(shared, 
                         cols = -Group, names_to = 'OTU', values_to = 'counts')) %>% 
  left_join(metadata, by = c('Group' = 'group')) %>% 
  #filter(early_euth == F) %>% 
  group_by(mouse_id, day, Family, early_euth) %>% 
  summarise(counts = sum(counts)) %>% 
  inner_join(select(histology, mouse_id, hist_score), by = 'mouse_id') %>% 
  mutate(outcome = ifelse(early_euth, 'severe', hist_score),
         outcome = factor(outcome, levels = c('low', 'high', 'severe'))) %>% 
  filter(outcome != 'mid') %>% 
  group_by(Family) %>% 
  mutate(present = sum(counts > 0)) %>% 
  filter(present > 30) %>% 
  ggplot(aes(x = day, y = counts, group = interaction(day, outcome), color = outcome)) + 
    geom_boxplot() + 
    facet_wrap(Family~., scales = 'free_y') + 
    theme_bw()

plot_temporal_genera <- function(taxa_list){
  taxonomy_df %>% 
    filter(Genus %in% taxa_list) %>% 
    left_join(pivot_longer(shared, 
                           cols = -Group, names_to = 'OTU', values_to = 'counts')) %>% 
    left_join(metadata, by = c('Group' = 'group')) %>%
    group_by(mouse_id, day, Genus, early_euth) %>% 
    summarise(counts = sum(counts)) %>% 
    inner_join(select(histology, mouse_id, hist_score), by = 'mouse_id') %>% 
    mutate(outcome = ifelse(early_euth, 'severe', hist_score),
           outcome = factor(outcome, levels = c('low', 'high', 'severe'))) %>% 
    filter(outcome != 'mid') %>% 
    group_by(Genus) %>% 
    mutate(present = sum(counts > 0)) %>% 
    filter(present > 30) %>% 
    ggplot(aes(x = day, y = counts, group = interaction(day, outcome), color = outcome)) + 
      geom_boxplot() + 
      facet_wrap(Genus~., scales = 'free_y') + 
      theme_bw()
}

# manual selection/exploration
plot_temporal_genera(c('Limosilactobacillus', 'Ihubacter', 'Anaeroplasma', 
                      'Murimonas', 'Sellimonas', 'Lactonifactor', 
                      'Enterocloster', 'Akkermansia', 'Clostridium', 
                      'Anaerostipes', 'Monoglobus', 'Coprobacillus'))

plot_temporal_genera(c('Bacteroides', 'Prevotellaceae_unclassified', 'Parabacteroides',
                  'Muribaculaceae_unclassified', 'Duncaniella', 'Phocaeicola'))

plot_temporal_genera(c('Bifidobacterium', 'Eggerthella', 'Deinococcus',
                'Ligilactobacillus', 'Clostridium_IV', 'Turicibacter',
                'Clostridium_sensu_stricto'))
```

#### identify temporal patterns differentiating high/low clinical score

potential last figure or supplemental to show significant temporal trends

```{r}

#lefse_ temporal trend hi/low day 0-10
lefse_hilo <- c("Anaeromassilibacillus", "Anaeroplasma", "Bacteroidales_unclassified", "Bifidobacterium", "Blautia", "Clostridioides", "Clostridium_IV", "Clostridium_XlVa", "Duncaniella", "Escherichia/Shigella", "Faecalibacterium", "Fournierella", "Helicobacter", "Hungatella", "Ihubacter", "Intestinimonas", "Klebsiella", "Kosakonia", "Lactobacillales_unclassified", "Legionella", "Muribaculaceae_unclassified", "Neglecta", "Odoribacter", "Oscillibacter", "Paraclostridium", "Paramuribaculum", "Prevotellaceae_unclassified", "Ruthenibacterium")

temporal_lefse_plot <- taxonomy_df %>% 
  filter(Genus %in% lefse_hilo) %>% 
  left_join(pivot_longer(shared, 
                         cols = -Group, names_to = 'OTU', values_to = 'counts')) %>% 
  left_join(metadata, by = c('Group' = 'group')) %>%
  group_by(mouse_id, day, Genus, early_euth) %>% 
  summarise(counts = sum(counts)) %>% 
  inner_join(select(histology, mouse_id, hist_score), by = 'mouse_id') %>% 
  mutate(outcome = ifelse(early_euth, 'severe', hist_score),
         outcome = factor(outcome, levels = c('low', 'high', 'severe')),
         rel_abun = counts/2107 * 100) %>% 
  filter(outcome != 'mid') %>% 
  group_by(Genus) %>% 
  mutate(present = sum(counts > 0)) %>% 
  filter(present > 60) %>% 
  ggplot(aes(x = day, y = rel_abun, group = interaction(day, outcome), color = outcome)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
                 position = position_dodge(width = 0.7)) +
    geom_hline(yintercept = 1/21.07, linetype = 'dashed', size = 0.25) + 
    facet_wrap(Genus~., scales = 'free_y') + 
    scale_y_log10() + 
    theme_bw()

ggsave(here('results/figures/Figure_XX_temporal_lefse.jpg'),
       temporal_lefse_plot,
       height = 9, width = 6)
```

```{r}
# presentation plot
#
#strain_exp_donor_selection_plot <- taxonomy_df %>% 
#  filter(Genus %in% c('Anaerotignum',
#                      'Hungatella', 'Eisenbergiella', 'Clostridium_sensu_stricto', 
#                      'Enterocloster')) %>% 
#  left_join(pivot_longer(shared, 
#                         cols = -Group, names_to = 'OTU', values_to = 'counts')) %>% 
#  left_join(metadata, by = c('Group' = 'group')) %>%
#  filter(day == 0) %>% 
#  group_by(human_source, mouse_id, Genus, early_euth) %>% 
#  summarise(counts = sum(counts)) %>% 
#  inner_join(select(histology, mouse_id, hist_score), by = 'mouse_id') %>% 
#  mutate(Genus = gsub('_', ' ', Genus),
#  	     Genus = paste0('*', Genus, '*'),
#         outcome = case_when(early_euth == T ~ 'Severe', 
#                             hist_score == 'low' ~ 'Low Clinical Score',
#                             hist_score == 'high' ~ 'High Clinical Score'),
#         outcome = factor(outcome, levels = c('Low Clinical Score', 'High Clinical Score', 'Severe')),
#         relative_abundance = counts/2107 * 100,
#         relative_abundance = ifelse(relative_abundance == 0, 0.035, relative_abundance)) %>% 
#  filter(human_source %in% c('DA00369', 'DA00430', 'DA00578')) %>% 
#  mutate(donor_source = case_when(human_source == 'DA00369' ~ 'F',
#                                    human_source == 'DA00430' ~ 'G',
#                                    human_source == 'DA00578' ~ 'M'),
#         donor_source = factor(donor_source, levels = c('G' ,'F', 'M'))) %>% 
#  ggplot(aes(x = donor_source, y = relative_abundance, color = outcome)) + 
#    geom_point() + 
#    geom_hline(yintercept = 100/2107, linetype = 'dashed', size = 0.25) +
#    facet_grid(.~Genus) + 
#    scale_color_manual(breaks = c('Low Clinical Score', 'High Clinical Score', 'Severe'),
#                       values = c("#23980d", "#e26df8", "#154975")) + 
#    scale_y_log10(breaks = c(0.1, 1, 10), labels = c('0.1', '1', '10')) + 
#    theme_bw() +
#    labs(x = 'Donor Source', y = 'Relative Abundance', color = NULL) +
#    theme(legend.position = 'bottom',
#          strip.text = ggtext::element_markdown(),
#          strip.background = element_blank())
#
#
#
#ggsave('~/Desktop/donor_selection.jpg',
#       strain_exp_donor_selection_plot,
#       height = 4, width = 8)
```