### 2021-12-14 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\usepackage[labelfont=bf]{caption}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\floatplacement{figure}{H} % Keep the figure up top of the page
\DeclareUnicodeCharacter{0301}{*************************************}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\pagenumbering{arabic}
\linenumbers
\doublespacing

\hypertarget{the-gut-bacterial-community-potentiates-clostridioides-difficile-infection-severity.}{%
\section{\texorpdfstring{The gut bacterial community potentiates
\emph{Clostridioides difficile} infection
severity.}{The gut bacterial community potentiates Clostridioides difficile infection severity.}}\label{the-gut-bacterial-community-potentiates-clostridioides-difficile-infection-severity.}}

\vspace{30mm}

\textbf{Running title:} Microbiota potentiates \emph{Clostridioides
difficile} infection severity

\vspace{20mm}

Nicholas A. Lesniak\(^1\), Alyxandria M. Schubert\(^1\), Kaitlyn
Flynn\(^1\), Jhansi L Leslie\(^1\), Hamide Sinani\(^1\), Ingrid L.
Bergin\(^3\), Vincent B Young\(^2\), Patrick D. Schloss\(^{1,\dagger}\)

\vspace{30mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}\\
1. Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI\\
2. Division of Infectious Diseases, Department of Internal Medicine,
University of Michigan Medical School, Ann Arbor, MI\\
3. Unit for Laboratory Animal Medicine, University of Michigan, Ann
Arbor, MI

\newpage

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

The severity of \emph{Clostridioides difficile} infections (CDI) has
increased over the last few decades. Patient age, white blood cell
count, creatinine levels as well as \emph{C. difficile} ribotype and
toxin genes have been associated with disease severity. However, it is
unclear whether there is an associate between the gut microbiota disease
severity. The gut microbiota is known to interact with \emph{C.
difficile} during infection. Perturbations to the gut microbiota are
necessary for \emph{C. difficile} to colonize the gut. The gut
microbiota is thought to impair \emph{C. difficile} colonization through
bile acid metabolism, nutrient consumption and bacteriocin production.
It has not been shown whether the gut microbiota can affect CDI disease
severity. Here we demonstrate that gut bacterial communities can
contribute to disease severity. We derived diverse gut communities by
colonizing germ-free mice with different human fecal communities. The
mice were then infected with a single \emph{C. difficile} ribotype 027
clinical isolate which resulted in moribundity and histopathologic
differences. The variation in severity was associated with the human
fecal community that the mice received. Generally, facultative anaerobes
with pathogenic potential, such as \emph{Escherichia},
\emph{Helicobacter}, and \emph{Klebsiella}, were associated with more
severe outcomes. Bacterial groups associated with fiber degradation,
such as \emph{Coprobacillus}, were associated with less severe outcomes.
Finally, we showed that, when infected with a different \emph{C.
difficile} isolate, the outcome of the infection was similar between the
mice that received the same fecal donor community. These data indicate
that gut bacteria can influence CDI disease severity.

\hypertarget{importance}{%
\subsection{Importance}\label{importance}}

\emph{Clostridioides difficile} infections (CDI) can be asymptomatic or
range in severity from mild diarrhea to severe outcomes including
surgical interventions and death. Models that predict severity and guide
treatment decisions are based on clinical factors and \emph{C.
difficile} characteristics. Although the gut microbiome plays a role in
protecting against CDI, it has not been investigated for its effect on
CDI disease severity or incorporated into attempts to predict disease
severity. We demonstrated that variation in the microbiome of mice
colonized with human feces yielded a range of disease outcomes. These
results revealed groups of bacteria associated with both severe and mild
\emph{C. difficile} infection outcomes. Gut bacterial community data
from patients with CDI could improve our ability to identify patients at
risk of developing more severe disease and improve interventions which
treat the gut bacteria to reduce host damage.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\emph{Clostridioides difficile} infections (CDI) have increased in
incidence and severity since \emph{C. difficile} was first identified as
the cause of antibiotic-associated pseudomembranous colitis (1). CDI
disease severity can range from asymptomatic to mild diarrhea to toxic
megacolon and death. The Infectious Diseases Society of America (IDSA)
and Society for Healthcare Epidemiology of America (SHEA) guidelines
categorize severe disease based on white blood cell count greater than
15,000 cells/mm\(^3\) and/or a serum creatinine greater than 1.5 mg/dL,
and complicated disease based on systolic blood pressure below 90,
ileus, or toxic megacolon (2). Since these measures are only present
once the disease has advanced, the IDSA classification does not allow
for identification of patients at risk of severe CDI. Prediction models
have been developed to score a patient's risk for severe CDI outcomes
based on clinical factors but have not been robust for broad application
(3--6). Thus, we have limited ability to prevent patients from
developing severe CDI. Ideally, we could identify patients that are
high-risk of severe disease so their treatments could be adjusted to
halt the infection before the disease worsens.

Missing from CDI severity prediction models are the effects of the
indigenous gut bacteria. \emph{C. difficile} interacts with the gut
community in many ways. The indigenous bacteria of a healthy intestinal
community provide a protective barrier preventing \emph{C. difficile}
from infecting the gut. A range of mechanisms can disrupt this barrier,
including antibiotics, medications, or dietary changes, and lead to
increased susceptibility to CDI (7--9). Once in the gut, the indigenous
bacteria can either promote or inhibit \emph{C. difficile} through
producing molecules or modifying the environment (10, 11). Bile acids
metabolized by the gut bacteria can inhibit \emph{C. difficile} growth
and affect toxin production (12, 13). Bacteria in the gut also can
compete more directly with \emph{C. difficile} through antibiotic
production or nutrient consumption (14, 15). While the relationship
between the gut bacteria and \emph{C. difficile} has been established,
the effect the gut bacteria can have on CDI disease severity is unclear.

Studies investigating the severity of CDI disease involving the
microbiome have not shown clear evidence of a relationship between the
microbiome and disease severity. These studies were not setup to
identify a relationship between the microbiome and disease severity.
Studies that have used clinical data cannot control for host, microbiome
or \emph{C. difficile} ribotype variation (16). While disease severity
varies, it cannot be attributed to specifically the microbiome since all
factors vary. Murine experiments typically use a single mouse colony and
different \emph{C. difficile} ribotypes to create severity differences
(3). Recently, our group has begun uncovering the effect microbiome
variation has on \emph{C. difficile} infection. We showed the variation
in the bacterial communities between mice from different mouse colonies
resulted in different clearance rates of \emph{C. difficile} (17). We
also showed varied ability of mice to spontaneously eliminate \emph{C.
difficile} infection when they were treated with different antibiotics
prior to \emph{C. difficile} challenge (18). However, there have not
been any studies investigating whether variation in the microbiome
affects the resultant disease severity of \emph{C. difficile}
infections.

Here, we sought to determine whether the gut bacteria contribute to
disease severity. We hypothesized that since the gut bacteria affect the
metabolism of \emph{C. difficile} and its infection dynamics, it can
also affect the disease severity of the infection. To test this
hypothesis, we colonized germ-free C57BL/6 mice with human fecal
communities to create varied gut communities. We then challenged the
mice with a single \emph{C. difficile} isolate, ribotype 027 (RT027)
human isolate 431, and followed the mice for the development of severe
outcomes of moribundity and histopathologic cecal tissue damage. Since
the murine host and \emph{C. difficile} isolate were the same and only
the gut community varied, the variation in disease severity we observed
was attributable to the microbiome.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\textbf{\emph{C. difficile} is able to infect germ-free mice colonized
with human fecal microbial communities without perturbation.} To produce
gut microbiomes with greater variation than those found in conventional
mouse colonies, we colonized germ-free mice with bacteria from human
feces (19). We inoculated germ-free C57BL/6 mice with homogenized feces
from one of 15 human fecal samples via oral gavage. The human fecal
samples were selected because they represented diverse community
structures based on community clustering (20). The gut communities were
allowed to establish themselves for two weeks post inoculation (21). We
then surveyed the bacterial members of the gut communities by 16S rRNA
gene sequencing of fecal pellets (Figure 1A). The bacterial communities
from each mouse grouped more closely to those communities from mice that
received the same human fecal donor community than to the mice who
received a different human fecal donor community (Figure 1B). The
communities were primarily composed of populations of \emph{Clostridia},
\emph{Bacteroidia}, \emph{Erysipelotrichia}, \emph{Bacilli}, and
\emph{Gammaproteobacteria}. However, the gut bacterial communities of
each donor group of mice harbored unique relative abundance
distributions of the shared bacterial classes.

Next, we tested this set of mice with their human-derived gut microbial
communities for susceptibility to \emph{C. difficile} infection. A
typical mouse model of CDI requires pre-treatment with antibiotics, such
as clindamycin, to become susceptible to \emph{C. difficile}
colonization (22, 23). However, we wanted to avoid modifying the
communities with an antibiotic to maintain their unique microbial
compositions and ecological relationships. Since these communities came
from people who had increased risk of CDI, such as hospitalization or
recent antibiotic use (20), we decided to test if \emph{C. difficile}
was able to infect these mice without any antibiotic perturbation. We
hypothesized that \emph{C. difficile} would be able to colonize a subset
of mice whose communities were susceptible from perturbations either the
donor experienced or from the ecological selection when colonizing the
murine gut. Mice were challenged with 10\(^{3}\) \emph{C. difficile}
RT027 human isolate 431 spores. The mice were followed for 10 days
post-challenge and their stool was collected and plated for \emph{C.
difficile} colony forming units (CFU) to determine the extent of the
infection. Surprisingly, mice from all donors were colonized (Figure 2).
Only two mice were able to resist \emph{C. difficile} infection and both
received their community from the same donor, N3. By colonizing
germ-free mice with different human fecal communities, we were able to
generate diverse gut communities in mice, which were susceptible to
\emph{C. difficile} infection without further modification of the gut
community.

\textbf{Infection severity varies by initial community.} After we
challenged the mice with \emph{C. difficile}, we investigated the
outcome from the infection and its relationship to the initial
community. We followed the mice for 10 days post-challenge for
colonization density, toxin production, and mortality. Seven mice, from
Donors N1, N2 and N5, were not colonized on the day after \emph{C.
difficile} challenge but were infected (\textgreater10\(^{6}\)) by the
end of the experiment. All mice that received their community from Donor
M1 through M6 succumbed to the infection and became moribund within 3
days post-challenge. The remaining mice, except the uninfected Donor N3
mice, maintained \emph{C. difficile} infection through the end of the
experiment (Figure 2). At 10 days post-challenge, or earlier for the
moribund mice, mice were euthanised and fecal pellets were assayed for
toxin activity and cecal tissue was collected and scored for
histopathologic signs of disease (Figure 3). Overall, there was greater
toxin activity detected in the stool of the moribund mice (\emph{P} =
0.003). However, when looking at each group of mice, we observed a range
in toxin activity for both the moribund and non-moribund mice (Figure
3A). Non-moribund mice from Donors N4 through N9 had comparable toxin
activity as the moribund mice. Additionally, not all moribund mice had
toxin activity detected in their stool. This discordance between toxin
activity and disease severity has been previously described (24, 25).
Next we examined the cecal tissue for histopathologic damage. Moribund
mice had high levels of epithelial damage, tissue edema, and
inflammation (Figure S1) similar to previously reported histopathologic
findings for \emph{C. difficile} RT027 (26). As observed with toxin
activity, the moribund mice had higher histopathologic scores than the
non-moribund mice (\emph{P} \textless{} 0.001). However, unlike the
toxin activity, all moribund mice had consistently high histopathologic
summary scores (Figure 3B). The non-moribund mice, Donor groups N1
through N9, had a range in tissue damage from none detected to similar
levels as the moribund mice, which grouped by community donor. Together,
the toxin activity, histopathologic score, and moribundity showed
variation across the donor groups but were largely consistent within
each donor group.

\textbf{Microbial community members explain variation in CDI severity.}
We next interrogated the bacterial communities at the time of \emph{C.
difficile} challenge (day 0) for their relationship to infection
outcomes using linear discriminant analysis (LDA) effect size (LEfSe)
analysis to identify individual bacterial populations that could explain
the variation in disease severity. We split the mice into groups by
severity level based on their moribundity and histopathologic score. We
dichotomized the histopathologic scores into high and low groups by
splitting on the median score. This analysis revealed 20 genera that
were significantly different by the disease severity (Figure 4A).
Bacterial genera \emph{Turicibacter}, \emph{Streptococcus},
\emph{Staphylococcus}, \emph{Pseudomonas}, \emph{Phocaeicola},
\emph{Parabacteroides}, \emph{Bacteroides}, and
\emph{Escherichia/Shigella} were detected at higher relative abundances
in the mice that became moribund. Populations of \emph{Anaerotignum},
\emph{Coprobacillus}, \emph{Enterocloster}, and \emph{Murimonas} were
more abundant in the non-moribund mice that would develop only low
intestinal injury. To understand the role of toxin activity in disease
severity, we applied LEfSe to identify the genera most likely to explain
the observed differences in toxin activity (Figure 4B). Many genera that
associated with toxin production were also associated with moribundity,
such as populations of \emph{Escherichia/Shigella} and
\emph{Bacteroides}. Likewise, there were genera such as
\emph{Anaerotignum}, \emph{Enterocloster}, and \emph{Murimonas} that
were associated with no toxin production that also exhibited greater
relative abundance in communities from non-moribund mice with a low
histopathologic score. Lastly, we tested for correlations between the
endpoint relative abundances of bacterial operational taxonomic units
(OTUs) and the histopathologic summary score (Figure 4C). The endpoint
relative abundance of \emph{Bacteroides} was positively correlated with
histopathologic score, as its day 0 relative abundance did with disease
severity (Figure 4A). Populations of \emph{Klebsiella} and
\emph{Prevotellaceae} were positively correlated with the
histopathologic score and were increased in the group of mice with
detectable toxin. This analysis identified bacterial genera that were
associated with the variation in moribundity, histopathologic score, and
toxin activity of CDI.

We next determined whether collectively, bacterial community membership
and relative abundance could be predictive of the CDI disease outcome.
We trained random forest models with bacterial community relative
abundance data from the day of colonization at each taxonomic rank to
predict toxin production, moribundity, and day 10 post-challenge
histopathologic summary score. For predicting if toxin would be
produced, microbial populations aggregated by phylum rank classification
performed similarly as models using lower taxonomic ranks (AUROC = 0.83,
Figure S2). \emph{C. difficile} was more likely to produce toxin when
the community it infected had less abundant populations of
\emph{Verrucomicrobia} and \emph{Campilobacterota} and had more abundant
populations of \emph{Proteobacteria} (Figure 5A). Next, we assessed the
ability of the community to predict moribundity. Bacteria grouped by
class rank classification was sufficient to predict which mice would
succumb to the infection before the end of the experiment (AUROC = 0.91,
Figure S2). The features with the greatest effect showed that
communities with greater populations of bacteria belonging to
\emph{Bacilli} and \emph{Firmicutes} and reduced populations of
\emph{Erysipelotrichia} were more likely to result in moribundity
(Figure 5B). Only one other class of bacteria was decreased in moribund
mice, a group of unclassified \emph{Clostridia}. Lastly, the relative
abundances of genera were able to predict a high or low histopathologic
score (histopathologic scores were dichotomized as in previous analysis,
AUROC = 0.99, Figure S2). No genera had a significantly greater effect
on the model performance than any others, indicating the model was
reliant on many genera for the correct prediction. The model used some
of the genera identified in the LEfSe analysis, such as
\emph{Coprobacillus}, \emph{Anaerostipes}, and \emph{Hungatella}.
Communities with greater abundances of \emph{Hungatella},
\emph{Eggerthella}, \emph{Bifidobacterium}, \emph{Duncaniella} and
\emph{Neisseria} were more likely to have high histopathologic scores.
These models have shown that the relative abundance of bacterial
populations and their relationship to each other could be used to
predict the variation in moribundity, histopathologic score, and toxin
activity of CDI.

\textbf{Disease severity consistent by donor community across isolates.}
Since Carlson et al.~showed \emph{C. difficile} isolates with lower
germination rates were more likely to result in severe disease (27), we
next tested whether the effect of the community on CDI outcomes was
conserved across \emph{C. difficile} isolates with higher rates of
germination. We colonized mice with human fecal slurry of either Donor
N7, N6, or M4 and allowed two weeks for the community to establish. We
then challenged the mice with a different \emph{C. difficile} isolate
and all mice were infected. The moribundity and histopathologic score
outcomes when challenged with a different \emph{C. difficile} human
isolate were comparable to the outcomes of challenges with \emph{C.
difficile} RT027 human isolate 431 by their donor group (Figure 6). For
the isolates and communities we tested, the effect of the gut bacterial
community on disease severity was greater than the effect of the
variation in these isolates.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

Challenging mice colonized with different human fecal communities with a
single \emph{C. difficile} isolate, RT027 isolate 431, demonstrated that
variation in the gut bacteria affects the \emph{C. difficile} infection
disease severity. Our analysis revealed an association between the
relative abundance of bacterial community members and disease severity.
We used the relative abundance patterns of community members associated
with disease outcomes to select donor communities to test whether the
disease severity was conserved when challenged with a different \emph{C.
difficile} isolate. The disease outcomes from the challenge with the
second \emph{C. difficile} isolate matched the outcomes of the challenge
with \emph{C. difficile} RT027 isolate 431. Therefore, the effect of the
community variation on disease severity \emph{in vivo} was greater than
the effect of \emph{C. difficile} isolate variation (27, 28). Overall,
these results have demonstrated that the bacterial community contributed
to the disease severity of \emph{C. difficile} infection.

CDI severity can range from no symptoms to mild symptoms, such as
diarrhea, to severe symptoms, such as toxic megacolon or death.
Physicians use prediction tools to identify patients most at risk of
developing a severe infection using white blood cell counts, serum
albumin level, or serum creatinine level (2, 29, 30). Those levels are
driven by the activities in the intestine (31). Research into the
drivers of this variation have revealed factors that make \emph{C.
difficile} more virulent. Strains are categorized for their virulence by
the presence and production toxin TcdA, toxin TcdB, and binary toxin and
the prevalence in outbreaks, such as ribotypes 027 and 078 (22, 32--35).
However, other studies have shown that virulence is not necessarily
linked with toxin production (36) or the strain (37). Furthermore, there
is variation in the genome, growth rate, sporulation, germination, and
toxin production in different isolates of a strain (27, 28, 38, 39).
This variation may help explain why severe CDI prediction tools often
miss identifying many patients with CDI that will develop severe disease
(3--6). Therefore, it is necessary to gain a full understanding of all
factors contributing to disease variation to improve our ability to
predict severity.

The state of the gut bacterial community determines the ability of
\emph{C. difficile} to colonize and persist in the intestine. \emph{C.
difficile} is unable to colonize an unperturbed healthy gut community
and is only able to become established after a perturbation (23). Once
colonized, the different communities lead to different metabolic
responses and dynamics of the \emph{C. difficile} population (11, 18,
40). The gut bacteria metabolize primary bile acids into secondary bile
acids (41, 42). The concentration of these bile acids affect the
germination, growth, toxin production and biofilm formation (12, 13, 43,
44). Members of the bacterial community also affect other metabolites
\emph{C. difficile} utilizes. \emph{Bacteroides thetaiotaomicron}
produce sialidases which release sialic acid from the mucosa for
\emph{C. difficile} to utilize (45, 46). The nutrient environment
affects toxin production (47). Thus, many of the actions of the gut
bacteria modulate \emph{C. difficile} in ways that could affect the
infection and resultant disease.

Myriad studies have explored the relationship between the microbiome and
CDI but none have established a relationship between the variation of
disease severity and the microbiome. CDI studies often use an infection
model with a single homogeneous bacterial community. Collins et al.~used
multiple human communities to colonize mice, however the communities
were pooled prior to gavaging into germ-free mice (48), resulting in a
single community. Studies examining difference in disease often use
different \emph{C. difficile} strains or ribotypes in mice with similar
microbiota as a proxy for variation in disease, such as strain 630 for
non-severe and RT027 for severe (22, 32, 33, 49). Studies have also
demonstrated variation in infection through tapering antibiotic dosage
(18, 23, 50) or by reducing the amount of \emph{C. difficile} cells or
spores used for the challenge (22, 50). These studies either lack
variation in the initial microbiome or have variation in the \emph{C.
difficile} infection itself, therefore we cannot make any association
between variation in severity and the microbiome.

With our recent observations that the initial community affected the
ability of \emph{C. difficile} to persist in the gut (17, 18) and the
existing research describing the myriad interactions between the
microbiome and \emph{C. difficile}, we hypothesized that the microbiome
can modulate CDI disease severity. Our data have demonstrated gut
bacterial relative abundances associate with variation in toxin
production, histopathologic scoring of the cecal tissue and mortality.
This analysis revealed populations of \emph{Akkermansia},
\emph{Anaerostipes}, \emph{Coprobacillus}, \emph{Enterocloster},
\emph{Lactonifactor}, and \emph{Monoglobus} were more abundant in the
microbiome of non-moribund mice which had low histopathologic scores and
no detected toxin. The protective role of these genera are supported by
previous studies. \emph{Coprobacillus}, \emph{Lactonifactor}, and
\emph{Monoglobus} have been shown to be involved in dietary fiber
fermentation and associated with healthy communities (51--54).
\emph{Anaerostipes} and \emph{Coprobacillus}, which produce short chain
fatty acids, have been associated with healthy communities (55--57).
Furthermore, \emph{Coprobacillus}, which was abundant in mice with low
histopathologic scores but rare in all other mice, has been shown to
contain a putative type I lantibiotic gene cluster and inhibit \emph{C.
difficile} colonization (58--60). \emph{Akkermansia} and
\emph{Enterocloster} were also identified as more abundant in mice which
had a low histopathologic scores but have contradictory supporting
evidence in the current literature. In our data, \emph{Akkermansia} was
most abundant in the non-moribund mice with low histopathologic scores
but there were some moribund mice which had increased populations of
\emph{Akkermansia}. This could be attributed to either a more protective
mucus layer was present inhibiting colonization (60, 61) or mucus
consumption by \emph{Akkermansia} could have been crossfeeding \emph{C.
difficile} or exposing a niche for \emph{C. difficile} (62--64).
Similarly, \emph{Enterocloster} was more abundant and associated with
low histopathologic scores. It has been associated with healthy
populations and has been used to mono-colonize germ-free mice to reduce
the ability of \emph{C. difficile} to colonize (65, 66). However,
\emph{Enterocloster} has also been involved in infections, such as
bacteremia (67, 68). These data have exemplified populations of bacteria
that have the potential to be either protective or harmful. Thus, the
disease outcome is not likely based on the abundance of individual
populations of bacteria, rather it is the result of the interactions of
the community.

The groups of bacteria that were associated with either a higher
histopathologic score or moribundity are members of the indigenous gut
community that also have been associated with disease, often referred to
as opportunistic pathogens. Many of the populations with pathogenic
potential that associated with worse outcomes are also facultative
anaerobes. \emph{Enterococcus}, \emph{Klebsiella},
\emph{Shigella/Escherichia}, \emph{Staphylococcus}, and
\emph{Streptococcus} have been shown to expand after antibiotic use (20,
69, 70) and are commonly detected in CDI cases (71--74). In addition to
these populations, \emph{Eggerthella}, \emph{Prevotellaceae} and
\emph{Helicobacter}, which associated with worse outcomes, have also
been associated with intestinal inflammation (75--77). Recently,
\emph{Helicobacter hepaticus} was shown to be sufficient to cause
susceptibility to CDI in IL-10 deficient C57BL/6 mice (78). In our
experiments, when \emph{Helicobacter} was present, the infection
resulted in a high histopathologic score. While we did not use IL-10
deficient mice, it is possible the bacterial community or host response
are similarly modified by \emph{Helicobacter}, allowing \emph{C.
difficile} infection and host damage. Aside from \emph{Helicobacter},
these groups of bacteria that associated with more severe outcomes did
not have a conserved association between their relative abundance and
the disease severity across all mice.

Since we observed groups of bacteria that were associated with less
severe disease it may be appropriate to apply the damage-response
framework for microbial pathogenesis to CDI (79, 80). This framework
posits that disease is not driven by a single entity, rather it is an
emergent property of the responses of the host immune system, infecting
microbe, \emph{C. difficile}, and the indigenous microbes at the site of
infection. In the first set of experiments, we used the same host,
C57BL/6 mice, the same infecting microbe, \emph{C. difficile} RT027
human isolate 431, with different gut bacterial communities. The
bacterial groups in those communities were often present in both
moribund and non-moribund and across the range of histopathologic
scores. Thus, it was not merely the presence of the bacteria but their
activity in response to the other microbes and host which affect the
extent of the host damage. Additionally, while each mouse and \emph{C.
difficile} population had the same genetic background, they too were
reacting to the specific microbial community. Disease severity is driven
by the cumulative effect of the host immune response and the activity of
\emph{C. difficile} and the gut bacteria. \emph{C. difficile} can drive
host damage through the production of toxin. The gut microbiota can
contribute to the host damage through the balance of metabolic and
competitive activities, such as bacteriocin production or mucin
degradation. For example, low levels of mucin degradation can provide
nutrients to other community members producing a diverse non-damaging
community (81). However, if mucin degradation becomes too great it
reduces the protective function of the mucin layer and exposes the
epithelial cells. This over-harvesting can contribute to the host damage
due to other members producing toxin. Thus, the resultant intestinal
damage is the balance of all activities in the gut environment. Host
damage is the emergent property of numerous damage-response curves, such
as one for host immune response, one for \emph{C. difficile} activity
and another for microbiome community activity, each of which are a
composite curve of the individual activities from each group, such as
antibody production, neutrophil infiltration, toxin production,
sporulation, fiber and mucin degradation. Therefore, while we have
identified populations of interest, it may be necessary to target
multiple types of bacteria to reduce the community interactions
contributing to host damage.

Here we have shown myriad bacterial groups and their relative abundances
associated with variation in severity of CDI disease. Further
understanding of how the microbiome affects severity in clinical
patients could reduce the amount of adverse outcomes of CDI. When a
patient is diagnosed with CDI, the gut community composition, in
addition to the traditionally obtained clinical information, may improve
our severity prediction and guide prophylactic treatment. Treating the
microbiome at the time of diagnosis may prevent the infection from
becoming more severe. Based on the general trends of our results,
promoting fiber metabolizing bacteria and reducing facultative anaerobes
may bolster the efficacy of current CDI treatments, reducing disease
severity.

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\textbf{Animal care.} 6- to 13-week old male and female germ-free
C57BL/6 were obtained from a single breeding colony in the University of
Michigan Germ-free Mouse Core. Mice were housed in cages of 2-4 mice per
cage and maintained in germ-free isolators at the University of Michigan
germ-free facility. All mouse experiments were approved by the
University Committee on Use and Care of Animals at the University of
Michigan. \textbf{Were the mice maintained in gnotobiotic isolators
throughout the experiment? Or were they transferred to BSL2 conditions
prior to administering the human FMTs?}

\textbf{\emph{C. difficile} experiments.} Human fecal samples were
obtained as part of Schubert et al.~and selected based on community
clusters (20) to result in diverse community structures. Feces were
homogenized by mixing 200 mg of sample with 5 ml of PBS. Mice were
inoculated with 100 \(\mu\)l of the fecal homogenate via oral gavage.
For the set of experiments testing multiple \emph{C. difficile} clinical
isolates, we selected a subset of donor fecal samples based on their
associated mouse transplant communities and the outcomes in the
experiments with \emph{C. difficile} RT027 isolate 431 that fit the
following criteria: one community we would expect to result in a low
histopathologic score (\textless{} 5), a high histopathologic score
(\textgreater{} 5), and a high histopathologic score which becomes
moribund. Additionally, we selected the donors based their relative
abundances of genera identified through the our analysis (Figure S3).
Two weeks after the fecal community inoculation, mice were challenged
with \emph{C. difficile}. \emph{C. difficile} isolates came from Carlson
et al.~which had previously been isolated and characterized (27, 28) and
have recently been further characterized (38). Spores concentration were
determined both before and after challenge (82). \(10^{3}\) \emph{C.
difficile} spores were given to each mouse via oral gavage. \textbf{How
were human FMTs prepared (anaerobically vs aerobically, etc)?}

\textbf{Sample collection.} Fecal samples were collected on the day of
\emph{C. difficile} challenge and the following 10 days. Each day, a
fecal sample was also collected and weighed and the remaining sample was
frozen at -20\(^\circ\)C. Anaerobically, the weighed fecal samples were
serially diluted in PBS, plated on TCCFA plates, and incubated at
37\(^\circ\)C for 24 hours. The plates were then counted for the number
of colony forming units (CFU) (83).

\textbf{DNA sequencing.} From the frozen fecal samples, total bacterial
DNA was extracted using MOBIO PowerSoil-htp 96-well soil DNA isolation
kit. We amplified the 16S rRNA gene V4 region and sequenced the
resulting amplicons using an Illumina MiSeq as described previously
(84).

\textbf{Sequence curation.} Sequences were processed with
mothur(v.1.44.3) as previously described (84, 85). In short, we used a
3\% dissimilarity cutoff to group sequences into operational taxonomic
units (OTUs). We used a naive Bayesian classifier with the Ribosomal
Database Project training set (version 18) to assign taxonomic
classifications to each OTU (86). We sequenced a mock community of a
known community composition and 16s rRNA gene sequences. We processed
this mock community with our samples to calculate the error rate for our
sequence curation, which was an error rate of 0.19\%.

\textbf{Toxin cytotoxicity assay.} To prepare the sample for the
activity assay, fecal material was diluted 1:10 weight per volume using
sterile PBS and then filter sterilized through a 0.22-\(\mu\)m filter.
Toxin activity was assessed using a Vero cell rounding-based
cytotoxicity assay as described previously (33). The cytotoxicity titer
was determined for each sample as the last dilution, which resulted in
at least 80\% cell rounding. Toxin titers are reported as the log10 of
the reciprocal of the cytotoxicity titer.

\textbf{Histopathology evaluation.} Mouse cecal tissue was placed in
histopathology cassettes and fixed in 10\% formalin, then stored in 70\%
ethanol. McClinchey Histology Labs, Inc.~(Stockbridge, MI) embedded the
samples in paraffin, sectioned, and created the hematoxylin and
eosin-stained slides. The slides were scored using previously described
criteria by a board-certified veterinary pathologist scored who was
blinded to the experimental groups (33).

\textbf{Statistical analysis and modeling.} To compare community
structures, we calculated Yue and Clayton dissimilarity matrices
(\(\theta\)\textsubscript{YC}) in mothur (87). We rarefied samples to
2,107 sequences per sample to limit uneven sampling biases. We tested
for differences in individual taxonomic groups that would explain the
outcome differences with LEfSe (88) in mothur. Remaining statistical
analysis and data visualization was performed in R (v4.0.5) with the
tidyverse package (v1.3.1). We tested for significant differences in
\(\beta\)-diversity (\(\theta\)\textsubscript{YC}) using the Wilcoxon
rank sum test. We used Spearman's correlation to identified which OTUs
that had a correlation between their relative abundance and the
histopathologic summary score. \emph{P} values were then corrected for
multiple comparisons with a Benjamini and Hochberg adjustment for a type
I error rate of 0.05 (89). We built random forest models using the
mikropml package (90) with OTUs from day 0 samples using mtry values of
1 through 10, 15, 20, 25, 40, 50, 100. The split for training and
testing varied by model to avoid overfitting the data. The toxin and
moribundity models were trained with 60\% of the data. The
histopathologic score model was trained with 80\% of the data. Lastly,
we did not analyze murine communities to donor community or clinical
data because germ-free mice colonized with human fecal communities
become more like indigenous murine communities once stabilized.
Furthermore, we are not trying to make any claims about human bacteria
and their association with human CDI.

\textbf{Code availability.} Scripts necessary to reproduce our analysis
and this paper are available in an online repository
(\url{https://github.com/SchlossLab/Lesniak_Severity_XXXX_2022}).

\textbf{Sequence data accession number.} All 16S rRNA gene sequence data
and associated metadata are available through the Sequence Read Archive
via accession PRJNA787941.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

Thank you to Sarah Lucas and Sarah Tomkovich for critical discussion in
the development and execution of this project. We also thank the
University of Michigan Germ-free Mouse Core for assistance with our
germfree mice, funded in part by U2CDK110768. This work was supported by
several grants from the National Institutes for Health R01GM099514,
U19AI090871, U01AI12455, and P30DK034933. Additionally, NAL was
supported by the Molecular Mechanisms of Microbial Pathogenesis training
grant (NIH T32 AI007528). The funding agencies had no role in study
design, data collection and analysis, decision to publish, or
preparation of the manuscript.

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Kelly2008}{}%
1. \textbf{Kelly CP}, \textbf{LaMont JT}. 2008. \emph{Clostridium
difficile} more difficult than ever. New England Journal of Medicine
\textbf{359}:1932--1940.
doi:\href{https://doi.org/10.1056/nejmra0707500}{10.1056/nejmra0707500}.

\leavevmode\hypertarget{ref-McDonald2018}{}%
2. \textbf{McDonald LC}, \textbf{Gerding DN}, \textbf{Johnson S},
\textbf{Bakken JS}, \textbf{Carroll KC}, \textbf{Coffin SE},
\textbf{Dubberke ER}, \textbf{Garey KW}, \textbf{Gould CV},
\textbf{Kelly C}, \textbf{Loo V}, \textbf{Sammons JS}, \textbf{Sandora
TJ}, \textbf{Wilcox MH}. 2018. Clinical practice guidelines for
\emph{Clostridium difficile} infection in adults and children: 2017
update by the infectious diseases society of america (IDSA) and society
for healthcare epidemiology of america (SHEA). Clinical Infectious
Diseases \textbf{66}:e1--e48.
doi:\href{https://doi.org/10.1093/cid/cix1085}{10.1093/cid/cix1085}.

\leavevmode\hypertarget{ref-Dieterle2020}{}%
3. \textbf{Dieterle MG}, \textbf{Putler R}, \textbf{Perry DA},
\textbf{Menon A}, \textbf{Abernathy-Close L}, \textbf{Perlman NS},
\textbf{Penkevich A}, \textbf{Standke A}, \textbf{Keidan M},
\textbf{Vendrov KC}, \textbf{Bergin IL}, \textbf{Young VB}, \textbf{Rao
K}. 2020. Systemic inflammatory mediators are effective biomarkers for
predicting adverse outcomes in \emph{Clostridioides difficile}
infection. mBio \textbf{11}.
doi:\href{https://doi.org/10.1128/mbio.00180-20}{10.1128/mbio.00180-20}.

\leavevmode\hypertarget{ref-Butt2013}{}%
4. \textbf{Butt E}, \textbf{Foster JA}, \textbf{Keedwell E},
\textbf{Bell JE}, \textbf{Titball RW}, \textbf{Bhangu A},
\textbf{Michell SL}, \textbf{Sheridan R}. 2013. Derivation and
validation of a simple, accurate and robust prediction rule for risk of
mortality in patients with \emph{Clostridium difficile} infection. BMC
Infectious Diseases \textbf{13}.
doi:\href{https://doi.org/10.1186/1471-2334-13-316}{10.1186/1471-2334-13-316}.

\leavevmode\hypertarget{ref-Perry2021}{}%
5. \textbf{Perry DA}, \textbf{Shirley D}, \textbf{Micic D},
\textbf{Patel CP}, \textbf{Putler R}, \textbf{Menon A}, \textbf{Young
VB}, \textbf{Rao K}. 2021. External validation and comparison of
\emph{Clostridioides difficile} severity scoring systems. Clinical
Infectious Diseases.
doi:\href{https://doi.org/10.1093/cid/ciab737}{10.1093/cid/ciab737}.

\leavevmode\hypertarget{ref-vanBeurden2017}{}%
6. \textbf{Beurden YH van}, \textbf{Hensgens MPM}, \textbf{Dekkers OM},
\textbf{Cessie SL}, \textbf{Mulder CJJ}, \textbf{Vandenbroucke-Grauls
CMJE}. 2017. External validation of three prediction tools for patients
at risk of a complicated course of \emph{Clostridium difficile}
infection: Disappointing in an outbreak setting. Infection Control \&
Hospital Epidemiology \textbf{38}:897--905.
doi:\href{https://doi.org/10.1017/ice.2017.89}{10.1017/ice.2017.89}.

\leavevmode\hypertarget{ref-Britton2014}{}%
7. \textbf{Britton RA}, \textbf{Young VB}. 2014. Role of the intestinal
microbiota in resistance to colonization by \emph{Clostridium
difficile}. Gastroenterology \textbf{146}:1547--1553.
doi:\href{https://doi.org/10.1053/j.gastro.2014.01.059}{10.1053/j.gastro.2014.01.059}.

\leavevmode\hypertarget{ref-Hryckowian2018}{}%
8. \textbf{Hryckowian AJ}, \textbf{Treuren WV}, \textbf{Smits SA},
\textbf{Davis NM}, \textbf{Gardner JO}, \textbf{Bouley DM},
\textbf{Sonnenburg JL}. 2018. Microbiota-accessible carbohydrates
suppress \emph{Clostridium difficile} infection in a murine model.
Nature Microbiology \textbf{3}:662--669.
doi:\href{https://doi.org/10.1038/s41564-018-0150-6}{10.1038/s41564-018-0150-6}.

\leavevmode\hypertarget{ref-VichVila2020}{}%
9. \textbf{Vila AV}, \textbf{Collij V}, \textbf{Sanna S}, \textbf{Sinha
T}, \textbf{Imhann F}, \textbf{Bourgonje AR}, \textbf{Mujagic Z},
\textbf{Jonkers DMAE}, \textbf{Masclee AAM}, \textbf{Fu J},
\textbf{Kurilshikov A}, \textbf{Wijmenga C}, \textbf{Zhernakova A},
\textbf{Weersma RK}. 2020. Impact of commonly used drugs on the
composition and metabolic function of the gut microbiota. Nature
Communications \textbf{11}.
doi:\href{https://doi.org/10.1038/s41467-019-14177-z}{10.1038/s41467-019-14177-z}.

\leavevmode\hypertarget{ref-Abbas2020}{}%
10. \textbf{Abbas A}, \textbf{Zackular JP}. 2020. Microbe-microbe
interactions during \emph{Clostridioides difficile} infection. Current
Opinion in Microbiology \textbf{53}:19--25.
doi:\href{https://doi.org/10.1016/j.mib.2020.01.016}{10.1016/j.mib.2020.01.016}.

\leavevmode\hypertarget{ref-Jenior2017}{}%
11. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. \emph{Clostridium difficile} colonizes
alternative nutrient niches during infection across distinct murine gut
microbiomes. mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\leavevmode\hypertarget{ref-Sorg2008}{}%
12. \textbf{Sorg JA}, \textbf{Sonenshein AL}. 2008. Bile salts and
glycine as cogerminants for \emph{Clostridium difficile} spores. Journal
of Bacteriology \textbf{190}:2505--2512.
doi:\href{https://doi.org/10.1128/jb.01765-07}{10.1128/jb.01765-07}.

\leavevmode\hypertarget{ref-Thanissery2017}{}%
13. \textbf{Thanissery R}, \textbf{Winston JA}, \textbf{Theriot CM}.
2017. Inhibition of spore germination, growth, and toxin activity of
clinically relevant \emph{C. difficile} strains by gut microbiota
derived secondary bile acids. Anaerobe \textbf{45}:86--100.
doi:\href{https://doi.org/10.1016/j.anaerobe.2017.03.004}{10.1016/j.anaerobe.2017.03.004}.

\leavevmode\hypertarget{ref-Aguirre2021}{}%
14. \textbf{Aguirre AM}, \textbf{Yalcinkaya N}, \textbf{Wu Q},
\textbf{Swennes A}, \textbf{Tessier ME}, \textbf{Roberts P},
\textbf{Miyajima F}, \textbf{Savidge T}, \textbf{Sorg JA}. 2021. Bile
acid-independent protection against \emph{Clostridioides difficile}
infection. PLOS Pathogens \textbf{17}:e1010015.
doi:\href{https://doi.org/10.1371/journal.ppat.1010015}{10.1371/journal.ppat.1010015}.

\leavevmode\hypertarget{ref-Kang2019}{}%
15. \textbf{Kang JD}, \textbf{Myers CJ}, \textbf{Harris SC},
\textbf{Kakiyama G}, \textbf{Lee I-K}, \textbf{Yun B-S},
\textbf{Matsuzaki K}, \textbf{Furukawa M}, \textbf{Min H-K},
\textbf{Bajaj JS}, \textbf{Zhou H}, \textbf{Hylemon PB}. 2019. Bile acid
7\(\alpha\)-dehydroxylating gut bacteria secrete antibiotics that
inhibit \emph{Clostridium difficile}: Role of secondary bile acids. Cell
Chemical Biology \textbf{26}:27--34.e4.
doi:\href{https://doi.org/10.1016/j.chembiol.2018.10.003}{10.1016/j.chembiol.2018.10.003}.

\leavevmode\hypertarget{ref-Seekatz2016}{}%
16. \textbf{Seekatz AM}, \textbf{Rao K}, \textbf{Santhosh K},
\textbf{Young VB}. 2016. Dynamics of the fecal microbiome in patients
with recurrent and nonrecurrent \emph{Clostridium difficile} infection.
Genome Medicine \textbf{8}.
doi:\href{https://doi.org/10.1186/s13073-016-0298-8}{10.1186/s13073-016-0298-8}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
17. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence \emph{Clostridioides difficile}
clearance in mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.

\leavevmode\hypertarget{ref-Lesniak2021}{}%
18. \textbf{Lesniak NA}, \textbf{Schubert AM}, \textbf{Sinani H},
\textbf{Schloss PD}. 2021. Clearance of \emph{Clostridioides difficile}
colonization is associated with antibiotic-specific bacterial changes.
mSphere \textbf{6}.
doi:\href{https://doi.org/10.1128/msphere.01238-20}{10.1128/msphere.01238-20}.

\leavevmode\hypertarget{ref-Nagpal2018}{}%
19. \textbf{Nagpal R}, \textbf{Wang S}, \textbf{Woods LCS},
\textbf{Seshie O}, \textbf{Chung ST}, \textbf{Shively CA},
\textbf{Register TC}, \textbf{Craft S}, \textbf{McClain DA},
\textbf{Yadav H}. 2018. Comparative microbiome signatures and
short-chain fatty acids in mouse, rat, non-human primate, and human
feces. Frontiers in Microbiology \textbf{9}.
doi:\href{https://doi.org/10.3389/fmicb.2018.02897}{10.3389/fmicb.2018.02897}.

\leavevmode\hypertarget{ref-Schubert2014}{}%
20. \textbf{Schubert AM}, \textbf{Rogers MAM}, \textbf{Ring C},
\textbf{Mogle J}, \textbf{Petrosino JP}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Schloss PD}. 2014. Microbiome data
distinguish patients with \emph{Clostridium difficile} infection and
non-\emph{C. difficile}-associated diarrhea from healthy controls. mBio
\textbf{5}.
doi:\href{https://doi.org/10.1128/mbio.01021-14}{10.1128/mbio.01021-14}.

\leavevmode\hypertarget{ref-Gillilland2012}{}%
21. \textbf{Gillilland MG}, \textbf{Erb-Downward JR}, \textbf{Bassis
CM}, \textbf{Shen MC}, \textbf{Toews GB}, \textbf{Young VB},
\textbf{Huffnagle GB}. 2012. Ecological succession of bacterial
communities during conventionalization of germ-free mice. Applied and
Environmental Microbiology \textbf{78}:2359--2366.
doi:\href{https://doi.org/10.1128/aem.05239-11}{10.1128/aem.05239-11}.

\leavevmode\hypertarget{ref-Chen2008}{}%
22. \textbf{Chen X}, \textbf{Katchar K}, \textbf{Goldsmith JD},
\textbf{Nanthakumar N}, \textbf{Cheknis A}, \textbf{Gerding DN},
\textbf{Kelly CP}. 2008. A mouse model of \emph{Clostridium
difficile}-associated disease. Gastroenterology \textbf{135}:1984--1992.
doi:\href{https://doi.org/10.1053/j.gastro.2008.09.002}{10.1053/j.gastro.2008.09.002}.

\leavevmode\hypertarget{ref-Schubert2015}{}%
23. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\leavevmode\hypertarget{ref-Akerlund2006}{}%
24. \textbf{Åkerlund T}, \textbf{Svenungsson B}, \textbf{Lagergren},
\textbf{Burman LG}. 2006. Correlation of disease severity with fecal
toxin levels in patients with \emph{Clostridium difficile}-associated
diarrhea and distribution of PCR ribotypes and toxin yields in vitro of
corresponding isolates. Journal of Clinical Microbiology
\textbf{44}:353--358.
doi:\href{https://doi.org/10.1128/jcm.44.2.353-358.2006}{10.1128/jcm.44.2.353-358.2006}.

\leavevmode\hypertarget{ref-Vitucci2020}{}%
25. \textbf{Vitucci JC}, \textbf{Pulse M}, \textbf{Tabor-Simecka L},
\textbf{Simecka J}. 2020. Epidemic ribotypes of \emph{Clostridium} (now
\emph{Clostridioides}) \emph{difficile} are likely to be more virulent
than non-epidemic ribotypes in animal models. BMC Microbiology
\textbf{20}.
doi:\href{https://doi.org/10.1186/s12866-020-1710-5}{10.1186/s12866-020-1710-5}.

\leavevmode\hypertarget{ref-Cowardin2016}{}%
26. \textbf{Cowardin CA}, \textbf{Buonomo EL}, \textbf{Saleh MM},
\textbf{Wilson MG}, \textbf{Burgess SL}, \textbf{Kuehne SA},
\textbf{Schwan C}, \textbf{Eichhoff AM}, \textbf{Koch-Nolte F},
\textbf{Lyras D}, \textbf{Aktories K}, \textbf{Minton NP}, \textbf{Petri
WA}. 2016. The binary toxin CDT enhances \emph{Clostridium difficile}
virulence by suppressing protective colonic eosinophilia. Nature
Microbiology \textbf{1}.
doi:\href{https://doi.org/10.1038/nmicrobiol.2016.108}{10.1038/nmicrobiol.2016.108}.

\leavevmode\hypertarget{ref-Carlson2015}{}%
27. \textbf{Carlson PE}, \textbf{Kaiser AM}, \textbf{McColm SA},
\textbf{Bauer JM}, \textbf{Young VB}, \textbf{Aronoff DM}, \textbf{Hanna
PC}. 2015. Variation in germination of \emph{Clostridium difficile}
clinical isolates correlates to disease severity. Anaerobe
\textbf{33}:64--70.
doi:\href{https://doi.org/10.1016/j.anaerobe.2015.02.003}{10.1016/j.anaerobe.2015.02.003}.

\leavevmode\hypertarget{ref-Carlson2013}{}%
28. \textbf{Carlson PE}, \textbf{Walk ST}, \textbf{Bourgis AET},
\textbf{Liu MW}, \textbf{Kopliku F}, \textbf{Lo E}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Hanna PC}. 2013. The relationship between
phenotype, ribotype, and clinical disease in human \emph{Clostridium
difficile} isolates. Anaerobe \textbf{24}:109--116.
doi:\href{https://doi.org/10.1016/j.anaerobe.2013.04.003}{10.1016/j.anaerobe.2013.04.003}.

\leavevmode\hypertarget{ref-Lungulescu2011}{}%
29. \textbf{Lungulescu OA}, \textbf{Cao W}, \textbf{Gatskevich E},
\textbf{Tlhabano L}, \textbf{Stratidis JG}. 2011. CSI: A severity index
for \emph{Clostridium difficile} infection at the time of admission.
Journal of Hospital Infection \textbf{79}:151--154.
doi:\href{https://doi.org/10.1016/j.jhin.2011.04.017}{10.1016/j.jhin.2011.04.017}.

\leavevmode\hypertarget{ref-Zar2007}{}%
30. \textbf{Zar FA}, \textbf{Bakkanagari SR}, \textbf{Moorthi KMLST},
\textbf{Davis MB}. 2007. A comparison of vancomycin and metronidazole
for the treatment of \emph{Clostridium difficile}-associated diarrhea,
stratified by disease severity. Clinical Infectious Diseases
\textbf{45}:302--307.
doi:\href{https://doi.org/10.1086/519265}{10.1086/519265}.

\leavevmode\hypertarget{ref-diMasi2018}{}%
31. \textbf{Masi A di}, \textbf{Leboffe L}, \textbf{Polticelli F},
\textbf{Tonon F}, \textbf{Zennaro C}, \textbf{Caterino M}, \textbf{Stano
P}, \textbf{Fischer S}, \textbf{Hägele M}, \textbf{Müller M},
\textbf{Kleger A}, \textbf{Papatheodorou P}, \textbf{Nocca G},
\textbf{Arcovito A}, \textbf{Gori A}, \textbf{Ruoppolo M}, \textbf{Barth
H}, \textbf{Petrosillo N}, \textbf{Ascenzi P}, \textbf{Bella SD}. 2018.
Human serum albumin is an essential component of the host defense
mechanism against \emph{Clostridium difficile} intoxication. The Journal
of Infectious Diseases \textbf{218}:1424--1435.
doi:\href{https://doi.org/10.1093/infdis/jiy338}{10.1093/infdis/jiy338}.

\leavevmode\hypertarget{ref-AbernathyClose2020}{}%
32. \textbf{Abernathy-Close L}, \textbf{Dieterle MG}, \textbf{Vendrov
KC}, \textbf{Bergin IL}, \textbf{Rao K}, \textbf{Young VB}. 2020. Aging
dampens the intestinal innate immune response during severe
\emph{Clostridioides difficile} infection and is associated with altered
cytokine levels and granulocyte mobilization. Infection and Immunity
\textbf{88}.
doi:\href{https://doi.org/10.1128/iai.00960-19}{10.1128/iai.00960-19}.

\leavevmode\hypertarget{ref-Theriot2011}{}%
33. \textbf{Theriot CM}, \textbf{Koumpouras CC}, \textbf{Carlson PE},
\textbf{Bergin II}, \textbf{Aronoff DM}, \textbf{Young VB}. 2011.
Cefoperazone-treated mice as an experimental platform to assess
differential virulence of \emph{Clostridium difficile} strains. Gut
Microbes \textbf{2}:326--334.
doi:\href{https://doi.org/10.4161/gmic.19142}{10.4161/gmic.19142}.

\leavevmode\hypertarget{ref-Goorhuis2008}{}%
34. \textbf{Goorhuis A}, \textbf{Bakker D}, \textbf{Corver J},
\textbf{Debast SB}, \textbf{Harmanus C}, \textbf{Notermans DW},
\textbf{Bergwerff AA}, \textbf{Dekker FW}, \textbf{Kuijper EJ}. 2008.
Emergence of \emph{Clostridium difficile} infection due to a new
hypervirulent strain, polymerase chain reaction ribotype 078. Clinical
Infectious Diseases \textbf{47}:1162--1170.
doi:\href{https://doi.org/10.1086/592257}{10.1086/592257}.

\leavevmode\hypertarget{ref-OConnor2009}{}%
35. \textbf{O'Connor JR}, \textbf{Johnson S}, \textbf{Gerding DN}. 2009.
\emph{Clostridium difficile} infection caused by the epidemic
BI/NAP1/027 strain. Gastroenterology \textbf{136}:1913--1924.
doi:\href{https://doi.org/10.1053/j.gastro.2009.02.073}{10.1053/j.gastro.2009.02.073}.

\leavevmode\hypertarget{ref-Rao2015}{}%
36. \textbf{Rao K}, \textbf{Micic D}, \textbf{Natarajan M},
\textbf{Winters S}, \textbf{Kiel MJ}, \textbf{Walk ST}, \textbf{Santhosh
K}, \textbf{Mogle JA}, \textbf{Galecki AT}, \textbf{LeBar W},
\textbf{Higgins PDR}, \textbf{Young VB}, \textbf{Aronoff DM}. 2015.
\emph{Clostridium difficile} ribotype 027: Relationship to age,
detectability of toxins A or B in stool with rapid testing, severe
infection, and mortality. Clinical Infectious Diseases
\textbf{61}:233--241.
doi:\href{https://doi.org/10.1093/cid/civ254}{10.1093/cid/civ254}.

\leavevmode\hypertarget{ref-Walk2012}{}%
37. \textbf{Walk ST}, \textbf{Micic D}, \textbf{Jain R}, \textbf{Lo ES},
\textbf{Trivedi I}, \textbf{Liu EW}, \textbf{Almassalha LM},
\textbf{Ewing SA}, \textbf{Ring C}, \textbf{Galecki AT}, \textbf{Rogers
MAM}, \textbf{Washer L}, \textbf{Newton DW}, \textbf{Malani PN},
\textbf{Young VB}, \textbf{Aronoff DM}. 2012. \emph{Clostridium
difficile} ribotype does not predict severe infection. Clinical
Infectious Diseases \textbf{55}:1661--1668.
doi:\href{https://doi.org/10.1093/cid/cis786}{10.1093/cid/cis786}.

\leavevmode\hypertarget{ref-Saund2021}{}%
38. \textbf{Saund K}, \textbf{Pirani A}, \textbf{Lacy B}, \textbf{Hanna
PC}, \textbf{Snitkin ES}. 2021. Strain variation in \emph{Clostridioides
difficile} toxin activity associated with genomic variation at both
PaLoc and non-PaLoc loci.
doi:\href{https://doi.org/10.1101/2021.12.08.471880}{10.1101/2021.12.08.471880}.

\leavevmode\hypertarget{ref-He2010}{}%
39. \textbf{He M}, \textbf{Sebaihia M}, \textbf{Lawley TD},
\textbf{Stabler RA}, \textbf{Dawson LF}, \textbf{Martin MJ},
\textbf{Holt KE}, \textbf{Seth-Smith HMB}, \textbf{Quail MA},
\textbf{Rance R}, \textbf{Brooks K}, \textbf{Churcher C}, \textbf{Harris
D}, \textbf{Bentley SD}, \textbf{Burrows C}, \textbf{Clark L},
\textbf{Corton C}, \textbf{Murray V}, \textbf{Rose G}, \textbf{Thurston
S}, \textbf{Tonder A van}, \textbf{Walker D}, \textbf{Wren BW},
\textbf{Dougan G}, \textbf{Parkhill J}. 2010. Evolutionary dynamics of
\emph{Clostridium difficile} over short and long time scales.
Proceedings of the National Academy of Sciences \textbf{107}:7527--7532.
doi:\href{https://doi.org/10.1073/pnas.0914322107}{10.1073/pnas.0914322107}.

\leavevmode\hypertarget{ref-Jenior2018}{}%
40. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. \emph{Clostridium difficile} alters the
structure and metabolism of distinct cecal microbiomes during initial
infection to promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\leavevmode\hypertarget{ref-Staley2016}{}%
41. \textbf{Staley C}, \textbf{Weingarden AR}, \textbf{Khoruts A},
\textbf{Sadowsky MJ}. 2016. Interaction of gut microbiota with bile acid
metabolism and its influence on disease states. Applied Microbiology and
Biotechnology \textbf{101}:47--64.
doi:\href{https://doi.org/10.1007/s00253-016-8006-6}{10.1007/s00253-016-8006-6}.

\leavevmode\hypertarget{ref-Long2017}{}%
42. \textbf{Long SL}, \textbf{Gahan CGM}, \textbf{Joyce SA}. 2017.
Interactions between gut bacteria and bile in health and disease.
Molecular Aspects of Medicine \textbf{56}:54--65.
doi:\href{https://doi.org/10.1016/j.mam.2017.06.002}{10.1016/j.mam.2017.06.002}.

\leavevmode\hypertarget{ref-Sorg2010}{}%
43. \textbf{Sorg JA}, \textbf{Sonenshein AL}. 2010. Inhibiting the
initiation of \emph{Clostridium difficile} spore germination using
analogs of chenodeoxycholic acid, a bile acid. Journal of Bacteriology
\textbf{192}:4983--4990.
doi:\href{https://doi.org/10.1128/jb.00610-10}{10.1128/jb.00610-10}.

\leavevmode\hypertarget{ref-Dubois2019}{}%
44. \textbf{Dubois T}, \textbf{Tremblay YDN}, \textbf{Hamiot A},
\textbf{Martin-Verstraete I}, \textbf{Deschamps J}, \textbf{Monot M},
\textbf{Briandet R}, \textbf{Dupuy B}. 2019. A microbiota-generated bile
salt induces biofilm formation in \emph{Clostridium difficile}. npj
Biofilms and Microbiomes \textbf{5}.
doi:\href{https://doi.org/10.1038/s41522-019-0087-4}{10.1038/s41522-019-0087-4}.

\leavevmode\hypertarget{ref-Ng2013}{}%
45. \textbf{Ng KM}, \textbf{Ferreyra JA}, \textbf{Higginbottom SK},
\textbf{Lynch JB}, \textbf{Kashyap PC}, \textbf{Gopinath S},
\textbf{Naidu N}, \textbf{Choudhury B}, \textbf{Weimer BC},
\textbf{Monack DM}, \textbf{Sonnenburg JL}. 2013. Microbiota-liberated
host sugars facilitate post-antibiotic expansion of enteric pathogens.
Nature \textbf{502}:96--99.
doi:\href{https://doi.org/10.1038/nature12503}{10.1038/nature12503}.

\leavevmode\hypertarget{ref-Ferreyra2014}{}%
46. \textbf{Ferreyra JA}, \textbf{Wu KJ}, \textbf{Hryckowian AJ},
\textbf{Bouley DM}, \textbf{Weimer BC}, \textbf{Sonnenburg JL}. 2014.
Gut microbiota-produced succinate promotes \emph{C. difficile} infection
after antibiotic treatment or motility disturbance. Cell Host \& Microbe
\textbf{16}:770--777.
doi:\href{https://doi.org/10.1016/j.chom.2014.11.003}{10.1016/j.chom.2014.11.003}.

\leavevmode\hypertarget{ref-MartinVerstraete2016}{}%
47. \textbf{Martin-Verstraete I}, \textbf{Peltier J}, \textbf{Dupuy B}.
2016. The regulatory networks that control \emph{Clostridium difficile}
toxin synthesis. Toxins \textbf{8}:153.
doi:\href{https://doi.org/10.3390/toxins8050153}{10.3390/toxins8050153}.

\leavevmode\hypertarget{ref-Collins2015}{}%
48. \textbf{Collins J}, \textbf{Auchtung JM}, \textbf{Schaefer L},
\textbf{Eaton KA}, \textbf{Britton RA}. 2015. Humanized microbiota mice
as a model of recurrent \emph{Clostridium difficile} disease. Microbiome
\textbf{3}.
doi:\href{https://doi.org/10.1186/s40168-015-0097-2}{10.1186/s40168-015-0097-2}.

\leavevmode\hypertarget{ref-Lawley2012}{}%
49. \textbf{Lawley TD}, \textbf{Clare S}, \textbf{Walker AW},
\textbf{Stares MD}, \textbf{Connor TR}, \textbf{Raisen C},
\textbf{Goulding D}, \textbf{Rad R}, \textbf{Schreiber F},
\textbf{Brandt C}, \textbf{Deakin LJ}, \textbf{Pickard DJ},
\textbf{Duncan SH}, \textbf{Flint HJ}, \textbf{Clark TG},
\textbf{Parkhill J}, \textbf{Dougan G}. 2012. Targeted restoration of
the intestinal microbiota with a simple, defined bacteriotherapy
resolves relapsing \emph{Clostridium difficile} disease in mice. PLoS
Pathogens \textbf{8}:e1002995.
doi:\href{https://doi.org/10.1371/journal.ppat.1002995}{10.1371/journal.ppat.1002995}.

\leavevmode\hypertarget{ref-Reeves2011}{}%
50. \textbf{Reeves AE}, \textbf{Theriot CM}, \textbf{Bergin IL},
\textbf{Huffnagle GB}, \textbf{Schloss PD}, \textbf{Young VB}. 2011. The
interplay between microbiome dynamics and pathogen dynamics in a murine
model of \emph{Clostridium difficile} infection. Gut Microbes
\textbf{2}:145--158.
doi:\href{https://doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}.

\leavevmode\hypertarget{ref-Mabrok2011}{}%
51. \textbf{Mabrok HB}, \textbf{Klopfleisch R}, \textbf{Ghanem KZ},
\textbf{Clavel T}, \textbf{Blaut M}, \textbf{Loh G}. 2011. Lignan
transformation by gut bacteria lowers tumor burden in a gnotobiotic rat
model of breast cancer. Carcinogenesis \textbf{33}:203--208.
doi:\href{https://doi.org/10.1093/carcin/bgr256}{10.1093/carcin/bgr256}.

\leavevmode\hypertarget{ref-Kim2019}{}%
52. \textbf{Kim CC}, \textbf{Healey GR}, \textbf{Kelly WJ},
\textbf{Patchett ML}, \textbf{Jordens Z}, \textbf{Tannock GW},
\textbf{Sims IM}, \textbf{Bell TJ}, \textbf{Hedderley D},
\textbf{Henrissat B}, \textbf{Rosendale DI}. 2019. Genomic insights from
\emph{Monoglobus pectinilyticus}: A pectin-degrading specialist
bacterium in the human colon. The ISME Journal \textbf{13}:1437--1456.
doi:\href{https://doi.org/10.1038/s41396-019-0363-6}{10.1038/s41396-019-0363-6}.

\leavevmode\hypertarget{ref-doPrado2021}{}%
53. \textbf{Prado SBR do}, \textbf{Minguzzi BT}, \textbf{Hoffmann C},
\textbf{Fabi JP}. 2021. Modulation of human gut microbiota by dietary
fibers from unripe and ripe papayas: Distinct polysaccharide degradation
using a colonic in vitro fermentation model. Food Chemistry
\textbf{348}:129071.
doi:\href{https://doi.org/10.1016/j.foodchem.2021.129071}{10.1016/j.foodchem.2021.129071}.

\leavevmode\hypertarget{ref-Muthuramalingam2019}{}%
54. \textbf{Muthuramalingam K}, \textbf{Singh V}, \textbf{Choi C},
\textbf{Choi SI}, \textbf{Kim YM}, \textbf{Unno T}, \textbf{Cho M}.
2019. Dietary intervention using (1, 3)/(1, 6)-\(\beta\)-glucan, a
fungus-derived soluble prebiotic ameliorates high-fat diet-induced
metabolic distress and alters beneficially the gut microbiota in mice
model. European Journal of Nutrition \textbf{59}:2617--2629.
doi:\href{https://doi.org/10.1007/s00394-019-02110-5}{10.1007/s00394-019-02110-5}.

\leavevmode\hypertarget{ref-Han2019}{}%
55. \textbf{Han S-H}, \textbf{Yi J}, \textbf{Kim J-H}, \textbf{Lee S},
\textbf{Moon H-W}. 2019. Composition of gut microbiota in patients with
toxigenic \emph{Clostridioides} (\emph{Clostridium}) \emph{difficile}:
Comparison between subgroups according to clinical criteria and toxin
gene load. PLOS ONE \textbf{14}:e0212626.
doi:\href{https://doi.org/10.1371/journal.pone.0212626}{10.1371/journal.pone.0212626}.

\leavevmode\hypertarget{ref-Duncan2004}{}%
56. \textbf{Duncan SH}, \textbf{Louis P}, \textbf{Flint HJ}. 2004.
Lactate-utilizing bacteria, isolated from human feces, that produce
butyrate as a major fermentation product. Applied and Environmental
Microbiology \textbf{70}:5810--5817.
doi:\href{https://doi.org/10.1128/aem.70.10.5810-5817.2004}{10.1128/aem.70.10.5810-5817.2004}.

\leavevmode\hypertarget{ref-Ye2018}{}%
57. \textbf{Ye J}, \textbf{Lv L}, \textbf{Wu W}, \textbf{Li Y},
\textbf{Shi D}, \textbf{Fang D}, \textbf{Guo F}, \textbf{Jiang H},
\textbf{Yan R}, \textbf{Ye W}, \textbf{Li L}. 2018. Butyrate protects
mice against methionine--choline-deficient diet-induced non-alcoholic
steatohepatitis by improving gut barrier function, attenuating
inflammation and reducing endotoxin levels. Frontiers in Microbiology
\textbf{9}.
doi:\href{https://doi.org/10.3389/fmicb.2018.01967}{10.3389/fmicb.2018.01967}.

\leavevmode\hypertarget{ref-Walsh2017}{}%
58. \textbf{Walsh CJ}, \textbf{Guinane CM}, \textbf{O'Toole PW},
\textbf{Cotter PD}. 2017. A profile hidden markov model to investigate
the distribution and frequency of LanB-encoding lantibiotic modification
genes in the human oral and gut microbiome. PeerJ \textbf{5}:e3254.
doi:\href{https://doi.org/10.7717/peerj.3254}{10.7717/peerj.3254}.

\leavevmode\hypertarget{ref-Sandiford2018}{}%
59. \textbf{Sandiford SK}. 2018. Current developments in lantibiotic
discovery for treating \emph{Clostridium difficile} infection. Expert
Opinion on Drug Discovery \textbf{14}:71--79.
doi:\href{https://doi.org/10.1080/17460441.2019.1549032}{10.1080/17460441.2019.1549032}.

\leavevmode\hypertarget{ref-Stein2013}{}%
60. \textbf{Stein RR}, \textbf{Bucci V}, \textbf{Toussaint NC},
\textbf{Buffie CG}, \textbf{Rätsch G}, \textbf{Pamer EG}, \textbf{Sander
C}, \textbf{Xavier JB}. 2013. Ecological modeling from time-series
inference: Insight into dynamics and stability of intestinal microbiota.
PLoS Computational Biology \textbf{9}:e1003388.
doi:\href{https://doi.org/10.1371/journal.pcbi.1003388}{10.1371/journal.pcbi.1003388}.

\leavevmode\hypertarget{ref-Nakashima2021}{}%
61. \textbf{Nakashima T}, \textbf{Fujii K}, \textbf{Seki T},
\textbf{Aoyama M}, \textbf{Azuma A}, \textbf{Kawasome H}. 2021. Novel
gut microbiota modulator, which markedly increases \emph{Akkermansia
muciniphila} occupancy, ameliorates experimental colitis in rats.
Digestive Diseases and Sciences.
doi:\href{https://doi.org/10.1007/s10620-021-07131-x}{10.1007/s10620-021-07131-x}.

\leavevmode\hypertarget{ref-Geerlings2018}{}%
62. \textbf{Geerlings S}, \textbf{Kostopoulos I}, \textbf{Vos W de},
\textbf{Belzer C}. 2018. \emph{Akkermansia muciniphila} in the human
gastrointestinal tract: When, where, and how? Microorganisms
\textbf{6}:75.
doi:\href{https://doi.org/10.3390/microorganisms6030075}{10.3390/microorganisms6030075}.

\leavevmode\hypertarget{ref-Deng2018}{}%
63. \textbf{Deng H}, \textbf{Yang S}, \textbf{Zhang Y}, \textbf{Qian K},
\textbf{Zhang Z}, \textbf{Liu Y}, \textbf{Wang Y}, \textbf{Bai Y},
\textbf{Fan H}, \textbf{Zhao X}, \textbf{Zhi F}. 2018. \emph{Bacteroides
fragilis} prevents \emph{Clostridium difficile} infection in a mouse
model by restoring gut barrier and microbiome regulation. Frontiers in
Microbiology \textbf{9}.
doi:\href{https://doi.org/10.3389/fmicb.2018.02976}{10.3389/fmicb.2018.02976}.

\leavevmode\hypertarget{ref-Engevik2020}{}%
64. \textbf{Engevik MA}, \textbf{Engevik AC}, \textbf{Engevik KA},
\textbf{Auchtung JM}, \textbf{Chang-Graham AL}, \textbf{Ruan W},
\textbf{Luna RA}, \textbf{Hyser JM}, \textbf{Spinler JK},
\textbf{Versalovic J}. 2020. Mucin-degrading microbes release
monosaccharides that chemoattract \emph{Clostridioides difficile} and
facilitate colonization of the human intestinal mucus layer. ACS
Infectious Diseases \textbf{7}:1126--1142.
doi:\href{https://doi.org/10.1021/acsinfecdis.0c00634}{10.1021/acsinfecdis.0c00634}.

\leavevmode\hypertarget{ref-Reeves2012}{}%
65. \textbf{Reeves AE}, \textbf{Koenigsknecht MJ}, \textbf{Bergin IL},
\textbf{Young VB}. 2012. Suppression of \emph{Clostridium difficile} in
the gastrointestinal tracts of germfree mice inoculated with a murine
isolate from the family \emph{Lachnospiraceae}. Infection and Immunity
\textbf{80}:3786--3794.
doi:\href{https://doi.org/10.1128/iai.00647-12}{10.1128/iai.00647-12}.

\leavevmode\hypertarget{ref-Ma2021}{}%
66. \textbf{Ma L}, \textbf{Keng J}, \textbf{Cheng M}, \textbf{Pan H},
\textbf{Feng B}, \textbf{Hu Y}, \textbf{Feng T}, \textbf{Yang F}. 2021.
Gut microbiome and serum metabolome alterations associated with isolated
dystonia. mSphere \textbf{6}.
doi:\href{https://doi.org/10.1128/msphere.00283-21}{10.1128/msphere.00283-21}.

\leavevmode\hypertarget{ref-Haas2020}{}%
67. \textbf{Haas KN}, \textbf{Blanchard JL}. 2020. Reclassification of
the \emph{Clostridium clostridioforme} and \emph{Clostridium sphenoides}
clades as \emph{Enterocloster} gen. nov. And \emph{Lacrimispora} gen.
nov., Including reclassification of 15 taxa. International Journal of
Systematic and Evolutionary Microbiology \textbf{70}:23--34.
doi:\href{https://doi.org/10.1099/ijsem.0.003698}{10.1099/ijsem.0.003698}.

\leavevmode\hypertarget{ref-Finegold2005}{}%
68. \textbf{Finegold SM}, \textbf{Song Y}, \textbf{Liu C}, \textbf{Hecht
DW}, \textbf{Summanen P}, \textbf{Könönen E}, \textbf{Allen SD}. 2005.
\emph{Clostridium clostridioforme}: A mixture of three clinically
important species. European Journal of Clinical Microbiology \&
Infectious Diseases \textbf{24}:319--324.
doi:\href{https://doi.org/10.1007/s10096-005-1334-6}{10.1007/s10096-005-1334-6}.

\leavevmode\hypertarget{ref-VanInsberghe2020}{}%
69. \textbf{VanInsberghe D}, \textbf{Elsherbini JA}, \textbf{Varian B},
\textbf{Poutahidis T}, \textbf{Erdman S}, \textbf{Polz MF}. 2020.
Diarrhoeal events can trigger long-term \emph{Clostridium difficile}
colonization with recurrent blooms. Nature Microbiology
\textbf{5}:642--650.
doi:\href{https://doi.org/10.1038/s41564-020-0668-2}{10.1038/s41564-020-0668-2}.

\leavevmode\hypertarget{ref-GarzaGonzalez2019}{}%
70. \textbf{Garza-González E}, \textbf{Mendoza-Olazarán S},
\textbf{Morfin-Otero R}, \textbf{Ramírez-Fontes A},
\textbf{Rodríguez-Zulueta P}, \textbf{Flores-Treviño S},
\textbf{Bocanegra-Ibarias P}, \textbf{Maldonado-Garza H},
\textbf{Camacho-Ortiz A}. 2019. Intestinal microbiome changes in fecal
microbiota transplant (FMT) vs. FMT enriched with \emph{Lactobacillus}
in the treatment of recurrent \emph{Clostridioides difficile} infection.
Canadian Journal of Gastroenterology and Hepatology \textbf{2019}:1--7.
doi:\href{https://doi.org/10.1155/2019/4549298}{10.1155/2019/4549298}.

\leavevmode\hypertarget{ref-Shafiq2020}{}%
71. \textbf{Shafiq M}, \textbf{Alturkmani H}, \textbf{Zafar Y},
\textbf{Mittal V}, \textbf{Lodhi H}, \textbf{Ullah W}, \textbf{Brewer
J}. 2020. Effects of co-infection on the clinical outcomes of
\emph{Clostridium difficile} infection. Gut Pathogens \textbf{12}.
doi:\href{https://doi.org/10.1186/s13099-020-00348-7}{10.1186/s13099-020-00348-7}.

\leavevmode\hypertarget{ref-Keith2020}{}%
72. \textbf{Keith JW}, \textbf{Dong Q}, \textbf{Sorbara MT},
\textbf{Becattini S}, \textbf{Sia JK}, \textbf{Gjonbalaj M},
\textbf{Seok R}, \textbf{Leiner IM}, \textbf{Littmann ER}, \textbf{Pamer
EG}. 2020. Impact of antibiotic-resistant bacteria on immune activation
and \emph{Clostridioides difficile} infection in the mouse intestine.
Infection and Immunity \textbf{88}.
doi:\href{https://doi.org/10.1128/iai.00362-19}{10.1128/iai.00362-19}.

\leavevmode\hypertarget{ref-Zackular2016}{}%
73. \textbf{Zackular JP}, \textbf{Moore JL}, \textbf{Jordan AT},
\textbf{Juttukonda LJ}, \textbf{Noto MJ}, \textbf{Nicholson MR},
\textbf{Crews JD}, \textbf{Semler MW}, \textbf{Zhang Y}, \textbf{Ware
LB}, \textbf{Washington MK}, \textbf{Chazin WJ}, \textbf{Caprioli RM},
\textbf{Skaar EP}. 2016. Dietary zinc alters the microbiota and
decreases resistance to \emph{Clostridium difficile} infection. Nature
Medicine \textbf{22}:1330--1334.
doi:\href{https://doi.org/10.1038/nm.4174}{10.1038/nm.4174}.

\leavevmode\hypertarget{ref-Berkell2021}{}%
74. \textbf{Berkell M}, \textbf{Mysara M}, \textbf{Xavier BB},
\textbf{Werkhoven CH van}, \textbf{Monsieurs P}, \textbf{Lammens C},
\textbf{Ducher A}, \textbf{Vehreschild MJGT}, \textbf{Goossens H},
\textbf{Gunzburg J de}, \textbf{Bonten MJM}, \textbf{Malhotra-Kumar S}.
2021. Microbiota-based markers predictive of development of
\emph{Clostridioides difficile} infection. Nature Communications
\textbf{12}.
doi:\href{https://doi.org/10.1038/s41467-021-22302-0}{10.1038/s41467-021-22302-0}.

\leavevmode\hypertarget{ref-Gardiner2014}{}%
75. \textbf{Gardiner BJ}, \textbf{Tai AY}, \textbf{Kotsanas D},
\textbf{Francis MJ}, \textbf{Roberts SA}, \textbf{Ballard SA},
\textbf{Junckerstorff RK}, \textbf{Korman TM}. 2014. Clinical and
microbiological characteristics of \emph{Eggerthella lenta} bacteremia.
Journal of Clinical Microbiology \textbf{53}:626--635.
doi:\href{https://doi.org/10.1128/jcm.02926-14}{10.1128/jcm.02926-14}.

\leavevmode\hypertarget{ref-Iljazovic2020}{}%
76. \textbf{Iljazovic A}, \textbf{Roy U}, \textbf{Gálvez EJC},
\textbf{Lesker TR}, \textbf{Zhao B}, \textbf{Gronow A}, \textbf{Amend
L}, \textbf{Will SE}, \textbf{Hofmann JD}, \textbf{Pils MC},
\textbf{Schmidt-Hohagen K}, \textbf{Neumann-Schaal M}, \textbf{Strowig
T}. 2020. Perturbation of the gut microbiome by \emph{Prevotella} spp.
enhances host susceptibility to mucosal inflammation. Mucosal Immunology
\textbf{14}:113--124.
doi:\href{https://doi.org/10.1038/s41385-020-0296-4}{10.1038/s41385-020-0296-4}.

\leavevmode\hypertarget{ref-Nagalingam2013}{}%
77. \textbf{Nagalingam NA}, \textbf{Robinson CJ}, \textbf{Bergin IL},
\textbf{Eaton KA}, \textbf{Huffnagle GB}, \textbf{Young VB}. 2013. The
effects of intestinal microbial community structure on disease
manifestation in IL-\(10^{-/-}\) mice infected with \emph{Helicobacter
hepaticus}. Microbiome \textbf{1}.
doi:\href{https://doi.org/10.1186/2049-2618-1-15}{10.1186/2049-2618-1-15}.

\leavevmode\hypertarget{ref-AbernathyClose2021}{}%
78. \textbf{Abernathy-Close L}, \textbf{Barron MR}, \textbf{George JM},
\textbf{Dieterle MG}, \textbf{Vendrov KC}, \textbf{Bergin IL},
\textbf{Young VB}. 2021. Intestinal inflammation and altered gut
microbiota associated with inflammatory bowel disease render mice
susceptible to \emph{Clostridioides difficile} colonization and
infection. mBio.
doi:\href{https://doi.org/10.1128/mbio.02733-20}{10.1128/mbio.02733-20}.

\leavevmode\hypertarget{ref-Pirofski2008}{}%
79. \textbf{Pirofski L-a}, \textbf{Casadevall A}. 2008. The
damage-response framework of microbial pathogenesis and infectious
diseases, pp. 135--146. \emph{In} Advances in experimental medicine and
biology. Springer New York.

\leavevmode\hypertarget{ref-Casadevall2014}{}%
80. \textbf{Casadevall A}, \textbf{Pirofski L-a}. 2014. What is a host?
Incorporating the microbiota into the damage-response framework.
Infection and Immunity \textbf{83}:2--7.
doi:\href{https://doi.org/10.1128/iai.02627-14}{10.1128/iai.02627-14}.

\leavevmode\hypertarget{ref-Tailford2015}{}%
81. \textbf{Tailford LE}, \textbf{Crost EH}, \textbf{Kavanaugh D},
\textbf{Juge N}. 2015. Mucin glycan foraging in the human gut
microbiome. Frontiers in Genetics \textbf{6}.
doi:\href{https://doi.org/10.3389/fgene.2015.00081}{10.3389/fgene.2015.00081}.

\leavevmode\hypertarget{ref-Sorg2009}{}%
82. \textbf{Sorg JA}, \textbf{Dineen SS}. 2009. Laboratory maintenance
of \emph{Clostridium difficile}. Current Protocols in Microbiology
\textbf{12}.
doi:\href{https://doi.org/10.1002/9780471729259.mc09a01s12}{10.1002/9780471729259.mc09a01s12}.

\leavevmode\hypertarget{ref-Winston2016}{}%
83. \textbf{Winston JA}, \textbf{Thanissery R}, \textbf{Montgomery SA},
\textbf{Theriot CM}. 2016. Cefoperazone-treated mouse model of
clinically-relevant \emph{Clostridium difficile} strain R20291. Journal
of Visualized Experiments.
doi:\href{https://doi.org/10.3791/54850}{10.3791/54850}.

\leavevmode\hypertarget{ref-Kozich2013}{}%
84. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
85. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

\leavevmode\hypertarget{ref-Wang2007}{}%
86. \textbf{Wang Q}, \textbf{Garrity GM}, \textbf{Tiedje JM},
\textbf{Cole JR}. 2007. Naïve bayesian classifier for rapid assignment
of rRNA sequences into the new bacterial taxonomy. Applied and
Environmental Microbiology \textbf{73}:5261--5267.
doi:\href{https://doi.org/10.1128/aem.00062-07}{10.1128/aem.00062-07}.

\leavevmode\hypertarget{ref-Yue2005}{}%
87. \textbf{Yue JC}, \textbf{Clayton MK}. 2005. A similarity measure
based on species proportions. Communications in Statistics - Theory and
Methods \textbf{34}:2123--2131.
doi:\href{https://doi.org/10.1080/sta-200066418}{10.1080/sta-200066418}.

\leavevmode\hypertarget{ref-Segata2011}{}%
88. \textbf{Segata N}, \textbf{Izard J}, \textbf{Waldron L},
\textbf{Gevers D}, \textbf{Miropolsky L}, \textbf{Garrett WS},
\textbf{Huttenhower C}. 2011. Metagenomic biomarker discovery and
explanation. Genome Biology \textbf{12}:R60.
doi:\href{https://doi.org/10.1186/gb-2011-12-6-r60}{10.1186/gb-2011-12-6-r60}.

\leavevmode\hypertarget{ref-Benjamini1995}{}%
89. \textbf{Benjamini Y}, \textbf{Hochberg Y}. 1995. Controlling the
false discovery rate: A practical and powerful approach to multiple
testing. Journal of the Royal Statistical Society: Series B
(Methodological) \textbf{57}:289--300.
doi:\href{https://doi.org/10.1111/j.2517-6161.1995.tb02031.x}{10.1111/j.2517-6161.1995.tb02031.x}.

\leavevmode\hypertarget{ref-Topcuoglu2021}{}%
90. \textbf{Topçuoğlu B}, \textbf{Lapp Z}, \textbf{Sovacool K},
\textbf{Snitkin E}, \textbf{Wiens J}, \textbf{Schloss P}. 2021.
Mikropml: User-friendly R package for supervised machine learning
pipelines. Journal of Open Source Software \textbf{6}:3073.
doi:\href{https://doi.org/10.21105/joss.03073}{10.21105/joss.03073}.
\end{cslreferences}

\newpage

\includegraphics{../results/figures/figure_1.jpg}

\textbf{Figure 1. Human fecal microbial communities established diverse
gut bacterial communities in germ-free mice.} (A) Relative abundance of
10 most abundant bacterial classes of the gut bacterial communities on
timepoint day 0, 14 days post-colonization of germ-free C57Bl/6 mice
with human fecal samples. Each column of abundances represents an
individual mouse. Each group of mice that received the same donor are
grouped together and labeled with a letter (N for non-moribund mice and
M for moribund mice) and number. (B) Median (points) and interquartile
range (lines) of \(\beta\)-diversity (\(\theta\)\textsubscript{YC})
between an individual mouse and either all others which were inoculated
with feces from the same donor or from a different donor. The
\(\beta\)-diversity among the same donor comparison group was
significantly less than the \(\beta\)-diversity of the different donor
group (\emph{P} \textless{} 0.05, calculated by Wilcoxon rank sum test).

\hfill\break

\includegraphics{../results/figures/figure_2.jpg}

\textbf{Figure 2. All donor groups resulted in \emph{C. difficile}
infection but with different outcomes.} \emph{C. difficile} CFU per gram
of stool was measured the day after challenge with 10\(^{3}\) \emph{C.
difficile} RT027 human isolate 431 spores and at the end of the
experiment, 10 days post-challenge. Each point represents an individual
mouse. Mice are grouped by donor and labeled by the donor letter (N for
non-moribund mice and M for moribund mice) and number. Points are
colored by donor group. Mice from donor groups N1 through N6 succumbed
to the infection prior to day 10 and were not plated on day 10
post-challenge. LOD = Limit of detection.

\hfill\break

\includegraphics{../results/figures/figure_3.jpg}

\textbf{Figure 3. Histopathologic score and toxin activity varied across
donor groups.} (A) Fecal toxin activity was detected in some mice post
\emph{C. difficile} challenge in both moribund and non-moribund mice.
(B) Cecum scored for histopathologic damage from mice at the end of the
experiment. Samples were collected for histopathologic scoring on day 10
post-challenge for non-moribund mice or the day the mouse succumbed to
the infection for the moribund group (day 2-3 post-challenge). Each
point represents an individual mouse. Mice are grouped by donor and
labeled by the donor letter (N for non-moribund mice and M for moribund
mice) and number. Points are colored by donor group. Missing points are
from mice that had insufficient sample collected for assaying fecal
toxin.

\hfill\break

\includegraphics{../results/figures/figure_4.jpg}

\textbf{Figure 4. Individual fecal bacterial community members of the
murine gut associated with \emph{C .difficile} infection outcomes.} (A
and B) Relative abundance of genera at the time of \emph{C. difficile}
challenge (Day 0) that varied significantly by the moribundity and
histopathologic summary score or toxin activity by LEfSe analysis.
Median (points) and interquartile range (lines) are plotted. Genera are
ordered alphabetically to ease comparisons across analyses. (A) Relative
abundances were compared across infection outcome of moribund (colored
black) or non-moribund with either a high histopathologic score (score
greater than the median score of 5, colored green) or a low
histopathologic summary score (score less than the median score of 5,
colored light green). (B) Relative abundances were compared between mice
which toxin activity was detected (Toxigenic, colored dark purple) and
which no toxin activity was detected (Non-toxigenic, colored light
purple). (C) Endpoint bacterial OTUs correlated with histopathologic
summary score. Each individual mouse is plotted (transparent gray
point). Spearman's correlations were statistically significant after
Benjamini-Hochberg correction for multiple comparisons. * next to a name
indicates that the bacterial group was unclassified at lower
classification ranks.

\hfill\break

\includegraphics{../results/figures/figure_5.jpg}

\textbf{Figure 5. Fecal bacterial community members of the murine gut at
the time of \emph{C. difficile} (Day 0) predicted outcomes of the
infection.} Day 0 bacterial community members grouped by different
classification rank were modeled with random forest to predict the
infection outcome. The models used the highest taxonomic classification
rank that performed as well as the lower ranks. Median (solid points)
and interquartile range (lines) of the group relative abundance are
plotted. Bacterial groups are ordered by their importance to the model;
taxonomic group at the top of the plot had the greatest decrease in
performance when its relative abundances were permuted. * next to the
name indicates that the bacterial group was unclassified at lower
classification ranks. (A) Bacterial members grouped by phyla predicted
which mice would have toxin activity detected at any point throughout
the infection (Toxigenic, dark purple). (B) Bacterial members grouped by
class predicted which mice would become moribund (dark blue). (C)
Bacterial members grouped by genera predicted if the mice would have a
high (score greater than the median score of 5, colored dark green) or
low (score less than the median score of 5, colored light green)
histopathologic summary score.

\hfill\break

\includegraphics{../results/figures/figure_6.jpg}

\textbf{Figure 6. \emph{C. difficile} infection outcome were similar
within human fecal donor groups across different \emph{C. difficile}
isolates.} Mice colonized with human fecal donor communities N7, N6, or
M4 were challenged with the isolate from the previous experiments in
this study, \emph{C. difficile} RT027 human isolate 431 (blue points),
or a different \emph{C. difficile} isolate (purple points). \emph{C.
difficile} CFUs were enumerated at the end of the experiment (day 10
post-challenge) and are plotted grouped by donor source (top plot).
Histopathologic scores from the tissue collected at the endpoint (either
day 10 post-challenge or day mice succumbed to infection) are plotted
grouped by donor source (bottom plot). Each point represents an
individual mouse. Mice that had received fecal communities from human
donor M4 succumbed to the infection prior to day 10 post-challenge so
were not plated for CFU on day 10 post-challenge (DECD = deceased). RT =
ribotype, Isolate = \emph{C. difficile} human isolate id, LOD = limit of
detection.

\hfill\break

\includegraphics{../results/figures/figure_S1.jpg}

\textbf{Figure S1. Histopathologic score of tissue damage at the
endpoint of the infection.} Tissue collected at the endpoint, either day
10 post-challenge (Non-moribund) or day mice succumbed to infection
(Moribund), were scored from histopathologic damage. Each point
represents an individual mouse. Mice (points) are grouped and colored by
their human fecal community donor. Missing points are from mice that had
insufficient sample for histopathologic scoring.

\hfill\break

\includegraphics{../results/figures/figure_S2.jpg}

\textbf{Figure S2. Random forest models predicted outcomes of the
\emph{C. difficile} challenge.} (A-C) Taxonomic classification rank
model performance. Relative abundance at the time of \emph{C. difficile}
challenge (Day 0) of the bacterial community members grouped by
different classification rank were modeled with random forest to predict
the infection outcome. The models used the highest taxonomic
classification rank that performed as well as the lower ranks. Black
rectangle highlights classification rank used to model each outcome.
(D-F) Model feature importance. Bacterial groups are ordered by their
decrease in area under receiver-operator curve (AUC) when its relative
abundances was permuted. Individual relative abundances were added to F
since differences in AUC were outside the interquartile range. *
indicates bacterial group was unclassified at lower classification
ranks. For all plots, median (solid points) and interquartile range
(lines) are plotted. (A) Toxin production modeled which mice would have
toxin detected during the experiment. (B) Moribundity modeled which mice
would succumb to the infection prior to day 10 post-challenge. (C)
Histopathologic score modeled which mice would have a high (score
greater than the median score of 5) or low (score less than the median
score of 5) histopathologic summary score. (D) Bacterial phyla which
affected the performance of predicting toxin activity when permuted. (E)
Bacterial classes which affected the performance of predicting
moribundity when permuted. (D) Bacterial genera which affected the
performance of predicting histopathologic score when permuted.

\hfill\break

\includegraphics{../results/figures/figure_S3.jpg}

\textbf{Figure S3. Subset of donor communities were selected to test for
conserved \emph{C. difficile} infection outcome with a different
\emph{C. difficile} isolate.} (A) Relative abundance of genera in mice
from previous experiments which received human fecal donor communities
N7, N6, or M4. Each column represents an individual mouse. (B) Features
from the random forest model that predicted moribundity using relative
abundance of bacteria were grouped by their genus classification rank
and are ordered by their decrease in AUC when each feature was permuted.
Median (solid points) and interquartile range (lines) change in AUC are
plotted. (C) Relative abundance of the most important genera used by the
model to predict moribundity. Median (solid points) and interquartile
range (lines) of the group relative abundances are plotted. * next to a
name indicates that the bacterial group was unclassified at lower
classification ranks.

\end{document}