### 2022-01-26 exploratory/notebook/initial_analysis/initial_analysis_notebook.Rmd

---
title: "initial_analysis_notebook"
author: "Nick Lesniak"
date: "4/13/2021"
output: rmarkdown::github_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

### Process sequencing data

- removed unused fastq files

### run fastqs through mothur

```{r read in data}
library(tidyverse)
library(here)

# calculate error rate
error_df <- read.table(here('data/mothur/sample.error.count'), header = T)
total_seqs <- sum(error_df$Sequences)
total_errors <- sum(error_df$Errors * error_df$Sequences)
error_rate <- total_errors / (total_seqs * 253) * 100

read_tsv(here('data/mothur/sample.final.shared'),
         col_types = cols(.default = col_double(),
                          Group = col_character())) %>% 
  select(-numOtus, -label) %>% 
  pivot_longer(cols = -Group, values_to = 'count', names_to = 'otu') %>% 
  group_by(Group) %>% 
  summarise(counts = sum(count)) %>% 
  filter(counts < 2020) %>% 
  arrange(counts) %>% 
  mutate(list = paste0(Group, ' (', counts, ' seqs)')) %>% 
  pull(list) %>% 
  paste(., collapse = ', ')
```

subsampled to 2017, eliminating the 25 samples listed above with their counts
error rate for the data was `r round(error_rate, 2)`%

```{r}
# 
source(here('code/Sum_OTU_by_Tax.R'))

# load in data
metadata <- read_tsv(here('data/process/metadata_tidy.tsv'),
                     col_type = 'cccccDddddDcdddddcdl') 
toxin <- read_tsv(here('data/process/toxin_tidy.tsv'),
                  col_type = 'cccccdc')
histology <- read_tsv(here('data/process/histology_tidy.tsv'),
                      col_type = 'ccccdddcdcc')
human_source <- read_tsv(here('data/process/human_source_tidy.tsv'),
                         col_type = cols(.default = col_character(), age = col_double()))
shared <- read_tsv(here('data/mothur/sample.final.0.03.subsample.shared'),
         col_types = cols(.default = col_double(),
                          Group = col_character())) %>% 
  mutate(Group = ifelse(grepl('_D|DA', Group), Group,
      gsub('D', '_D', Group))) %>% 
  select(-label, -numOtus) %>% 
  pivot_longer(-Group, names_to = 'otu', values_to = 'counts') %>% 
  mutate(relative_abundance = counts/2107 * 100)
taxonomy <- read_tsv(here('data/process/final.taxonomy.tidy.tsv'),
                     col_types = cols(.default = col_character()))
alpha_df <- read_tsv(here('data/mothur/sample.final.groups.ave-std.summary'),
         col_types ='dccdddddd') %>% 
  filter(method == 'ave') %>% 
  mutate(group = ifelse(grepl('_D|DA', group), group,
                        gsub('D', '_D', group)))
nmds_df <- read_tsv(here('data/mothur/sample.final.thetayc.0.03.lt.ave.nmds.axes'),
         col_types ='cdd') %>% 
    mutate(group = ifelse(grepl('_D|DA', group), group,
                        gsub('D', '_D', group)))
source(here('code/read_dist.R'))
beta_df <- read_dist(here('data/mothur/sample.final.thetayc.0.03.lt.ave.dist')) %>% 
      mutate(rows = ifelse(grepl('_D|DA', rows), rows,
                        gsub('D', '_D', rows)),
             columns = ifelse(grepl('_D|DA', columns), columns,
                        gsub('D', '_D', columns)))
donor_colors <- c("#41bbc5", "#7e2640", "#9ad859", "#682dbd", "#728e24", 
                  "#e26df8", "#23980d", "#fd048f", "#46e33e", "#fa1bfc", 
                  "#155126", "#f9b2d1", "#154975", "#e2c627", "#6294ce")
```

# Figure 1 - Human communities reporducibly colonize mice

Sampled spread of diverse donors for inoculating germ-free mice

```{r fig.width=7, fig.height=5}

human_source_list <- shared %>% 
  filter(grepl('DA', Group)) %>% 
  pull(Group) %>% unique

#nmds_df %>% 
#  filter(group %in% human_source_list) %>% 
#  ggplot(aes(x = axis1, y = axis2, color = outcome)) + 
#    geom_point() + 
#    labs(x = 'NMDS axis 1', y = 'NMDS axis 2', color = NULL) + 
#    theme_bw()
# theta yc shows comparison better
```

Unique communities without conserved structure
```{r, fig.height= 6, fig.width=6}
donor_plot <- shared %>% 
  filter(counts > 0,
         grepl('DA', Group)) %>% 
  left_join(taxonomy, by = c('otu' = 'OTU')) %>% 
  rename(taxa_level = Class) %>% 
  group_by(Group, taxa_level) %>% 
  summarise(relative_abundance = log10(sum(relative_abundance) + 0.0001)) %>% 
  group_by(taxa_level) %>% 
  mutate(median_ra = median(relative_abundance)) %>% 
  ggplot(aes(x = Group, 
             y = reorder(taxa_level, median_ra), 
             fill = relative_abundance)) + 
    geom_tile() + 
    scale_fill_gradient2(low="white", mid='#0B775E', high = 'black',
		  	limits = c(-2,2), na.value = NA, midpoint = .3,
			  breaks = c(-2.5, -1, 0, 1, 2), labels = c('', '0.1', '1', '10', '100')) + 
	  theme_bw() + 
    labs(x = NULL, y = NULL, fill = NULL) + 
    scale_x_discrete(guide = guide_axis(angle = 45)) + 
    facet_grid(.~Group, scales = 'free_x') + 
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          legend.position = 'none')

day_0_plot <- metadata %>% 
  filter(day == 0,
         cdiff_strain == 431) %>% 
  select(group, human_source, mouse_id) %>% 
  left_join(filter(shared, counts > 0), by = c('group' = 'Group')) %>% 
  left_join(taxonomy, by = c('otu' = 'OTU')) %>% 
  rename(taxa_level = Class) %>% 
  group_by(group, taxa_level, human_source) %>% 
  summarise(relative_abundance = log10(sum(relative_abundance) + 0.0001)) %>% 
  group_by(taxa_level) %>% 
  filter(!is.na(taxa_level)) %>% 
  mutate(median_ra = mean(relative_abundance),
         human_source = as.character(as.numeric(gsub('DA', '', human_source))),
         human_source = factor(human_source, levels =
                                 c('10093', '10148', '431', '884', '1245',
                                   '581', '1324', '953', '1134',
                                   '369', '430', '10027', '10034',
                                   '10082', '578')),
         taxa_level = gsub('_unclassified', '', taxa_level),
         taxa_level = gsub('_', ' ', taxa_level),
         taxa_level = paste0('*', taxa_level, '*')) %>% 
  ggplot(aes(x = group, 
             y = reorder(taxa_level, median_ra), 
             fill = relative_abundance)) + 
    geom_tile() + 
    scale_fill_gradient2(low="white", mid='#0B775E', high = 'black',
		  	limits = c(-2,2), na.value = NA, midpoint = .3,
			  breaks = c(-2.5, -1, 0, 1, 2), labels = c('', '0.1', '1', '10', '100')) + 
	  theme_bw() + 
    labs(x = NULL, y = NULL, fill = NULL) + 
    scale_x_discrete(guide = guide_axis(angle = 45)) + 
    facet_grid(.~human_source, scales = 'free_x') + 
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          legend.position = 'bottom',
          axis.text.y = ggtext::element_markdown(angle = 45),
          strip.text.x = element_text(angle = 45),
          strip.background = element_blank(),
          legend.margin=margin(t=-0.1, r=0, b=-0.1, l=0, unit="cm"),
          legend.key.height = unit(0.3, 'cm'))

#cowplot::plot_grid(
#  cowplot::plot_grid(NULL, donor_plot, nrow = 1, rel_widths = c(.95,13)),
#  day_0_plot, ncol = 1, rel_heights = c(8,9))
day_0_plot

```

Unsure about alpha plots

```{r}
alpha_df <- alpha_df %>% 
  select(group, sobs, invsimpson) %>% 
  left_join(bind_rows(metadata, 
                      tibble(group = unique(metadata$human_source),
                             human_source = unique(metadata$human_source),
                             sample_type = 'donor')),
            by = c('group')) %>% 
  mutate(sample = case_when(sample_type == 'donor' ~ 'Donor',
                            day %in% c(0:10) ~ paste('Day', day),
                            day == mouse_endpoint ~ 'Endpoint',
                            T ~ 'NA'),
         sample = factor(sample, levels = c('Donor', 'Day 0', 'Day 1', 'Day 2', 'Day 3', 
                                            'Day 4', 'Day 5', 'Day 6', 'Day 7', 'Day 8', 'Day 9', 'Day 10'))) %>% 
  filter(sample != 'NA')

alpha_df %>% 
  filter(day == 0) %>% 
  #kruskal.test(x = .$invsimpson, g = as.factor(.$human_source)) # p = 0.002
  ggplot(aes(x = human_source, y = invsimpson, color = human_source)) + 
    geom_point(alpha = 0.3) + 
    theme_bw() + 
    #facet_wrap(human_source~.) + 
    guides(color = 'none') +  
    ylim(0,NA)

alpha_df %>% 
  filter(day == 0) %>% 
  #kruskal.test(x = .$sobs, g = as.factor(.$human_source)) # p = 9.9e-5
  ggplot(aes(x = human_source, y = sobs, color = human_source)) + 
    geom_point(alpha = 0.3) + 
    theme_bw() + 
    #facet_wrap(human_source~.) + 
    guides(color = 'none') +
    ylim(0,NA)

# how many otus are unique to each donor and what portion of community?
otu_dist <- alpha_df %>% 
  filter(day == 0) %>% 
  select(group, human_source) %>% 
  left_join(shared, by = c('group' = 'Group')) %>% 
  filter(relative_abundance > 0.1) %>% 
  select(human_source, otu) 
# 353 otus
#otu_dist %>% 
#  pull(otu) %>% 
#  unique %>% length 
otu_dist %>% 
  distinct %>% 
  group_by(otu) %>% 
  mutate(otu_present_n_samples = length(human_source)) %>% 
  left_join(select(taxonomy, tax_otu_label, OTU), by = c('otu' = 'OTU')) %>% 
  left_join(shared, by = c('otu')) %>% 
  ggplot(aes(x = as.factor(otu_present_n_samples), y = relative_abundance)) + 
    geom_violin() +
    scale_y_log10() + 
    facet_wrap(human_source~.)

```

```{r nmds, eval = F, include = F}
initial_nmds_plot <- nmds_df %>% 
  left_join(filter(select(metadata, group, human_source, day), day %in% c(0, 10)), by = c('group')) %>% 
  mutate(human_source = case_when(!is.na(human_source) ~ human_source,
                                  grepl('DA', group) ~ group,
                                  grepl('CON3', group) ~ 'DA00430',
                                  T ~ 'Unknown'),
         sample_type = ifelse(grepl('DA', group), 'human', 'murine')) %>% 
  filter(sample_type != 'human', day == 0) %>% 
  ggplot(aes(x = axis1, y = axis2, color = human_source)) +
    geom_point() + 
    scale_color_manual(values = donor_colors) + 
    #facet_grid(~sample_type) + 
    theme_bw() + 
    guides(color = 'none') + 
    labs(x = 'NMDS axis 1', y = 'NMDS axis 2')

# plot for presentation
#ggsave('~/Desktop/intial_nmds.jpg',
#       initial_nmds_plot,
#       height = 4, width = 4)

initial_nmds_plot

```

```{r}
day_0_samples <- metadata %>% 
  filter(day == 0) %>% 
  select(group, human_source)
day_10_samples <- metadata %>% 
  filter(day == 10) %>% 
  select(group, human_source)

# compare distance between:
#   donor and recipient
#   co-recipients
#   other cages
#   other donors

donor_v_donor <- beta_df %>% 
  filter(grepl('DA', rows),
         grepl('DA', columns)) %>% 
  mutate(comparison = 'Donor vs Donor')

donor_v_recipent <- bind_rows(
    day_0_samples %>% 
      inner_join(beta_df, by = c('group' = 'rows', 'human_source' = 'columns')),
    day_0_samples %>% 
      inner_join(beta_df, by = c('group' = 'columns', 'human_source' = 'rows'))) %>% 
  mutate(comparison = 'Donor vs Recipient Initial')

donor_v_recipent_end <- bind_rows(
    day_10_samples %>% 
      inner_join(beta_df, by = c('group' = 'rows', 'human_source' = 'columns')),
    day_10_samples %>% 
      inner_join(beta_df, by = c('group' = 'columns', 'human_source' = 'rows'))) %>% 
  mutate(comparison = 'Donor vs Recipient End')

corecipient <- day_0_samples %>% 
  left_join(day_0_samples, by = c('human_source')) %>% 
  filter(group.x != group.y) %>% 
  inner_join(beta_df, by = c('group.x' = 'rows', 'group.y' = 'columns')) %>% 
  mutate(comparison = 'Recipients compared to\nothers w/same donor')

other_donors <- beta_df %>% 
  filter((grepl('DA', rows) & grepl('D0', columns)) | 
         (grepl('D0', rows) & grepl('DA', columns))) %>% 
  anti_join(donor_v_recipent, by = c('rows' = 'group', 'columns' = 'human_source')) %>% 
  anti_join(donor_v_recipent, by = c('rows' = 'human_source', 'columns' = 'group')) %>% 
  mutate(comparison = 'Donor vs Non-recipients')

other_recipients <- beta_df %>% 
  filter(grepl('D0', rows) & grepl('D0', columns)) %>% 
  anti_join(corecipient, by = c('rows' = 'group.x', 'columns' = 'group.y')) %>% 
  mutate(comparison = 'Recipients compared to\nothers w/different donor')

day0_v_day10 <- beta_df %>% 
  separate(rows, c('cageA', 'mouseA', 'dayA')) %>% 
  separate(columns, c('cageB', 'mouseB', 'dayB')) %>% 
  filter(mouseA == mouseB & cageA == cageB) %>% 
  filter((dayA == 'D0' & dayB == 'D10') |
         (dayA == 'D10' & dayB == 'D0')) %>% 
  mutate(comparison = 'Initial vs End')

beta_plot <- bind_rows(donor_v_donor, donor_v_recipent, corecipient, other_donors, 
          other_recipients, day0_v_day10, donor_v_recipent_end) %>% 
  mutate(comparison = factor(comparison, levels = 
                              c('Donor vs Donor',
                                'Donor vs Non-recipients', 
                                'Recipients compared to\nothers w/different donor',
                                'Recipients compared to\nothers w/same donor', 
                                'Donor vs Recipient End',
                                'Initial vs End',
                                'Donor vs Recipient Initial'))) %>% 
  ggplot(aes(x = comparison, y = distances)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    geom_jitter(alpha = 0.1, width = 0.3) + 
    labs(x = NULL, y = 'Theta YC') +
    theme_bw() + 
    coord_flip()

refined_beta_plot <- bind_rows(corecipient, other_recipients) %>% 
  mutate(comparison = ifelse(comparison == 'Recipients compared to\nothers w/different donor',
                             'Different donor', 'Same donor'),
    comparison = factor(comparison, levels = 
                              c('Same donor',
                                'Different donor'))) %>% 
  ggplot(aes(x = comparison, y = distances)) + 
    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
    #geom_jitter(alpha = 0.1, width = 0.3) + 
    labs(x = NULL, y = 'Similarity') +
    scale_y_continuous(breaks = c(0, 1), limits = c(0,1),
                       labels = c('Complete\nsimilarity', 'No similarity')) + 
    theme_bw()
#wilcox.test(corecipient$distances, other_recipients$distances)
# presentation plot
#ggsave('~/Desktop/beta_plot.jpg',
#       refined_beta_plot,
#       width = 3, height = 5)

# how do donors distances look by donor?
#donor_v_donor %>% 
#  pivot_longer(cols = c(rows, columns)) %>% 
#  ggplot(aes(x = value, y = distances, color = value)) + geom_point()
#    geom_jitter(width = 0.25) + 
#    scale_color_manual(values = donor_colors) %>% 
#    theme_bw()

corecipient %>% 
  separate(group.x, c('cageA', 'mouseA', 'dayA')) %>% 
  separate(group.y, c('cageB', 'mouseB', 'dayB')) %>% 
  mutate(same_cage = cageA == cageB) %>% 
  ggplot(aes(x = human_source, y = distances, color = human_source, shape = same_cage)) +
    geom_point(position = position_jitterdodge(), width = 0.2) + 
    scale_color_manual(values = donor_colors) + 
    scale_shape_manual(values = c(18,16)) + 
    scale_x_discrete(guide = guide_axis(angle = 45)) + 
    theme_bw() +
    theme(legend.position = 'bottom') + 
    labs(y = 'Theta YC', x = 'Donor Source', color = NULL, shape = 'Same Cage') + 
    guides(color = 'none')

ggsave(here('results/figures/Figure_1.jpg'),
  cowplot::plot_grid(day_0_plot + theme(text=element_text(size = 9)), 
                     NULL,
                     beta_plot + theme(text=element_text(size = 9)), 
                     ncol = 1, 
                     rel_heights = c(6, 0.3, 3),
                     labels = c('A', 'B', NULL)),
  height = 9.0625, width = 6.875, unit = 'in')
```

# Figure 2 - Mice were colonized without any perturbation

```{r, include = T, eval = F}
metadata %>% 
  filter(cdiff_strain == 431) %>% 
  mutate(human_source = factor(human_source, levels = 
                                c('DA01324', 'DA00953', 'DA00581', 'DA01134', 'DA10148', 
                                  'DA00369', 'DA00430', 'DA10093', 'DA10027', 'DA10034',
                                  'DA10082', 'DA00884', 'DA00578', 'DA00431', 'DA01245'))) %>% 
  ggplot(aes(x = day, y = log10(cdiff_cfu + 60), color = human_source)) +
    stat_summary(fun = 'median', geom = 'line') + 
    geom_point(alpha = 0.3) +
    scale_color_manual(values = donor_colors) + 
    guides(color = 'none') + 
    theme_bw() + 
    labs(x = 'Day', y = 'C. difficile CFU (log10)')
```
```{r}
cfu_plot <- metadata %>% 
  filter(cdiff_strain == 431,
         day %in% c(1,10)) %>% 
  mutate(human_source = as.numeric(gsub('DA', '', human_source)),
         human_source = factor(human_source, levels = 
                                c('1324', '953', '581', '1134', '10148', 
                                  '369', '430', '10093', '10027', '10034',
                                  '10082', '884', '578', '431', '1245')),
         day = factor(paste('Day', day), levels = c('Day 1', 'Day 10'))) %>% 
  ggplot(aes(x = human_source, y = log10(cdiff_cfu + 60), color = human_source)) +
    geom_jitter(width = 0.2) + 
    facet_grid(day~.)+ 
    scale_x_discrete(guide = guide_axis(angle = 45)) + 
    scale_color_manual(values = donor_colors) + 
    theme_bw() + 
    labs(x = 'Donor Source', y = 'C. difficile CFU (log10)') + 
    guides(color = 'none')

ggsave(here('results/figures/Figure_2.jpg'),
       cfu_plot,
       width = 6, height = 4)

# presentation plot
#presentation_cfu_plot <- metadata %>% 
#  filter(cdiff_strain == 431,
#         day %in% c(1,10)) %>% 
#  mutate(human_source = as.numeric(gsub('DA', '', human_source)),
#         human_source = factor(human_source, levels = 
#                                c('1324', '953', '581', '1134', '10148', 
#                                  '369', '430', '10093', '10027', '10034',
#                                  '10082', '884', '578', '431', '1245'),
#                               labels = LETTERS[1:15]),
#         day = factor(paste('Day', day), levels = c('Day 1', 'Day 10'))) %>% 
#  ggplot(aes(x = human_source, y = cdiff_cfu, color = human_source)) +
#    geom_jitter(width = 0.2) + 
#    facet_grid(day~.) + 
#    scale_color_manual(values = donor_colors) + 
#    scale_y_continuous(limits = c(1.5,8.5),
#                       breaks = c(2,4,6,8),
#                       labels = c('10^2', '10^4', '10^6', '10^8')) + 
#    theme_bw() + 
#    labs(x = 'Donor Source', y = '*C. difficile* CFU') + 
#    guides(color = 'none') + 
#    theme(axis.text.y = ggtext::element_markdown(),
#          axis.title.y = ggtext::element_markdown(),
#          strip.text.y = element_text(angle = 0, size = 12),
#          strip.background = element_blank(),
#          panel.spacing = unit(2, "lines")) + 
#			geom_hline(yintercept = 2, linetype = 'dashed', size = .5, color = 'white') + 
#			geom_label(x = 14, y = 2, label = "LOD", fill = 'white', color = 'white') + 
#			geom_text(x = 14, y = 2, label = "LOD", color = 'lightgray', size = 12/.pt)
#ggsave('~/Desktop/cfu_plot.jpg',
#       presentation_cfu_plot,
#       width = 8, height = 5)
#
```

DA00581 is made up of 4 cages, and at day 1 we see 1 cage at ~10^7.5, 2 cages just above 10^5 and 1 cage at 0.
For the cage at 0 there are three mice, one remains at 0, the other two have transient low appearance of cdiff  as seen below

```{r}
metadata %>% filter(human_source == 'DA00581') %>% 
  ggplot(aes(x = day, y = log10(cdiff_cfu + 60), shape = mouse_id, color = cage_id)) +
    geom_line() + 
    geom_point() + 
    scale_shape_manual(values = 1:11) + 
    theme_bw() + 
    guides(shape = 'none') + 
    labs(x = 'Day', y = 'C. difficile CFU (log10)')

```

Points are added to allow identification of individuals. Here we see one mouse has low cfu on days 6 and 7, another mouse on day 10.

# Figure supplemental - weight loss

```{r, include = F, eval = F}
metadata %>% 
  filter(cdiff_strain == 431) %>% 
  mutate(human_source = factor(human_source, levels = 
                                c('DA01324', 'DA00953', 'DA00581', 'DA01134', 'DA10148', 
                                  'DA00369', 'DA00430', 'DA10093', 'DA10027', 'DA10034',
                                  'DA10082', 'DA00884', 'DA00578', 'DA00431', 'DA01245'))) %>% 
  ggplot(aes(x = day, y = percent_weightLoss, color = human_source)) +
    stat_summary(fun = 'median', geom = 'line') + 
    geom_point(alpha = 0.3) +
    scale_color_manual(values = donor_colors) + 
    guides(color = 'none') + 
    theme_bw() + 
    labs(x = 'Day', y = 'Weight (Relative to initial)')
```

```{r}
wt_loss_plot <- metadata %>% 
  filter(cdiff_strain == 431,
         day %in% c(2,10)) %>% 
  mutate(human_source = as.numeric(gsub('DA', '', human_source)),
         human_source = factor(human_source, levels = 
                                c('1324', '953', '581', '1134', '10148', 
                                  '369', '430', '10093', '10027', '10034',
                                  '10082', '884', '578', '431', '1245')),
         day = factor(paste('Day', day), levels = c('Day 2', 'Day 10'))) %>% 
  ggplot(aes(x = human_source, y = percent_weightLoss, color = human_source)) +
    geom_jitter(width = 0.2) + 
    facet_grid(day~.)+ 
    scale_x_discrete(guide = guide_axis(angle = 45)) + 
    scale_color_manual(values = donor_colors) + 
    theme_bw() + 
    labs(x = 'Donor Source', y = 'Weight (% relative to initial)') + 
    guides(color = 'none')

wt_loss_plot

ggsave(here('results/figures/Figure_S1.jpg'),
       wt_loss_plot,
       width = 6, height = 4)
```

```{r}
#metadata %>% 
#  filter(day == 0 | day == mouse_endpoint) %>% 
#  mutate(time_point = ifelse(day == 0, 'Initial', 'Endpoint'),
#         time_point = factor(time_point, levels = c('Initial', 'Endpoint'))) %>% 
#  left_join(nmds_df, by = c('group')) %>% 
#  ggplot(aes(x = axis1, y = axis2, color = early_euth)) +
#    geom_point() + 
#    facet_grid(.~time_point) +
#    #scale_color_manual(values  = donor_colors) + 
#    theme_bw() + 
#    labs(x = 'NMDS axis 1', y = 'NMDS axis 2', color = 'Moribund') + 
#    theme(legend.position = 'bottom')
#
# not sure if this adds to this figure

```

> NAs on Day 0 for DA01134, DA10034, DA00884  
> Last samples for moribund mice is from cecum

# Figure 3 - Difference in severity

 What is toxin distribution by outcome?

```{r histology}
#mouse_by_outcome <- histology %>% 
#  mutate(mouse_id = paste0(cage_id, '_', ear_tag)) %>% 
#  select(mouse_id, human_source, epithelial_damage, summary_score) %>% 
#  left_join(unique(select(metadata, mouse_id, early_euth, cdiff_strain),
#            by = c('mouse_id'))) %>% 
#  filter(cdiff_strain == 431) %>% 
#  group_by(human_source) %>% 
#  mutate(severity = mean(epithelial_damage) < 1) %>% 
## using epithelial damage
#  mutate(outcome = case_when(early_euth == T ~ 'moribund',
#                             severity == T ~ 'mild',
#                             severity == F ~ 'severe',
#                             T ~ 'NA')) %>% 
## using summary score
##    mutate(outcome_ss = case_when(early_euth == T ~ 'moribund',
##                             summary_score < 4.4 ~ 'mild', # mean = 4.4, median = 5
##                             summary_score > 4.4 ~ 'severe',
##                             T ~ 'NA')) %>% 
#  mutate(outcome = factor(outcome, levels = c('mild', 'severe', 'moribund'))) %>% 
#  ungroup %>% 
#  select(mouse_id, outcome)
# if not using histology to split non-severe into mild/moderate, no need for mouse_by_outcome

histology_plot <- histology %>% 
  left_join(distinct(select(metadata, mouse_id, early_euth)), by = 'mouse_id') %>% 
  filter(!is.na(early_euth)) %>% 
  mutate(early_euth = ifelse(early_euth, 'Severe', 'Mild'),
         mouse_id = paste0(cage_id, '_', ear_tag)) %>% 
  #left_join(mouse_by_outcome, by = c('mouse_id')) %>%
  pivot_longer(cols = c(edema_tissue, inflammation_tissue, epithelial_damage, summary_score),
               names_to = 'histology', values_to = 'score') %>% 
  filter(histology == 'summary_score') %>% 
  mutate(histology = ifelse(histology == 'summary_score', 'End point', NA),
         human_source = gsub('DA0{0,2}', '', human_source)) %>% 
  ggplot(aes(x = human_source, y = score, , color = early_euth, fill = early_euth)) +
    #stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5), position = position_dodge(width = 0.4)) +
    geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 2) + 
    #geom_point(position = position_jitterdodge(jitter.width = 0.3), alpha = 0.5) + 
    theme_bw() + 
    labs(x = NULL, y = 'Clinical\nScore', fill = 'Outcome') +
    scale_x_discrete(guide = guide_axis(angle = 45)) + 
    scale_y_continuous(breaks = c(1,3,5,7,9), 
                       labels = c(1,3,5,7,9)) + 
    facet_grid(histology~early_euth, scales = 'free', space = 'free_x') + 
    guides(color = 'none') + 
    theme(legend.position = 'bottom',
          strip.background.x = element_blank(),
          strip.text.x = element_blank(),
          legend.margin=margin(t=-0.3, r=0, b=-0.2, l=0, unit="cm"),
          legend.key.height = unit(0, 'cm'))

#if trying to split colonized
#> **Splitting by mean epithelial damage for donor < 1 = 12 mild, 15 severe, 17 moribund**  
#> Splitting of median summary score of 5 =  10 mild, 11 severe, 17 moribund  
#>        (6 have a summary score of 5)  
#> Splitting of mean summary score of 4.4 =  10 mild, 17 severe, 17 moribund

toxin_plot <- toxin %>% 
  inner_join(select(metadata, group, human_source, early_euth), by = c('group')) %>% 
#  left_join(mouse_by_outcome, by = c('mouse_id')) %>% 
  mutate(outcome = ifelse(early_euth == T, 'Severe', 'Mild')) %>% 
  filter(!is.na(outcome)) %>%
  mutate(day = as.numeric(toxin_sample_day),
         human_source = gsub('DA0{0,2}', '', human_source)) %>% 
  filter(!day %in% c(4,7),
         toxin_sample_type == 'stool') %>% 
  ggplot(aes(x = human_source, y = Log_repiricoal_dilution, 
             fill = outcome, color = outcome)) + 
    #stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5), position = position_dodge(width = 0.4)) +
    #geom_jitter(width= 0.3, height = 0, alpha = 0.3) + 
    geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 2) + 
    labs(x = NULL, y = 'Toxin activity\n ') + 
    facet_grid(day~outcome, scales = 'free_x', space = 'free_x') + 
    theme_bw() +
    theme(legend.position = 'none',
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.grid.minor.y = element_blank())

toxin_plot
histology_plot

ggsave(here('results/figures/Figure_3.jpg'),
  cowplot::plot_grid(toxin_plot, NULL, histology_plot, ncol = 1, 
                   rel_heights = c(8, 1, 4),
                   labels = c('A', 'B', NULL)),
  width = 6, height = 4)

# presentation plot
#presentation_histology_plot <- histology %>% 
#  left_join(distinct(select(metadata, mouse_id, early_euth)), by = 'mouse_id') %>% 
#  filter(!is.na(early_euth)) %>% 
#  mutate(early_euth = ifelse(early_euth, 'Severe', 'Non-severe'),
#         mouse_id = paste0(cage_id, '_', ear_tag)) %>% 
#  #left_join(mouse_by_outcome, by = c('mouse_id')) %>%
#  pivot_longer(cols = c(edema_tissue, inflammation_tissue, epithelial_damage, summary_score),
#               names_to = 'histology', values_to = 'score') %>% 
#  filter(histology == 'summary_score') %>% 
#  mutate(histology = ifelse(histology == 'summary_score', 'End point', NA),
#         human_source = gsub('DA0{0,2}', '', human_source),
#         human_source = factor(human_source, levels = 
#                                 c('1324', '953', '581', '1134', '10148', 
#                                   '369', '430', '10093', '10027', '10034',
#                                   '10082', '884', '578', '431', '1245'),
#                               labels = LETTERS[1:15])) %>% 
#  ggplot(aes(x = human_source, y = score, fill = human_source, color = human_source)) +
#    geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 1) + 
#    theme_bw() + 
#    labs(x = NULL, y = 'Clinical Score') +
#    scale_y_continuous(breaks = c(1,3,5,7,9), 
#                       labels = c(1,3,5,7,9)) + 
#    scale_fill_manual(values = donor_colors) + 
#    scale_color_manual(values = donor_colors) + 
#    facet_grid(.~early_euth, scales = 'free', space = 'free_x') + 
#    theme(legend.position = 'none',
#          strip.background.x = element_blank(),
#          strip.text.x = element_text(size = 12),
#          legend.margin=margin(t=-0.3, r=0, b=-0.2, l=0, unit="cm"),
#          legend.key.height = unit(0, 'cm'))
#
#ggsave('~/Desktop/histology_plot.jpg',
#       presentation_histology_plot,
#       width = 8, height = 3.5)
#
```

```{r, toxin/hist pvalues}

toxin %>% 
  filter(toxin_sample_day < 3,
         toxin_sample_type == 'stool') %>% 
  left_join(
    unique(select(metadata, mouse_id, early_euth)),
    by = c('mouse_id')) %>% 
  select(early_euth, Log_repiricoal_dilution) %>% 
  filter(!is.na(early_euth)) %>% 
  summarise(pvalue = 
    wilcox.test(
      pull(filter(., early_euth ==T), Log_repiricoal_dilution),
      pull(filter(., early_euth ==F), Log_repiricoal_dilution),
      alternative = 'greater')$p.value)

histology %>% 
  left_join(distinct(select(metadata, mouse_id, early_euth)), by = 'mouse_id') %>% 
  filter(!is.na(early_euth)) %>% 
  summarise(pvalue = 
    wilcox.test(
      pull(filter(., early_euth ==T), summary_score),
      pull(filter(., early_euth ==F), summary_score),
      alternative = 'greater')$p.value)

```

```{r}
## Can we use toxin to split severity?
#do mice split into +/- toxin based on outcome of moribund/mild/severe?

#toxin %>% 
#  mutate(mouse_id = ifelse(toxin_sample_type == 'cecalcon', 
#                           gsub('C$', '', mouse_id), mouse_id)) %>% 
#  group_by(mouse_id) %>% 
#  summarise(max_toxin = max(Log_repiricoal_dilution),
#            mean_toxin = mean(Log_repiricoal_dilution)) %>% 
#  left_join(mouse_by_outcome, by = c('mouse_id')) %>% 
#  filter(!is.na(human_source)) %>% 
#  pivot_longer(c(max_toxin, mean_toxin)) %>% 
#  ggplot(aes(x = outcome, y = value)) + 
#    stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
#    geom_jitter(alpha = 0.3, width = 0.3) + 
#    facet_wrap(.~name, scales = 'free_y') + 
#    labs(x = NULL, y = 'Log reciprocal toxin') + 
#    theme_bw()
```

```{r}
## Does toxin explain histological findings?
#toxin %>% 
#  mutate(mouse_id = ifelse(toxin_sample_type == 'cecalcon', 
#                           gsub('C$', '', mouse_id), mouse_id)) %>% 
#  group_by(mouse_id) %>% 
#  mutate(last_sample = max(toxin_sample_day)) %>% 
#  filter(toxin_sample_day == last_sample) %>% 
#  left_join(select(histology, mouse_id, summary_score), by = c('mouse_id')) %>% 
#  left_join(mouse_by_outcome, by = c('mouse_id')) %>% 
#  filter(!is.na(summary_score)) %>% 
#  ggplot(aes(y = Log_repiricoal_dilution, x = summary_score)) + 
#    geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5,
#                 aes(fill = as.factor(summary_score))) + 
#    #stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5),
#    #             aes(color = as.factor(summary_score))) +
#    #geom_jitter(width = 0.25, alpha = 0.5) + 
#    scale_color_manual(values = donor_colors) + 
#    scale_fill_manual(values = donor_colors) + 
#    theme_bw() + 
#    scale_x_continuous(breaks = 0:10) + 
#    guides(color = 'none', fill = 'none')
#
#toxin %>% 
#  mutate(mouse_id = ifelse(toxin_sample_type == 'cecalcon', 
#                           gsub('C$', '', mouse_id), mouse_id)) %>% 
#  group_by(mouse_id) %>% 
#  mutate(last_sample = max(toxin_sample_day)) %>% 
#  filter(toxin_sample_day == last_sample) %>% 
#  left_join(select(histology, mouse_id, epithelial_damage), by = c('mouse_id')) %>% 
#  left_join(mouse_by_outcome, by = c('mouse_id')) %>% 
#  filter(!is.na(epithelial_damage)) %>% 
#  ggplot(aes(y = Log_repiricoal_dilution, x = epithelial_damage, 
#             color = as.factor(epithelial_damage))) + 
#    geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5,
#                 aes(fill = as.factor(epithelial_damage))) + 
#    #stat_summary(fun.data = 'median_hilow', fun.args = (conf.int=0.5)) +
#    #geom_jitter(width = 0.25, alpha = 0.5) + 
#    scale_color_manual(values = donor_colors) + 
#    scale_fill_manual(values = donor_colors) + 
#    theme_bw() + 
#    scale_x_continuous(breaks = 0:10) + 
#    guides(color = 'none', fill = 'none')
#
# not applicable to our story
```

Mice were challenged with a strain isolated that matched Clostridioides difficile ribotype 027

> what is published about histology of 027 in C57/B6 mice?   
> how much variation is there in individual isolates of RT027?
> RT027/BI/NAP1, toxinotype III
> VPI 10463 is ribotype 087, toxinotype 0

```{r}
#does cdiff correlate with toxin or histology?

cfu_toxin_corr <- metadata %>% 
  inner_join(toxin, by = 'group') %>% 
  filter(cdiff_strain == 431,
         cdiff_cfu != 0) %>% 
  mutate(cfu = log10(cdiff_cfu)) %>% 
	mutate(spearman_p = cor.test(cfu, Log_repiricoal_dilution, 
	                             method = 'spearman', exact = F)$p.value,
	       spearman_rho = cor.test(cfu, Log_repiricoal_dilution, 
	                               method = 'spearman', exact = F)$estimate) 
cfu_toxin_corr %>% 
  ggplot(aes(x = cfu, y = Log_repiricoal_dilution)) + 
    geom_point(alpha = 0.3) + 
    theme_bw() + 
    guides(color = 'none') + 
    geom_smooth(method = 'lm', se=F) + 
    labs(x = 'C. difficile CFU (log 10)', y = 'Toxin (log reciprocal dilution)',
         subtitle = paste0('rho = ', unique(cfu_toxin_corr$spearman_rho), 
                           ', pvalue = ', unique(cfu_toxin_corr$spearman_p)))

#hist_cfu_corr <- metadata %>% 
#  filter(day == mouse_endpoint,
#         cdiff_strain == 431) %>% 
#  left_join(histology, by = c('mouse_id')) %>% 
#  mutate(cfu = log10(cdiff_cfu),
#         score = summary_score) %>% 
#	mutate(spearman_p = cor.test(cfu, score, method = 'spearman', exact = F)$p.value,
#	       spearman_rho = cor.test(cfu, score, method = 'spearman', exact = F)$estimate)
#hist_cfu_corr %>% 
#  ggplot(aes(x = cfu, y = score)) + 
#    geom_point(alpha = 0.3) + 
#    theme_bw() + 
#    guides(color = 'none') + 
#    geom_smooth(method = 'lm', se=F) + 
#    labs(x = 'C. difficile CFU (log 10)', y = 'Histology Score',
#         subtitle = paste0('rho = ', unique(hist_cfu_corr$spearman_rho), 
#                           ', pvalue = ', unique(hist_cfu_corr$spearman_p)))
# correlation of cfu and summary score, rho = 0.15, p value = 0.29

```
Expected correlation, since no cfu == no toxin, and cfu when greater popultion will produce more toxin

Did not calculation correlation for toxin/clinical score since toxin only detected in 5 of 53 (with high clinica score) however within range of others with similar clinical score and no toxin and bias in number of samples w/o toxin  
Correlations between CFU and histology score not significant.

Same day toxin vs Genus

```{r}
#metadata %>% 
#  left_join(mutate(histology, mouse_id = paste0(cage_ID, '_', mouse_ID)) %>% select(-day),
#            by = c('mouse_id')) %>% 
#  mutate(outcome = ifelse(summary_score < 4, 'Low', outcome)) %>% 
#ggplot(aes(x = day, 
#           y = log_cfu, 
#           color = outcome, 
#           group = mouse_id)) + 
#  geom_line() + 
#  facet_wrap(outcome~., ncol = 1) +
#  theme_bw()

shared <- read_tsv(here('data/mothur/sample.final.0.03.subsample.shared'),
         col_types = cols(.default = col_double(),
                          Group = col_character())) %>% 
  mutate(Group = ifelse(grepl('_D|DA', Group), Group,
      gsub('D', '_D', Group))) %>% 
  select(-label, -numOtus) %>% 
  pivot_longer(-Group, names_to = 'otu', values_to = 'counts') %>% 
  mutate(relative_abundance = counts/2107 * 100)

# correlations
# toxin to otu
toxin_otu_corr <- metadata %>% 
  inner_join(toxin, by = 'group') %>% 
  filter(cdiff_strain == 431) %>% 
  select(Group = group, Log_repiricoal_dilution) %>% 
  left_join(shared, by = c('Group')) %>% 
  left_join(taxonomy, by = c('otu' = 'OTU')) %>% 
  filter(!is.na(otu)) %>% 
  group_by(Group, Genus, Log_repiricoal_dilution) %>%
  summarise(counts = sum(counts),
            relative_abundance = sum(relative_abundance)) %>% 
  group_by(Genus) %>% 
  mutate(otu = Genus,
        sd = sd(counts)) %>% 
  filter(sd != 0) %>% 
  nest() %>% 
	mutate(spearman_p = map(.x = data, .f = ~ cor.test(.x$counts, .x$Log_repiricoal_dilution, 
	                                                   method = 'spearman', exact = F)$p.value),
	       spearman_rho = map(.x = data, .f = ~ cor.test(.x$counts, .x$Log_repiricoal_dilution, 
	                                                     method = 'spearman', exact = F)$estimate)) %>%
	unnest(c(spearman_p, spearman_rho)) %>% 
	mutate(ncomparisons = nrow(.),
  	  pvalue = p.adjust(spearman_p, method = 'BH', n = ncomparisons)) %>% # correct p values
	filter(pvalue < 0.05) %>% # select only those above 0.05 after pvalue correction
	unnest(data)

toxin_otu_corr %>% 
  filter(spearman_rho < -0.35 | spearman_rho > 0.35) %>% 
  mutate(relative_abundance = ifelse(relative_abundance == 0, 0.045, relative_abundance)) %>% 
  ggplot(aes(x = relative_abundance, y = Log_repiricoal_dilution, color = otu)) + 
    geom_point(alpha = 0.3) + 
    theme_bw() + 
    facet_wrap(otu~., scales = 'free_x') + 
    scale_x_log10() + 
    guides(color = 'none') + 
    geom_smooth(method = 'lm', se=F) + 
    labs(x = 'Relative Abundance', y = 'Toxin (log reciprocal dilution)')
```

```{r significant_same_day_toxin, fig,height = 14}
# use lefse instead
#metadata %>% 
#  inner_join(toxin, by = 'group') %>% 
#  filter(cdiff_strain == 431) %>% 
#  mutate(toxin = Log_repiricoal_dilution > 1) %>% 
#  select(Group = group, toxin) %>% 
#  left_join(shared, by = c('Group')) %>% 
#  left_join(taxonomy, by = c('otu' = 'OTU')) %>% 
#  filter(!is.na(otu)) %>% 
#  group_by(Group, Genus, toxin) %>%
#  summarise(counts = sum(counts),
#            relative_abundance = sum(relative_abundance)) %>% 
#  group_by(Genus) %>% 
#  mutate(otu = Genus,
#        sd = sd(counts)) %>% 
#  filter(sd != 0) %>% 
#  nest() %>% 
#	mutate(wilcox_p = map(.x = data, .f = ~ wilcox.test(pull(filter(.x, toxin ==T), counts),
#    pull(filter(.x, toxin ==F), counts))$p.value)) %>%
#	unnest(wilcox_p) %>% 
#	mutate(pvalue = p.adjust(wilcox_p, method = 'BH', n = nrow(.))) %>% # correct p values
#	filter(pvalue < 0.05) %>% # select only those above 0.05 after pvalue correction
#	unnest(data) %>% 
#  group_by(Genus) %>% 
#  mutate(Genus = gsub('_unclassified', '', Genus),
#         median_ra = median(relative_abundance),
#         relative_abundance = ifelse(relative_abundance == 0, 0.045, relative_abundance)) %>% 
#  ggplot(aes(x = reorder(Genus, median_ra), y = relative_abundance, color = toxin)) + 
#    geom_boxplot() + 
#    coord_flip() + 
#    scale_y_log10() + 
#    theme_bw() + 
#    labs(x = NULL, y = 'Relative Abundance')
#
```

```{r significant_day_0_toxin, fig,height = 14}
#Significant Day 0 OTUs for communities that do/dont produce toxin
# use lefse instead
#toxin %>% 
#  group_by(mouse_id) %>% 
#  summarise(toxin = max(Log_repiricoal_dilution) > 1) %>% 
#  inner_join(filter(metadata, cdiff_strain == 431, day == 0), 
#             by = 'mouse_id') %>% 
#  select(Group = group, toxin) %>% 
#  left_join(shared, by = c('Group')) %>% 
#  left_join(taxonomy, by = c('otu' = 'OTU')) %>% 
#  filter(!is.na(otu)) %>% 
#  group_by(Group, Genus, toxin) %>%
#  summarise(counts = sum(counts),
#            relative_abundance = sum(relative_abundance)) %>% 
#  group_by(Genus) %>% 
#  mutate(otu = Genus,
#        sd = sd(counts)) %>% 
#  filter(sd != 0) %>% 
#  nest() %>% 
#	mutate(wilcox_p = map(.x = data, .f = ~ wilcox.test(pull(filter(.x, toxin ==T), counts),
#    pull(filter(.x, toxin ==F), counts))$p.value)) %>%
#	unnest(wilcox_p) %>% 
#	mutate(pvalue = p.adjust(wilcox_p, method = 'BH', n = nrow(.))) %>% # correct p values
#	filter(pvalue < 0.05) %>% # select only those above 0.05 after pvalue correction
#	unnest(data) %>% 
#  group_by(Genus) %>% 
#  mutate(Genus = gsub('_unclassified', '', Genus),
#         median_ra = median(relative_abundance),
#         relative_abundance = ifelse(relative_abundance == 0, 0.045, relative_abundance)) %>% 
#  ggplot(aes(x = reorder(Genus, median_ra), y = relative_abundance, color = toxin)) + 
#    geom_boxplot() + 
#    coord_flip() + 
#    scale_y_log10() + 
#    theme_bw() + 
#    labs(x = NULL, y = 'Relative Abundance')

```

Correlation of day 0 otus with day 10 histology scores

```{r}
# correlation of histology to otu
metadata %>% 
  filter(cdiff_strain == 431,
         day == 0) %>% 
  inner_join(histology, by = 'mouse_id') %>% 
  filter(early_euth == F) %>% 
  select(Group = group, test = summary_score) %>% 
  left_join(shared, by = c('Group')) %>% 
  group_by(otu) %>%
  mutate(sd = sd(counts)) %>% 
  filter(!is.na(otu),
         sd != 0) %>% 
  nest() %>% 
  mutate(spearman_p = map(.x = data, .f = ~ cor.test(.x$counts, .x$test, 
                                                     method = 'spearman', exact = F)$p.value),
         spearman_rho = map(.x = data, .f = ~ cor.test(.x$counts, .x$test, 
                                                       method = 'spearman', exact = F)$estimate)) %>% # compare cleared vs colonized
  unnest(c(spearman_p, spearman_rho)) %>% 
  mutate(pvalue = p.adjust(spearman_p, method = 'BH', n = nrow(.))) %>% # correct p values
  filter(pvalue < 1) %>% # select only those above 0.05 after pvalue correction
  unnest(data) %>% 
  left_join(taxonomy, by = c('otu' = 'OTU')) %>% 
  mutate(tax_otu_label = gsub('_unclassified', '', tax_otu_label),
         tax_otu_label = paste0(tax_otu_label, '\nrho = ', round(spearman_rho, 3), 
                                ', *P* = ', round(pvalue, 3))) %>%  
  ggplot(aes(x = relative_abundance + 0.001, y = test, color = tax_otu_label)) + 
    geom_point(alpha = 0.3) +
    geom_smooth(method = 'lm', se = F) +
    scale_x_log10() + 
    facet_wrap(tax_otu_label ~ ., scales = 'free_x') + 
    theme_bw() + 
    labs(x = 'Relative Abundance', y = 'Clinical Score') + 
      guides(color = 'none')

```

Correlation of day 10 otus with histology scores

```{r}
# correlation of histology to otu
day_10_hist_genus_corr_plot <- metadata %>% 
  filter(cdiff_strain == 431,
         day == mouse_endpoint) %>% 
  inner_join(histology, by = 'mouse_id') %>% 
  filter(early_euth == F) %>% 
  select(Group = group, test = summary_score) %>% 
    left_join(shared, by = c('Group')) %>% 
    group_by(otu) %>%
    mutate(sd = sd(counts)) %>% 
    filter(!is.na(otu),
           sd != 0) %>% 
    nest() %>% 
  	mutate(spearman_p = map(.x = data, .f = ~ cor.test(.x$counts, .x$test, 
  	                                                   method = 'spearman', exact = F)$p.value),
  	       spearman_rho = map(.x = data, .f = ~ cor.test(.x$counts, .x$test, 
  	                                                     method = 'spearman', exact = F)$estimate)) %>% # compare cleared vs colonized
  	unnest(c(spearman_p, spearman_rho)) %>% 
  	mutate(pvalue = p.adjust(spearman_p, method = 'BH', n = nrow(.))) %>% # correct p values
  	filter(pvalue < 0.05) %>% # select only those above 0.05 after pvalue correction
  	unnest(data) %>% 
    left_join(taxonomy, by = c('otu' = 'OTU')) %>% 
    mutate(tax_otu_label = gsub(' \\(', '\n\\(', tax_otu_label)) %>% 
    ggplot(aes(x = relative_abundance + 0.001, y = test, color = tax_otu_label)) + 
      geom_point(alpha = 0.3) +
      geom_smooth(method = 'lm', se = F) +
      scale_x_log10() + 
      facet_wrap(tax_otu_label ~ ., scales = 'free_x') + 
      theme_bw() + 
      labs(x = 'Relative Abundance', y = 'Clinical Score') + 
      guides(color = 'none')

day_10_hist_genus_corr_plot
```
```

### LEfSe

```{r, fig.height = 9, fig.width = 6}
# Figures should be no more than 6.87 inches wide (approximately 17.46 cm or 41.25 picas wide) and 9.06 inches high (approximately 23.01 cm or 54.37 picas high)

metadata <- read_tsv(here('data/process/metadata_tidy.tsv'),
		col_type = 'cccccDddddDcdddddcdl') 
shared <- read_tsv(here('data/mothur/sample.final.0.03.subsample.shared'),
		col_types = cols(.default = col_double(),
			Group = col_character())) %>% 
	mutate(Group = ifelse(grepl('_D|DA', Group), Group,
		gsub('D', '_D', Group)))
taxonomy <- read_tsv(here('data/process/final.taxonomy.tidy.tsv'),
                     col_types = cols(.default = col_character()))

lefse_design <- data.table::fread(here('data/process/lefse/toxin_severity_Genus.design'))
    colnames(lefse_design) <- c('Group', 'class')

lefse_df <- data.table::fread(here('data/process/lefse/toxin_severity_Genus.0.03.lefse_summary'),
  	fill = T) %>% 
  filter(!is.na(LDA)) %>% 
  top_n(20, LDA) %>% 
  mutate(fctr_class = factor(Class),
    order = LDA + 10 * as.numeric(fctr_class),
  	tax_otu_label = OTU)

#shared_lefse <- shared %>% 
#  gather(OTU, abundance, matches('Otu\\d*')) %>% 
#	left_join(taxonomy, by = 'OTU') %>% 
#	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
#  right_join(lefse_design, by = c('Group')) %>% 
#	group_by(Group, Genus, order, class) %>% # removed mouse_id from interactive plot
#	summarise(abundance = sum(abundance)) %>% 
#	mutate(tax_otu_label = Genus) %>% 
#	mutate(abundance = abundance / 21.07,
#		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
#  #highlight_key(~mouse_id) %>% 
#	ggplot(aes(x = reorder(tax_otu_label, order), 
#			y = abundance, color = class)) + # , shape = mouse_id remove mouse_id from interactive plot
#		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
#    stat_summary(fun.data = 'median_hilow', aes(group = class),
#			fun.args = (conf.int=0.5), position = position_dodge(width = .8)) +
#		theme_bw() + 
#    labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
#		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
#			labels=c("0","0.1","1","10","100")) +
#		#scale_color_manual(values = c('#8de4d3','#f75b5f')) + 
#    theme(panel.grid.major.y = element_blank(),
#          panel.grid.minor.y = element_line(color = 'black'),
#			legend.position = 'top') + 
#		coord_flip() + 
#  geom_jitter(position = position_jitterdodge(dodge.width = .8, jitter.width = 0.1),
#              alpha = 0.1)
#
#cfu_lefse <- metadata %>% 
#  left_join(mutate(lefse_design, mouse_id = gsub('D0', '', Group)), 
#      by = c('mouse_id')) %>% 
#  mutate(cdiff_cfu = ifelse(cdiff_cfu == 0, log10(60), log10(cdiff_cfu)),
#         xaxis = 'C. difficile') %>% 
#	ggplot(aes(x = xaxis, y = cdiff_cfu, color = class)) + 
#		geom_hline(yintercept = log10(100), linetype = 'dashed', lwd = 0.3) + 
#    stat_summary(fun.data = 'median_hilow', aes(group = class),
#			fun.args = (conf.int=0.5), position = position_dodge(width = 1)) +
#		theme_bw() + 
#    labs(x = NULL, y = 'CFU', color = NULL) + 
#		#scale_color_manual(values = c('#8de4d3','#f75b5f')) + 
#    theme(panel.grid.major.y = element_blank(), 
#			panel.grid.minor = element_blank(),
#			legend.position = 'none') + 
#		coord_flip() + 
#  geom_jitter(position = position_jitterdodge(dodge.width = 1, jitter.width = 0.2), 
#              alpha = 0.1)
#
#cowplot::plot_grid(shared_lefse, 
#                   cowplot::plot_grid(NULL, cfu_lefse, nrow = 1, rel_widths = c(1.2, 8)),
#                   ncol = 1, rel_heights = c(9,1))
#

#library(plotly)
#gg <- ggplotly(abundance_plot + geom_jitter(position = position_jitterdodge(dodge.width = 0.5)), tooltip = c('mouse_id'))
#highlight(hide_legend(gg), on = "plotly_click", off = "plotly_doubleclick", 
#           color = "black", selectize = T, opacityDim = .2,
#           selected = attrs_selected(showlegend = F, marker = list(symbol = 'X')))

```

```{r, fig.height = 9}

plot_lefse <- function(file_name){
  lefse_design <- data.table::fread(here(paste0('data/process/lefse/', file_name, '.design')))
      colnames(lefse_design) <- c('Group', 'class')

  lefse_df <- data.table::fread(here(paste0('data/process/lefse/', file_name, '.0.03.lefse_summary')),
    	fill = T) %>% 
    filter(!is.na(LDA)) %>% 
    top_n(20, LDA) %>% 
    mutate(fctr_class = factor(Class),
      order = LDA + 10 * as.numeric(fctr_class),
    	tax_otu_label = OTU)

   shared %>% 
    gather(OTU, abundance, matches('Otu\\d*')) %>% 
  	left_join(taxonomy, by = 'OTU') %>% 
  	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
    right_join(lefse_design, by = c('Group')) %>% 
  	group_by(Group, Genus, order, class) %>% # removed mouse_id from interactive plot
  	summarise(abundance = sum(abundance)) %>% 
  	mutate(tax_otu_label = Genus) %>% 
  	mutate(abundance = abundance / 21.07,
  		abundance = ifelse(abundance == 0, 0.045, abundance)) %>% 
    #highlight_key(~mouse_id) %>% 
  	ggplot(aes(x = tax_otu_label, 
  			y = abundance, color = class)) + # , shape = mouse_id remove mouse_id from interactive plot
  		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
      stat_summary(fun.data = 'median_hilow', aes(group = class),
  			fun.args = (conf.int=0.5), position = position_dodge(width = .8)) +
  		theme_bw() + 
      labs(x = NULL, y = 'Relative Abundance', color = NULL, title = file_name) + 
  		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
  			labels=c("0","0.1","1","10","100")) +
  		#scale_color_manual(values = c('#8de4d3','#f75b5f')) + 
      theme(panel.grid.major.y = element_blank(),
            panel.grid.minor.y = element_line(color = 'black'),
  			legend.position = 'top') + 
  		coord_flip() + 
    geom_jitter(position = position_jitterdodge(dodge.width = .8, jitter.width = 0.1),
                alpha = 0.1)
}
```

#### day 0 prediction of future

```{r, fig.height = 9}
plot_lefse('day_0_toxin_production_Genus')
plot_lefse('severe_disease_Genus')
#plot_lefse('toxin_severity_Genus') # decieving because communities/cages get split like (ie the one mouse that was moribund w/o toxin)

plot_lefse('summary_score_hilo_day0_Genus')
plot_lefse('day_0_severity_Genus')
```

#### same day

```{r, fig.height = 9}
plot_lefse('toxin_presence_Genus')
plot_lefse('summary_score_hilo_day10_Genus')

day_0_toxin_production_plot <- plot_lefse('day_0_toxin_production_Genus')
day_0_severity_Genus_plot <- plot_lefse('day_0_severity_Genus')

ggsave(here('results/figures/Figure_4.jpg'),
  cowplot::plot_grid(day_0_severity_Genus_plot + theme(text=element_text(size = 9)),
                     cowplot::plot_grid(day_0_toxin_production_plot + theme(text=element_text(size = 9)),
                                        day_10_hist_genus_corr_plot + theme(text=element_text(size = 9)),
                                        ncol = 1, rel_heights = c(3,2),
                                        labels = c('B', 'C')),
                     nrow = 1, 
                     labels = c('A', NULL)),
  height = 7, width = 6.875, unit = 'in')

```

```{r presentation lefse plot}
#  lefse_design <- data.table::fread(here(paste0('data/process/lefse/day_0_severity_Genus.design')))
#      colnames(lefse_design) <- c('Group', 'class')
#    
#  lefse_df <- data.table::fread(here(paste0('data/process/lefse/day_0_severity_Genus.0.03.lefse_summary')),
#    	fill = T) %>% 
#    filter(!is.na(LDA)) %>% 
#    top_n(10, LDA) %>% 
#    mutate(fctr_class = factor(Class),
#      order = LDA + 10 * as.numeric(fctr_class),
#    	tax_otu_label = OTU)
#    
#lefse_presentation_plot <- shared %>% 
#    gather(OTU, abundance, matches('Otu\\d*')) %>% 
#  	left_join(taxonomy, by = 'OTU') %>% 
#  	right_join(lefse_df, by = c('Genus' = 'OTU')) %>% 
#    right_join(lefse_design, by = c('Group')) %>% 
#  	group_by(Group, Genus, order, class) %>% # removed mouse_id from interactive plot
#  	summarise(abundance = sum(abundance)) %>% 
#  	mutate(tax_otu_label = Genus) %>% 
#  	mutate(tax_otu_label = gsub('_', ' ', tax_otu_label),
#  	       tax_otu_label = paste0('*', tax_otu_label, '*'),
#  	       abundance = abundance / 21.07,
#  		     abundance = ifelse(abundance == 0, 0.035, abundance),
#  		     class = case_when(class == 'mild' ~ 'Low Clinical Score',
#  		                         class == 'moderate' ~ 'High Clinical Score',
#  		                         class == 'severe' ~ 'Severe'),
#  		     class = factor(class, levels = c('Low Clinical Score', 'High Clinical Score', 'Severe'))) %>% 
#  	ggplot(aes(x = tax_otu_label, y = abundance, color = class)) +
#      stat_summary(fun.data = 'median_hilow', aes(group = class),
#  			fun.args = (conf.int=0.5), position = position_dodge(width = .4)) +
#  		geom_hline(yintercept = 0.047, linetype = 'dashed', lwd = 0.3) + 
#  		theme_bw() + 
#      labs(x = NULL, y = 'Relative Abundance', color = NULL) + 
#  		scale_y_log10(breaks=c(0.01, 0.1, 1, 10, 100),
#  			labels=c("0","0.1","1","10","100")) +
#	    scale_x_discrete(guide = guide_axis(angle = 45)) + 
#  		scale_color_manual(values = c("#62ebc9", "#399283", "#0c1b37")) + 
#      theme(panel.grid.minor.y = element_blank(),
#  			    legend.position = 'top',
#            axis.text.x = ggtext::element_markdown())
#
#ggsave('~/Desktop/lefse_presentation_plot.jpg',
#       lefse_presentation_plot,
#       width = 9, height = 5)
```

What is the genus Clostridioides made up of?
Only stool?
toxin + disease mild = limited access to epithelial

Microbe-microbe Interactions during Clostridioides difficile Infection
  Arwa Abbas and Joseph P. Zackular. Curr Opin Microbiol. 2020 Feb; 53: 19–25. doi: 10.1016/j.mib.2020.01.016

    C. difficile infection (CDI) can cause a spectrum of disease from mild diarrhea to severe complications such as pseudomembranous colitis, toxic megacolon and death (reviewed in [3]). This broad spectrum of C. difficile-associated disease may be explained in part by bacterial genetic factors such as variation in the pathogenicity locus [4] and increased accessory gene content [5,6]. There is also likely a strong role for exogenous factors such as host genetics, comorbidities, treatment modalities and previous drug exposures. Here we focus on how resident microbiota can manipulate pathogen behavior and virulence.
        competition for essential nutrients, limiting access to mucosal surfaces, direct production of antimicrobial molecules, modulating the intestinal metabolome, and activating the host immune system against the pathogen of interest
    Exposure to low concentrations of deoxycholate, one of the most abundant cecal bile acids [39,40], reduced toxin production by most strains, without a concomitant reduction in general vegetative cell growth 
    B. thetatiotaomicron cross-feeds sialic acid to C. difficile
    succinate appears necessary for C. difficile expansion in the gut
    inhibit via production of bile acids, antibiotics, acidification
    lacto-conditioned media reduce toxin and ability to adhere to host cells
    Enterobacteriaceae and Enterococcus are known to thrive during intestinal inflammation and during CDI 
    Cdiff induces indole production via E. coli which inhibits anaerobes
    Cdiff ferments tyrosine to p-cresol
    C. difficile can also secrete proline-based cyclic dipeptides that can inhibit gut bacteria, including commensal Clostridium species

microbiome/Cdifficile/host interactions
  lactobacillus inhibits cdiff
  biofilm can protect cdiff
  microbiome can secrete inhibitory compounds - turbomycin, lantibiotics
  microbiome can cross feed microbiome via sialic acid/succinate or even proline or AAs
  cdiff can inhibit microbiome w/ indole or para-cresol