### 2022-04-22 results/tables/header.tex
update manuscript w/revisions

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\pagenumbering{gobble}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\usepackage[labelformat=empty]{caption} % Table caption label NULL
\usepackage[textfont=bf]{caption} % Table caption text bold
\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}