### 2022-04-22 results/tables/Table_S1.Rmd

---
title: ""
output:
  pdf_document:
    keep_tex: false
    includes:
      in_header: header.tex
geometry: margin=1.0in
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE}
######################################################################
#----------------- Read in necessary libraries -------------------#
######################################################################
deps = c("knitr","rmarkdown","tidyverse");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
######################################################################
#-------------- Define knitr options to allow kable ----------------#
######################################################################
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
```

```{r table_2, echo=FALSE, results='asis'}
######################################################################
# We load a human_source_tidy.tsv file to make Table S1
# This file is generated from subsetting MIMARKS_cdclinical.xlsx 
# from Schubert et al. mBio 2014 in code/tidy_raw_data.R
######################################################################

human_src_meta <- read_tsv(here::here('data/process/human_source_tidy.tsv'),
                           col_types = cols(.default = col_character())) %>% 
    select(plot_labels, everything()) %>%
    rename_if(is.character,
           ~ c("Sample Label", "Subject ID", "Subject Location", 
                     "Sample Medium", "Subject Age", "Subject Sex", 
                     "Recent Antibiotic Use", "Proton Pump Inhibitor Use",
                     "H2 Receptor Blocker Use", "Antacid Use", "Healthcare Worker",
                     "History of CDI", "Surgery within 6 Months", "Vegetarian",
                     "Weight", "Disease Status")) %>% 
    arrange(`Sample Label`)

#------------------------Table S1 generation------------------------#
TableS1 <- kable(human_src_meta,
                "latex",
                caption = "Table S1. Demographic information of subjects whose stool samples used to colonize germ-free mice.",
                booktabs = T, 
                linesep = "") %>%
  # latex options make table to be condensed
  # table itseld is left aligned
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),
                position = "left") %>%
  # Adjust header widths
  column_spec(1, width = '4em') %>%
  column_spec(3, width = '5em') %>%
  column_spec(4, width = '7em') %>%
  column_spec(5, width = '3em') %>%
  column_spec(6, width = '5em') %>%
  column_spec(7, width = '5em') %>%
  column_spec(8, width = '5em') %>%
  column_spec(9, width = '5em') %>%
  column_spec(10, width = '5em') %>%
  column_spec(11, width = '5em') %>%
  column_spec(12, width = '5em') %>%
  column_spec(13, width = '5em') %>%
  # Add bold header
  row_spec(0,bold=TRUE)

TableS1
```